### 2014-05-11 paper/proposal.tex
first lines of definition of the problem

\documentclass[11pt]{artikel3}
\usepackage{fullpage, setspace, graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{times}

\title{UCD
\\MSc Thesis Proposal:\\\emph{Real-time System For Continuous Capacity Planning in Cloud Environment With Dynamic Pricing}}
\author{Kamil Smuga}
\date{\today}

\begin{document}
\maketitle

\section{Problem}

New IaaS solutions are available on the market with growing number of supported pricing models. This makes the offering attractive enough to move entire businesses to the Cloud. The availability of different pricing schemes reveals new challenges in already unpredictive Capacity Planning process. Is it better to pay some amount upfront and less hourly rate? Is is better to invest in less heavy utilized instances or maybe in more medium ones? When exactly one pricing scheme becomes more attractive?

\begin{itemize}
  \item For a thesis, a hypothesis (`thesis statement') that you will test in your research.
  \item For a project, identify the work required (e.g. implementation and/or experiment) that needs to be completed. If you are completing a project, make sure to speak with your advisor about the expected deliverables; this will include a written project report.
\end{itemize}

{\bf Motivate your problem.} 
\begin{itemize}
  \item What is the significance of your problem? 
  \item What applications or new opportunities will solving your problem provide?
\end{itemize}

{\bf Related Work.}
%Demonstrate the connection between your chosen problem and its foundations in existing work. 
\begin{itemize}
  \item What are the key theoretical models (e.g. process-based, formal language/complexity models, probability-based) and algorithms have been applied toward this problem previously? 
  \item What limitation and/or opportunity do you plan to address in your project/thesis?
  \item
In the related research literature, how is success measured (e.g. metrics and/or coverage of problem aspects)?
\end{itemize}

\section{Methodology}

\begin{itemize}
\item For projects, which libraries or software tools will be used for development, and at the highest level, what is the software design?
\item For theory-based projects and theses, what are the key theorems to be developed and/or proven? What proof techniques will be used?
\item For other theses, what algorithms will be adapted or devised, and what algorithms will they be compared with in your experiments?
\end{itemize}

\section{Evaluation}

\begin{itemize}
\item What data and software will be needed for your evaluation? 
\item What metrics
will you use to measure success? Commonly these include some subset of time, space, and accuracy (recognition rate, precision, recall, etc.).
Almost always, this should include reference to the evaluation methods described in the related work.
\item For empirical theses, an experiment is needed: which algorithms will be compared, and using what sets of parameters for each? If people are involved in the experiments, how will the experiment control for unwanted bias or confounds? {\bf How does the experiment test the hypothesis?}
\item {\bf How will you know when you are done?}
\end{itemize}

\section{Evaluation Outcomes}

\begin{itemize}
\item For your chosen assessment methods, what are the possible outcomes? 
\item Under which outcomes are the project goals achieved, or the hypothesis confirmed or rejected?
\end{itemize}

\bibliographystyle{plain}
\bibliography{preproposal}

**(Omitted) As an exercise, modify this document to include the references in the {\tt plain.bib} file.

\end{document}

### 2014-05-11 paper/proposal.tex
motiviation in progress

@@
 The availability of different pricing schemes reveals new challenges in already unpredictive Capacity Planning process.
-Is it better to pay some amount upfront and less hourly rate?
+Is it better to pay some amount of money upfront and less hourly rate?
 Is is better to invest in less heavy utilized instances or maybe in more medium ones?
@@
 \begin{itemize}
-  \item For a thesis, a hypothesis (`thesis statement') that you will test in your research.
-  \item For a project, identify the work required (e.g.
-implementation and/or experiment) that needs to be completed.
-If you are completing a project, make sure to speak with your advisor about the expected deliverables; this will include a written project report.
+  \item Thesis statement: An analysis of current utilization of the infrastructure in terms of price effectiveness in dynamic pricing IaaS model.
+A calculation of optimal configuration and adjustements based on recent data and future trends calculated from the load.
+  \item Project work: The system will accept incoming data from the IaaS provider with utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
+The data will be stored, analysed in terms of price optimality and graphed on capacity planning dashboard.
 \end{itemize}
 
-{\bf Motivate your problem.} 
+{\bf Motivation.} 
 \begin{itemize}

### 2014-05-11 paper/proposal.tex
motivation finished

@@
   \item Project work: The system will accept incoming data from the IaaS provider with utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
-The data will be stored, analysed in terms of price optimality and graphed on capacity planning dashboard.
+The data will be stored, analysed in terms of price optimality (subject for research) and graphed on capacity planning dashboard.
 \end{itemize}
@@
 \begin{itemize}
-  \item What is the significance of your problem? 
-  \item What applications or new opportunities will solving your problem provide?
+  \item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either gather milions of users.
+Although the problem of scaling is hard enough by itself, a reality with limited budgets and business sustainability requires to tackle the scaling problem from the most optimal, cost-effective angle.
+Capacity planning can make a difference between making or losing the money in a dynamic, price variable environment.
+Based on a dynamic nature of bussiness size and external factors like pricing, the capacity planning needs to be continously adjusted and become a routine for daily operations.
+  \item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing and provider choice possible in an educated and more  confident manner. 
 \end{itemize}

### 2014-05-11 paper/proposal.tex
motivation

@@
   \item Thesis statement: An analysis of current utilization of the infrastructure in terms of price effectiveness in dynamic pricing IaaS model.
-A calculation of optimal configuration and adjustements based on recent data and future trends calculated from the load.
+A calculation of optimal configuration and adjustments based on recent data and future trends calculated from the load.
   \item Project work: The system will accept incoming data from the IaaS provider with utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
@@
 \begin{itemize}
-  \item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either gather milions of users.
+  \item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either gather millions of users.
 Although the problem of scaling is hard enough by itself, a reality with limited budgets and business sustainability requires to tackle the scaling problem from the most optimal, cost-effective angle.
 Capacity planning can make a difference between making or losing the money in a dynamic, price variable environment.
-Based on a dynamic nature of bussiness size and external factors like pricing, the capacity planning needs to be continously adjusted and become a routine for daily operations.
+Based on a dynamic nature of business size and external factors like pricing, the capacity planning needs to be continuously adjusted and become a routine for daily operations.
   \item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing and provider choice possible in an educated and more  confident manner. 

### 2014-05-14 paper/proposal.tex
methodology added

@@
 
-\title{UCD
-\\MSc Thesis Proposal:\\\emph{Real-time System For Continuous Capacity Planning in Cloud Environment With Dynamic Pricing}}
+\title{UCD\\MSc Thesis Proposal:\\\emph{Real-time Framework For Continuous Capacity Planning in Cloud Environment With Dynamic Pricing}}
 \author{Kamil Smuga}
@@
   \item Project work: The system will accept incoming data from the IaaS provider with utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
-The data will be stored, analysed in terms of price optimality (subject for research) and graphed on capacity planning dashboard.
+The data will be stored, analysed in terms of price optimality (subject for research) and graphed on a capacity planning dashboard.
 \end{itemize}
@@
   \item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either gather millions of users.
-Although the problem of scaling is hard enough by itself, a reality with limited budgets and business sustainability requires to tackle the scaling problem from the most optimal, cost-effective angle.
+Although the problem of scaling is hard enough by itself, the reality with limited budgets and business sustainability requires to tackle the scaling problem from the most optimal, cost-effective angle.
 Capacity planning can make a difference between making or losing the money in a dynamic, price variable environment.
-Based on a dynamic nature of business size and external factors like pricing, the capacity planning needs to be continuously adjusted and become a routine for daily operations.
+Based on a dynamic nature of business size and external factors like pricing, a capacity planning and management need to be continuously adjusted and become an integral part of daily operations.
   \item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing and provider choice possible in an educated and more  confident manner.
+
 \end{itemize}
@@
 %Demonstrate the connection between your chosen problem and its foundations in existing work.
-\begin{itemize}
-  \item What are the key theoretical models (e.g. process-based, formal language/complexity models, probability-based) and algorithms have been applied toward this problem previously? 
-  \item What limitation and/or opportunity do you plan to address in your project/thesis?
-  \item
-In the related research literature, how is success measured (e.g. metrics and/or coverage of problem aspects)?
-\end{itemize}
+
+Most of the work in Capacity Planning for Cloud field is conducted from IaaS provider perspective.
+This paper will focus on SaaS provider who is charged on VM uptime rather than cpu, memory, i/o utlization.
+Moreover, VM uptime price varies based on renting strategy.
+This specific problem seems to be somewhat unexplored in the research area.
 
@@
 \begin{itemize}
-\item For projects, which libraries or software tools will be used for development, and at the highest level, what is the software design?
-\item For theory-based projects and theses, what are the key theorems to be developed and/or proven? What proof techniques will be used?
-\item For other theses, what algorithms will be adapted or devised, and what algorithms will they be compared with in your experiments?
+\item High level software design:
+  \begin{itemize}
+   \item Input: Uptime and pricing scheme (on-demand, light, medium, heavy) metrics from the host.
+   \item Analysis: An algorithm that takes the data and indicates how optimal the usage is for a given pricing configuration.
+   \item Output: Send data to graphing system to create a dashboard that shows current usage and optimality ratio
+  \end{itemize}
 \end{itemize}

### 2014-06-16 paper/proposal.tex
fixed intendation
initial bibliography paragraph

@@
 
-Most of the work in Capacity Planning for Cloud field is conducted from IaaS provider perspective.
+Most of the work in Capacity Planning in cloud field is conducted from IaaS provider perspective (\bibitem{bib:join_pricing}).
 This paper will focus on SaaS provider who is charged on VM uptime rather than cpu, memory, i/o utlization.
@@
 \bibliography{preproposal}
+  \bibitem{bib:join_pricing} Ling Tang Sch.
+of Comput.
+Sci.
+& Eng., Nanjing Univ.
+of Sci.
+& Technol., Nanjing, China Jinghui Qian ; Lei Xu ; Yu Yan.
+Joint pricing and capacity planning for IaaS cloud.
+ICOIN, pages 34 - 39, 2014
+  \end{preproposal}
 

### 2014-06-18 paper/proposal.tex
proposal revised

@@
 
-\title{UCD\\MSc Thesis Proposal:\\\emph{Real-time Framework For Continuous Capacity Planning in Cloud Environment With Dynamic Pricing}}
+\title{UCD\\MSc Thesis Proposal:\\\emph{A Near Real Time System For Continuous Capacity Planning in Cloud Environment With Dynamic Pricing Based On Uptime}}
 \author{Kamil Smuga}
@@
 
-New IaaS solutions are available on the market with growing number of supported pricing models.
-This makes the offering attractive enough to move entire businesses to the Cloud.
-The availability of different pricing schemes reveals new challenges in already unpredictive Capacity Planning process.
+New IaaS solutions are becoming available with a growing number of supported pricing models.
+This makes the offering attractive enough to move the entire businesses to the Cloud.
+The availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive Capacity Planning process.
 Is it better to pay some amount of money upfront and less hourly rate?
@@
 \begin{itemize}
-  \item Thesis statement: An analysis of current utilization of the infrastructure in terms of price effectiveness in dynamic pricing IaaS model.
-A calculation of optimal configuration and adjustments based on recent data and future trends calculated from the load.
-  \item Project work: The system will accept incoming data from the IaaS provider with utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
-The data will be stored, analysed in terms of price optimality (subject for research) and graphed on a capacity planning dashboard.
+  \item Thesis statement: An analysis of current utilization of the infrastructure in terms of price effectiveness in a dynamic pricing IaaS model based on uptime.
+Estimation of an optimal configuration and possible adjustments based on a recent data.
+  \item Project work: The system will accept incoming data from an IaaS provider - utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
+The data will be stored, analysed in terms of price optimality and graphed on a capacity planning dashboard.
 \end{itemize}
@@
 \begin{itemize}
-  \item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either gather millions of users.
-Although the problem of scaling is hard enough by itself, the reality with limited budgets and business sustainability requires to tackle the scaling problem from the most optimal, cost-effective angle.
+  \item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either attract millions of users.
+Although the problem of scaling is hard enough by itself, the reality with limited budgets requires to tackle the scaling problem from the most optimal, cost-effective angle.
 Capacity planning can make a difference between making or losing the money in a dynamic, price variable environment.
-Based on a dynamic nature of business size and external factors like pricing, a capacity planning and management need to be continuously adjusted and become an integral part of daily operations.
-  \item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing and provider choice possible in an educated and more  confident manner.
+Based on a dynamic nature of business size and external factors like pricing, capacity planning and provisioning need to be continuously adjusted and becomes an integral part of daily operations.
+  \item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing scheme and choice of the provider possible in an educated and more confident manner.
 
@@
 
-Most of the work in Capacity Planning in cloud field is conducted from IaaS provider perspective (\bibitem{bib:join_pricing}).
-This paper will focus on SaaS provider who is charged on VM uptime rather than cpu, memory, i/o utlization.
-Moreover, VM uptime price varies based on renting strategy.
+  Most of the work in Cloud Capacity Planning field is conducted from IaaS provider perspective (\cite{bib:join_pricing}, \cite{bib:self_adaptive}).
+The papers investigate how to effectively and efficently provide an infrastructure services to the customers.
+  There is also a work that is focusing on IaaS consumer (\cite{bib:cost-aware-elasticity}, \cite{bib:elastic-capacity-planning}).
+These papers focus on finding an optimal configuration of resources considering cpu, memory, i/o utilisation.
+  This paper will focus on SaaS provider who is charged on VM uptime - a new pricing model - rather than purely based on cpu, memory, i/o utlization.
+Moreover, VM uptime price varies based on a renting strategy.
 This specific problem seems to be somewhat unexplored in the research area.
@@
    \item Input: Uptime and pricing scheme (on-demand, light, medium, heavy) metrics from the host.
-   \item Analysis: An algorithm that takes the data and indicates how optimal the usage is for a given pricing configuration.
-   \item Output: Send data to graphing system to create a dashboard that shows current usage and optimality ratio
+   \item Analysis: An algorithm that calculates the optimality factor for a given VM based on uptime.
+   \item Output: Send data to graphing system to create a dashboard that shows current usage and optimality ratio.
   \end{itemize}
@@
 
-\section{Evaluation}
-
-\begin{itemize}
-  \item What data and software will be needed for your evaluation? 
-  \item What metrics
-will you use to measure success? Commonly these include some subset of time, space, and accuracy (recognition rate, precision, recall, etc.).
-Almost always, this should include reference to the evaluation methods described in the related work.
-  \item For empirical theses, an experiment is needed: which algorithms will be compared, and using what sets of parameters for each?
-If people are involved in the experiments, how will the experiment control for unwanted bias or confounds?
-{\bf How does the experiment test the hypothesis?}
-  \item {\bf How will you know when you are done?}
-\end{itemize}
-
-\section{Evaluation Outcomes}
-
-\begin{itemize}
-  \item For your chosen assessment methods, what are the possible outcomes? 
-  \item Under which outcomes are the project goals achieved, or the hypothesis confirmed or rejected?
-\end{itemize}
-
-\bibliographystyle{plain}
-\bibliography{preproposal}
+\begin{thebibliography}{50}
   \bibitem{bib:join_pricing} Ling Tang Sch.
@@
 & Technol., Nanjing, China Jinghui Qian ; Lei Xu ; Yu Yan.
-Joint pricing and capacity planning for IaaS cloud.
-ICOIN, pages 34 - 39, 2014
-  \end{preproposal}
-
-**(Omitted) As an exercise, modify this document to include the references in the {\tt plain.bib} file.
+\textsl{Joint pricing and capacity planning for IaaS cloud}.
+ICOIN, pages 34-39, 2014.
+  \bibitem{bib:self_adaptive}  Yexi Jiang; Chang-shing Perng; Tao Li; Rong Chang.
+\textsl{Self-Adaptive Cloud Capacity Planning}.
+IEEE Ninth International Conference on Services Computing (SCC), pages 73-80, 2012.
+  \bibitem{bib:cost-aware-elasticity} Upendra Sharma, Prashant Shenoy, Sambit Sahu, Anees Shaikh.
+\textsl{Kingfisher: Cost-aware Elasticity in the Cloud}.
+INFOCOM, 2011 Proceedings IEEE, pp.206-210, April 2011.
+  \bibitem{bib:elastic-capacity-planning} Cheng Tian; Ying Wang; Bo Yin.\textsl{An elastic capacity planning method in cloud}. ICCSE, pages 239-244, 2012.
+\end{thebibliography)
 

### 2014-06-20 paper/proposal.tex
project tasks added

@@
 
-\title{UCD\\MSc Thesis Proposal:\\\emph{A Near Real Time System For Continuous Capacity Planning in Cloud Environment With Dynamic Pricing Based On Uptime}}
+\title{UCD\\MSc Thesis Proposal:\\\emph{An Approach To Continuous Capacity Planning In Cloud Environment With Uptime Based Pricing Model}}
 \author{Kamil Smuga}
@@
 This makes the offering attractive enough to move the entire businesses to the Cloud.
-The availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive Capacity Planning process.
+The availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive capacity planning process.
 Is it better to pay some amount of money upfront and less hourly rate?
 Is is better to invest in less heavy utilized instances or maybe in more medium ones?
-When exactly one pricing scheme becomes more attractive?
+When exactly one pricing scheme becomes more cost effective?
 
@@
 
+  \section{Project tasks}
+
+  \begin{itemize}
+  \item Mandatory:
+    \begin{itemize}
+      \item A formal definition of continous capacity planning approach.
+      \item A design of an algorithm that calculates an uptime optimality factor of VM utilisation based on a pricing scheme.
+    \end{itemize}
+
+  \item Discretionary
+    \begin{itemize}
+      \item An integration with an IaaS provider's API to feed data into the system.
+      \item A graphical representation of results calculated by the algorithm.
+    \end{itemize}
+
+  \item Exceptional
+    \begin{itemize}
+      \item Generate suggestions to change the configuration based on the calculated results.
+    \end{itemize}
+
 \begin{thebibliography}{50}

### 2014-06-30 paper/proposal.tex
wording

@@
 New IaaS solutions are becoming available with a growing number of supported pricing models.
-This makes the offering attractive enough to move the entire businesses to the Cloud.
-The availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive capacity planning process.
-Is it better to pay some amount of money upfront and less hourly rate?
-Is is better to invest in less heavy utilized instances or maybe in more medium ones?
-When exactly one pricing scheme becomes more cost effective?
+More often than not, a hosted Cloud environment is used to design and build an infrastructure for a product.
+The recent availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive capacity planning process.
+When exactly does one pricing scheme becomes more cost effective?
 
@@
 \begin{itemize}
-      \item A formal definition of continous capacity planning approach.
-      \item A design of an algorithm that calculates an uptime optimality factor of VM utilisation based on a pricing scheme.
+\item A definition of continous capacity planning approach.
+\item A design of an algorithm that calculates VM utilisation optimality factor based on a pricing scheme.
 \end{itemize}
@@
 \begin{itemize}
-      \item Generate suggestions to change the configuration based on the calculated results.
+\item Suggestions generation to change the configuration based on the calculated results.
+\item An improvement to the algorithm that considers expotential discounting to help to make a cost effective choice over time.
+\item Calculate trends based on historical data and suggest and an optimal configuration for this prediction.
 \end{itemize}

### 2014-06-30 paper/proposal.tex
wording

@@
 The recent availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive capacity planning process.
+One can choose between ad hoc provisioning vs upfront payments and reduced hourly rates.
+Reserved instances charged upfront are splitted into three groups: light, medium and heavy.
+Which one is better for a given utilisation model?
 When exactly does one pricing scheme becomes more cost effective?
@@
 \begin{itemize}
-\item New startups often attempt to solve globe wide problems and have to either process increasing amounts of data either attract millions of users.
-Although the problem of scaling is hard enough by itself, the reality with limited budgets requires to tackle the scaling problem from the most optimal, cost-effective angle.
-Capacity planning can make a difference between making or losing the money in a dynamic, price variable environment.
-Based on a dynamic nature of business size and external factors like pricing, capacity planning and provisioning need to be continuously adjusted and becomes an integral part of daily operations.
+\item
+
+Capacity planning can make a difference between earning and losing money.
+Considering a dynamic nature of business and external factors like pricing, capacity planning and provisioning need to be continuously adjusted and become an integral part of daily operations.
+Having a granual data about infrastructure utilisation is fundamental to make adjustments when it is needed.
+Frequency of adjustments may vary, however, it is important to have this possibility at any time.
 \item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing scheme and choice of the provider possible in an educated and more confident manner.

### 2014-06-30 paper/proposal.tex
changes from Christina

@@
 
-\title{UCD\\MSc Thesis Proposal:\\\emph{An Approach To Continuous Capacity Planning In Cloud Environment With Uptime Based Pricing Model}}
+\title{UCD\\MSc Thesis Proposal:\\\emph{An approach for Continuous Capacity Planning in Cloud Environments with Uptime-based Pricing Model}}
 \author{Kamil Smuga}
@@
 
-New IaaS solutions are becoming available with a growing number of supported pricing models.
+New Infrastructure as a Service solutions are becoming available with a growing number of supported pricing models.
 More often than not, a hosted Cloud environment is used to design and build an infrastructure for a product.
-The recent availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictive capacity planning process.
-One can choose between ad hoc provisioning vs upfront payments and reduced hourly rates.
-Reserved instances charged upfront are splitted into three groups: light, medium and heavy.
+The recent availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictable capacity planning process.
+There is a choice between ad-hoc provisioning and upfront payments with reduced hourly rates.
+Reserved instances charged upfront are categorised into three groups: light, medium and heavy.
 Which one is better for a given utilisation model?
 When exactly does one pricing scheme becomes more cost effective?
+Determining which machine type is better for a given utilisation model, or at which point the cost effectiveness of a pricing scheme changes, is vital for the companies subscribing to the IaaS.
 
 \begin{itemize}
-\item Thesis statement: An analysis of current utilization of the infrastructure in terms of price effectiveness in a dynamic pricing IaaS model based on uptime.
-Estimation of an optimal configuration and possible adjustments based on a recent data.
+\item Thesis statement: This thesis will focus on analyzing a system infrastructure utilisation in terms of price effectiveness with uptime-based pricing model.
 \item Project work: The system will accept incoming data from an IaaS provider - utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
@@
 Capacity planning can make a difference between earning and losing money.
-Considering a dynamic nature of business and external factors like pricing, capacity planning and provisioning need to be continuously adjusted and become an integral part of daily operations.
-Having a granual data about infrastructure utilisation is fundamental to make adjustments when it is needed.
+Considering the dynamic nature of business and external factors like pricing, capacity planning and provisioning need to be continuously adjusted and become an integral part of daily operations.
+Having a granular data about infrastructure utilisation is fundamental to make adjustments when it is needed.
 Frequency of adjustments may vary, however, it is important to have this possibility at any time.
-\item Solving the problem of visualisation and optimisation of infrastructure costs will make decisions about pricing scheme and choice of the provider possible in an educated and more confident manner.
+\item Solving the problem of visualisation and optimisation of infrastructure costs will facilitate more informed and effective pricing scheme and provider selection decisions.
 
@@
 Most of the work in Cloud Capacity Planning field is conducted from IaaS provider perspective (\cite{bib:join_pricing}, \cite{bib:self_adaptive}).
-The papers investigate how to effectively and efficently provide an infrastructure services to the customers.
+The papers investigate how to effectively and efficiently provide an infrastructure services to the customers.
 There is also a work that is focusing on IaaS consumer (\cite{bib:cost-aware-elasticity}, \cite{bib:elastic-capacity-planning}).
 These papers focus on finding an optimal configuration of resources considering cpu, memory, i/o utilisation.
-This paper will focus on SaaS provider who is charged on VM uptime - a new pricing model - rather than purely based on cpu, memory, i/o utlization.
+This paper will focus on SaaS provider who is charged on VM uptime - a new pricing model - rather than purely based on cpu, memory, i/o utilization.
 Moreover, VM uptime price varies based on a renting strategy.
@@
 \begin{itemize}
-\item A definition of continous capacity planning approach.
+\item A definition of continuous capacity planning approach.
 \item A design of an algorithm that calculates VM utilisation optimality factor based on a pricing scheme.
@@
 \item Suggestions generation to change the configuration based on the calculated results.
-\item An improvement to the algorithm that considers expotential discounting to help to make a cost effective choice over time.
+\item An improvement to the algorithm that considers exponential discounting to help to make a cost effective choice over time.
 \item Calculate trends based on historical data and suggest and an optimal configuration for this prediction.
@@
 of Sci.
-& Technol., Nanjing, China Jinghui Qian ; Lei Xu ; Yu Yan.
-\textsl{Joint pricing and capacity planning for IaaS cloud}.
+& Technol., Nanjing, China Jinghui Qian ; Lei Xu ; Yu Yan.\textsl{Joint pricing and capacity planning for IaaS cloud}.
 ICOIN, pages 34-39, 2014.

### 2014-08-07 paper/proposal.tex
latex syntax corrections
A bit of more detailed explanation in proposal

@@
 More often than not, a hosted Cloud environment is used to design and build an infrastructure for a product.
-The recent availability of different pricing schemes based on resource utilisation and uptime reveals new challenges in already unpredictable capacity planning process.
+The recent availability of different pricing schemes based on resource utilization and uptime reveals new challenges in already unpredictable capacity planning process.
 There is a choice between ad-hoc provisioning and upfront payments with reduced hourly rates.
-Reserved instances charged upfront are categorised into three groups: light, medium and heavy.
-Which one is better for a given utilisation model?
+Reserved instances charged upfront are categorized into three groups: light, medium and heavy.
+Which one is better for a given utilization model?
 When exactly does one pricing scheme becomes more cost effective?
-Determining which machine type is better for a given utilisation model, or at which point the cost effectiveness of a pricing scheme changes, is vital for the companies subscribing to the IaaS.
+Determining which machine type is better for a given utilization model, or at which point the cost effectiveness of a pricing scheme changes, is vital for the companies subscribing to the IaaS.
 
 \begin{itemize}
-\item Thesis statement: This thesis will focus on analyzing a system infrastructure utilisation in terms of price effectiveness with uptime-based pricing model.
-\item Project work: The system will accept incoming data from an IaaS provider - utilisation metrics and a number of running instances for a given type (light, medium, heavy, on demand).
-The data will be stored, analysed in terms of price optimality and graphed on a capacity planning dashboard.
+\item Thesis statement: This thesis will focus on analyzing a system infrastructure utilization in terms of price effectiveness with uptime-based pricing model.
+This work will emphasis an importance of knowledge of distributed system state to allow making optimal changes and plan for future.
+\item Project work: The system will accept incoming data from an IaaS provider - utilization metrics and a number of running instances for a given type (light, medium, heavy, on demand).
+The data will be stored, analyzed in terms of price optimality and graphed on a capacity planning dashboard.
 \end{itemize}
@@
 \begin{itemize}
-\item
-
-Capacity planning can make a difference between earning and losing money.
+\item Capacity planning can make a difference between earning and losing money.
 Considering the dynamic nature of business and external factors like pricing, capacity planning and provisioning need to be continuously adjusted and become an integral part of daily operations.
-Having a granular data about infrastructure utilisation is fundamental to make adjustments when it is needed.
+Having a granular data about infrastructure utilization is fundamental to make adjustments when it is needed.
 Frequency of adjustments may vary, however, it is important to have this possibility at any time.
-\item Solving the problem of visualisation and optimisation of infrastructure costs will facilitate more informed and effective pricing scheme and provider selection decisions.
+\item Solving the problem of visualization and optimization of infrastructure costs will facilitate more informed and effective pricing scheme and provider selection decisions.
+\item It is likely that an IaaS user over-provisions an infrastructure to handle all traffic.
+This creates a scenario where it is desired to know the state of a configuration and optimize if possible.
+Moreover, changes to existing configuration will affect optimality factors and require to be re-calculated systematically.
 
@@
 There is also a work that is focusing on IaaS consumer (\cite{bib:cost-aware-elasticity}, \cite{bib:elastic-capacity-planning}).
-These papers focus on finding an optimal configuration of resources considering cpu, memory, i/o utilisation.
+These papers focus on finding an optimal configuration of resources considering cpu, memory, i/o utilization.
 This paper will focus on SaaS provider who is charged on VM uptime - a new pricing model - rather than purely based on cpu, memory, i/o utilization.
@@
 \begin{itemize}
-\item Input: Uptime and pricing scheme (on-demand, light, medium, heavy) metrics from the host.
-\item Analysis: An algorithm that calculates the optimality factor for a given VM based on uptime.
-\item Output: Send data to graphing system to create a dashboard that shows current usage and optimality ratio.
+\item Input per single host: 
+	\begin{itemize}
+	\item Uptime - metric to report how long the system is up.
+By default, it is gathered as an aggregation of total runtime since reboot.
+Post-processing of this information will be done to distribute total number of running hours per day to align with pricing model.
+	\item Pricing scheme - one of: on-demand, light, medium, heavy - a property of reserved machine that pricing is based on.
+	\end{itemize}
+\item Analysis: An algorithm that calculates an optimality factor for a given host based on a daily value of Uptime and choice of Pricing scheme.
+\item Output: Algorithm results to be send to a graphing system to create dashboards to show current state and optimality ratio.
 \end{itemize}
@@
 \begin{itemize}
-\item A definition of continuous capacity planning approach.
-\item A design of an algorithm that calculates VM utilisation optimality factor based on a pricing scheme.
+\item A definition of continuous capacity planning approach. Explanation why granular knowledge about state of running system is fundamental to make educated choices over time. 
+\item A design of an algorithm that calculates VM utilization optimality factor based on uptime and a pricing scheme.
 \end{itemize}
@@
 \end{itemize}
+\end{itemize}
 
@@
 Sci.
-& Eng., Nanjing Univ.
+\& Eng., Nanjing Univ.
 of Sci.
-& Technol., Nanjing, China Jinghui Qian ; Lei Xu ; Yu Yan.\textsl{Joint pricing and capacity planning for IaaS cloud}.
+\& Technol., Nanjing, China Jinghui Qian ; Lei Xu ; Yu Yan.\textsl{Joint pricing and capacity planning for IaaS cloud}.
 ICOIN, pages 34-39, 2014.
@@
 \bibitem{bib:elastic-capacity-planning} Cheng Tian; Ying Wang; Bo Yin.\textsl{An elastic capacity planning method in cloud}. ICCSE, pages 239-244, 2012.
-\end{thebibliography)
-
+\end{thebibliography}
 \end{document}

### 2014-08-11 paper/proposal-ext.tex
extending proposal with design of algorithms in pseudocode

--- 2014-08-07 paper/proposal.tex
@@
 \usepackage{times}
+\usepackage[]{algorithm2e}
 
@@
 Post-processing of this information will be done to distribute total number of running hours per day to align with pricing model.
+
+	Post-processing algorithm to calculate total uptime per hour
+
+	\begin{algorithm}[H]
+ 		\KwData{Timestamp in epoch, Uptime in seconds}
+ 		\KwResult{total uptime in seconds per hour}
+ 		\If{Timestamp indicates a start of new hour e.g. 02:00:00 UTC} {
+ 			set startHour to Timestamp\;
+ 			set endHour to Timestamp + 1h\;
+ 			set totalUptime to 0\;
+ 			set lastUptime to Uptime\;
+ 			return;
+ 		}
+ 		\eIf{Timestamp is between startHour and endHour} {
+ 			\eIf{Uptime is greater than lastUptime} {
+ 				totalUptime = Uptime\;
+ 			}
+ 			{
+ 				totalUptime += Uptime\;
+ 			}
+ 		} (\textsl{Timestamp equals to endHour} \textbf{then} )
+ 		{
+ 			return calculate\_total\_number\_\of\_seconds\_for\_the\_hour()\;
+ 		}
+ 		return\;
+
+ \caption{Post-processing of uptime raw data}
+\end{algorithm}
+
 	\item Pricing scheme - one of: on-demand, light, medium, heavy - a property of reserved machine that pricing is based on.

### 2014-08-11 paper/proposal-ext.tex
Fixes to first algorithm

--- 2014-08-07 paper/proposal.tex
@@
 \usepackage{times}
+\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
 
@@
 Post-processing of this information will be done to distribute total number of running hours per day to align with pricing model.
+
+	\begin{algorithm}[H]
+ 		\KwData{Timestamp in epoch, Uptime in seconds}
+ 		\KwResult{total uptime in seconds per hour}
+ 		\If{Timestamp indicates a start of new hour e.g. 02:00:00 UTC} {
+ 			set startHour to Timestamp\;
+ 			set endHour to Timestamp + 1h\;
+ 			set totalUptime to Uptime\;
+ 			set initialUptime to Uptime\;
+ 			set lastUptime to Uptime\;
+ 			set dived to False\;
+ 			return;
+ 		}
+ 		\eIf{Timestamp is between startHour and endHour} {
+ 			\If{Uptime is smaller than lastUptime} {
+ 				totalUptime += lastUptime\;
+ 				set dived to True\;
+ 			}
+ 		} (\textsl{Timestamp equals endHour} \textbf{then} ) 
+ 		{
+ 			\eIf{dived} {
+ 				return totalUptime - initialUptime\;
+ 				}
+ 			{
+ 			return Uptime - initialUptime\;
+ 			}
+ 		}
+ \caption{Post-processing algorithm to calculate total uptime per hour}
+\end{algorithm}
+
 	\item Pricing scheme - one of: on-demand, light, medium, heavy - a property of reserved machine that pricing is based on.

### 2015-01-25 paper/paper.tex
initial report structure

\documentclass[]{final_report}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[normalem]{ulem}

%%%%%%%%%%%%%%%%%%%%%%
%%% Project details
%%%%%%%%%%%%%%%%%%%%%%
\def\studentname{Kamil Smuga}
\def\projecttitle{An approach for Continuous Capacity Planning in Cloud Environments with Uptime-based Pricing Model}
\def\supervisorname{Liam Murphy}
\def\moderatorname{Christina Thorpe}

\begin{document}

\maketitle
\tableofcontents\pdfbookmark[0]{Table of Contents}{toc}\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% ABSTRACT 
%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}

\textsl{New Infrastructure as a Service solutions are becoming available with a growing number of supported pricing models. More often than not, a hosted Cloud environment is used to design and build an infrastructure for a product. The recent availability of different pricing schemes based on resource utilization and uptime reveals new challenges in already unpredictable capacity planning process. There is a choice between ad-hoc provisioning and upfront payments with reduced hourly rates. Reserved instances charged upfront are categorized into three groups: light, medium and heavy. Which one is better for a given utilization model? When exactly does one pricing scheme becomes more cost effective? Determining which machine type is better for a given utilization model, or at which point the cost effectiveness of a pricing scheme changes, is vital for the companies subscribing to the IaaS. }

\end{abstract}
\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% INTRODUCTION 
%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}

\section{Objectives}

\section{Structure of the Document}

\newpage

%%%%%%%%%%%%%%%%%%%%%%
%%% BACKGROUND 
%%%%%%%%%%%%%%%%%%%%%%

\chapter{Background}

\section{The Problem Domain}

\section{IaaS consumer point of view}

\section{IaaS provider uptime-based pricing schemes}

\section{Literature review}

%%%%%%%%%%%%%%%%%%%%%%
%%% OPTIMALITY 
%%% ALGORITHM 
%%%%%%%%%%%%%%%%%%%%%%

\chapter{Design and Implementation of Optimality Algorithm Based on Uptime and Pricing Scheme}

\section{Design}

\section{Implementation}

%%%%%%%%%%%%%%%%%%%%%%
%%% ALGORITHM TESTING  
%%% ON GOOGLE DATASET 
%%%%%%%%%%%%%%%%%%%%%%

\chapter{Algorithm Testing on Google Cluster Usage Dataset}

%%%%%%%%%%%%%%%%%%%%%%
%%% CAPACITY PLANNING 
%%% STRATEGIES    
%%%%%%%%%%%%%%%%%%%%%%

\chapter{Capacity Planning Strategies}

\section{Peak-to-mean ratio based resource assignment}

\section{Only heavy instances}

%%%%%%%%%%%%%%%%%%%%%%
%%% CONCLUSION 
%%%%%%%%%%%%%%%%%%%%%%

\chapter{Conclusion and Further Work}

%%%%%%%%%%%%%%%%%%%%%%
%%% BIBLIOGRAPHY 
%%%%%%%%%%%%%%%%%%%%%%

\newpage
\begin{thebibliography}{99}
\bibitem{DAWSON:2000} Christian Dawson. \emph{The Essence of Computing Projects -- A Student's Guide}. 192 pages. ISBN: 013021972X. Pearson Education, 2000.
\end{thebibliography}
\label{endpage}

\end{document}

\end{article}

### 2015-01-31 paper/paper.tex
high level sections for algorithm design
cost per hour

@@
 \usepackage[normalem]{ulem}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\usepackage{csvsimple}
 
@@
 
+\section{Motivation}
+
 \section{IaaS consumer point of view}
@@
 
+\section{Taxonomy}
+
+\subsection{Upfront cost}
+
+Investement in compute resources comes with upfront cost.
+In case of physical machine it is hardware or lease cost.
+For virtual resources, IaaS providers offer per hour discounts for upfront charged schemes.
+\par
+Upfront cost is considered to be one time payment and will be represented by \textit{u}.
+
+\subsection{Cost per hour}
+
+Cost per hour represents an aggregated cost of running a single machine divided by number of hours in a calendar year.
+Aggregate cost is relatively easy to calculate for virtualized environments.
+IaaS providers usually charge an X amount per hour.
+It might vary based on total uptime per month or pricing scheme.
+This number will be considered as partial cost per hour and defined as \textit{pcph}.
+\par
+Cost of running of physical hardware is less straightforward to calculate as it includes: power consumption, operations, hardware replacements cost, data center related costs and many others that may vary based on individual cases.
+Although more troublesome to calculate, cost per hour per host for physical infrastructure is possible to calculate.
+\par
+Cost per hour will be defined as \textit{cph}.
+Considering the fact that upfront cost is more substantial for light usage than 24/7, cost per hour will be defined as a function of hours/day usage defined as \textit{h} and calculated from equation:
+\begin{equation}
+cph(h) = \frac{u + \sum_{i=1}^{366} pcph \times h}{365 \times h}
+\end{equation}
 \section{Design}
-
 \section{Implementation}
@@
 \chapter{Algorithm Testing on Google Cluster Usage Dataset}
+
+\section{Input data assumptions}
+
+Based on data published for AWS~\cite{AWS:2014}
+
+\csvautotabular{out}
+
+\includegraphics[width=\linewidth]{cph}
+\includegraphics[width=\linewidth]{cph_zoom}
+
+\section{Calculate cost per hour for each pricing scheme: light, medium and heavy}
+
+\section{Graph price per hour on y and number of hours/day on x to find cut points}
+
+\section{Calculate optimality factor based on distance from cut points}
 
@@
 \begin{thebibliography}{99}
-\bibitem{DAWSON:2000} Christian Dawson. \emph{The Essence of Computing Projects -- A Student's Guide}. 192 pages. ISBN: 013021972X. Pearson Education, 2000.
+\bibitem{AWS:2014} http://forecastcloudy.net/2012/04/02/amazon-web-services-aws-ec2-pricing-data/ https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-light-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-medium-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-heavy-linux.json
 \end{thebibliography}

### 2015-01-31 paper/calculate_cost_per_hour.py

@@ -1,0 +1,26 @@
+#!/usr/bin/env python
+
+lights = []
+mediums = []
+heavies = []
+
+# values for size: cc2.8xlarge region: us-east
+# 1 year reservation
+light = { "upfront" : "1762", "perhour" : "0.904" }
+medium = { "upfront" : "4146", "perhour" : "0.54" }
+heavy = { "upfront" : "5000", "perhour" : "0.361" }
+
+def calc_cost_per_hour(upfront, cost_per_hour, hours_per_day):
+    upfront = int(upfront)
+    cost_per_hour = float(cost_per_hour)
+    cph = (upfront + (365 * cost_per_hour * hours_per_day)) / (365 * hours_per_day)
+    return cph
+
+lights = [calc_cost_per_hour(light["upfront"], light["perhour"], x) for x in range(1,25)] 
+mediums = [calc_cost_per_hour(medium["upfront"], medium["perhour"], x) for x in range(1,25)] 
+heavies = [calc_cost_per_hour(heavy["upfront"], heavy["perhour"], x) for x in range(1,25)] 
+
+print "light,medium,heavy"
+for i in range(24):
+    print "%s,%s,%s" % (lights[i], mediums[i], heavies[i])
+

### 2015-02-01 calculate_cost_per_hour.py
calculate cost per hour for more examples

@@ -1,0 +1,64 @@
+#!/usr/bin/env python
+
+import matplotlib.pyplot as plot
+
+lights = []
+mediums = []
+heavies = []
+entities = []
+
+class Entity:
+    def __init__(self, name, light, medium, heavy):
+        self.name = name
+        self.light = light
+        self.medium = medium
+        self.heavy = heavy
+
+    def calc_cost_per_hour(self, mode, hours_per_day):
+        upfront = int(mode["upfront"])
+        cost_per_hour = float(mode["perhour"])
+        cph = (upfront + (365 * cost_per_hour * hours_per_day)) / (365 * hours_per_day)
+        return cph   
+
+# values for size: cc2.8xlarge region: us-east
+# 1 year reservation
+c_name = 'cc2.8xlarge'
+c_light = { "upfront" : "1762", "perhour" : "0.904" }
+c_medium = { "upfront" : "4146", "perhour" : "0.54" }
+c_heavy = { "upfront" : "5000", "perhour" : "0.361" }
+
+# values for size: m2.4xlarge region: us-east
+# 1 year reservation
+m_name = 'm2.4xlarge'
+m_light = { "upfront" : "1088", "perhour" : "0.676" }
+m_medium = { "upfront" : "2604", "perhour" : "0.52" }
+m_heavy = { "upfront" : "3156", "perhour" : "0.272" }
+
+# values for size: hs1.8xlarge region: us-east
+# 1 year reservation
+h_name = 'hs1.8xlarge'
+h_light = { "upfront" : "3968", "perhour" : "2.24" }
+h_medium = { "upfront" : "9200", "perhour" : "1.38" }
+h_heavy = { "upfront" : "11213", "perhour" : "0.92" }
+
+c_entity = Entity(c_name, c_light, c_medium, c_heavy)
+m_entity = Entity(m_name, m_light, m_medium, m_heavy)
+h_entity = Entity(h_name, h_light, h_medium, h_heavy)
+
+entities.extend([c_entity, m_entity, h_entity])
+
+for e in entities:
+    lights = [ e.calc_cost_per_hour(e.light, x) for x in range(1,25) ] 
+    mediums = [ e.calc_cost_per_hour(e.medium, x) for x in range(1,25) ]
+    heavies = [ e.calc_cost_per_hour(e.heavy, x) for x in range(1,25) ]
+
+    t = range(24)
+    plot.plot(t, lights, t, mediums, t, heavies)
[…]

### 2015-03-14 paper/paper.tex
paper high level design improvements

@@
 
+\emph{ /* To be revised. */} \par
 \textsl{New Infrastructure as a Service solutions are becoming available with a growing number of supported pricing models.
@@
 
+\emph{ /* Type of the project: The 'Big Idea' / proof by construction. */}
+
 \section{Objectives}
+This work aims to design an algorithm that answer IaaS consumer's price optimality questions and help to make adjustments based on uptime based charging per machine.
+
+%\section{Thesis statement}
+%\emph{ /* Based on~\cite{proofByConstruction} I have to provide a statement that will be proven by construction. Is it a valid approach? */} \par
+
+%To have a grasp on 'how optimal my infrastructure provisioning is' one should gather VM usage data, split and calculate optimality KPIs
+
+%To answer price optimality questions asked by IaaS customer, one should split and model usage data and run optimality KPIs based on
 
 \section{Structure of the Document}
+
+\textbf{Background}.
+This section will outline the background information related to IaaS consumer challenges related to understanding the whole picture for price optimality of rented infrastructure.
+\par
+\textbf{Design}. This section explains design and environment conditions for the algorithm to be useful. \par
+\textbf{Implementation}. This sections explains algorithm implementation details and data analytics software that was used - Apache Spark~\cite{spark}. \par
+\textbf{Evaluation}. The proposed algorithmic approach is applied on Google Trace data~\cite{googleTrace}. \par
+\textbf{Conclusion and Further Work}. The completed data analysis is discussed. Further work related to algorithm improvements, automation and lessons learned are presented.
 
@@
 \section{The Problem Domain}
+\emph{/* Description of IaaS world - providers and consumers. */}
+\subsection{IaaS consumer point of view}
+\subsection{IaaS provider uptime-based pricing schemes}
 
 \section{Motivation}
-
-\section{IaaS consumer point of view}
-
-\section{IaaS provider uptime-based pricing schemes}
+\emph{ /* Description of the problem - cloud based company runs a mix of software services (real-time streaming, batch, web and database servers) on common hardware.
+How do they know whether provisioned VMs run in the most optimal configuration?
+*/ }
 
 \section{Literature review}
+\emph{/* Haven't found anything that would tackle this specific problem so far.
+Might mention the most research is done from IaaS provider perspective?
+This includes resource scheduling.
+My work can lead to cost-aware scheduler research */}
+
+%good number of journal papers, a swathe of conference papers, a few recent workshop
+%papers and not too many web sites
 
 %%%%%%%%%%%%%%%%%%%%%%
-%%% OPTIMALITY 
-%%% ALGORITHM 
+%%% DESIGN 
 %%%%%%%%%%%%%%%%%%%%%%
 
-\chapter{Design and Implementation of Optimality Algorithm Based on Uptime and Pricing Scheme}
+\chapter{Design}
 
@@
 
-Investement in compute resources comes with upfront cost.
+Investment in compute resources comes with upfront cost.
 In case of physical machine it is hardware or lease cost.
@@
 \end{equation}
-\section{Design}
-\section{Implementation}
+
+\subsection{Intersection points between pricing schemes}
+\emph{/* How to find points when one pricing scheme becomes */}
+
+\section{Algorithm}
+
+\emph{/* High level design of the algorithm in pseudo-code? */}
+
+\subsection{Calculate cost per hour for a given pricing scheme}
+\subsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+\subsection{Collect VM usage data}
+\subsection{Group usage metrics by machine}
+\subsection{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
+\subsection{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}
 
 %%%%%%%%%%%%%%%%%%%%%%
+%%% IMPLEMENTATION 
+%%%%%%%%%%%%%%%%%%%%%%
+
+\chapter{Implementation}
+
+\emph{/* Implementation quirks related to make it work. Mostly map/reduce jobs written for Apache Spark. */}
+
+%%%%%%%%%%%%%%%%%%%%%%  
+%%% EVALUATION!!!
 %%% ALGORITHM TESTING  
@@
 
-\chapter{Algorithm Testing on Google Cluster Usage Dataset}
+\chapter{Evaluation}
 
-\section{Input data assumptions}
+\emph{/* Evaluation will go through each of Design stages and will perform necessary calculations. Some examples attached below. */}
 
-Based on data published for AWS~\cite{AWS:2014}
-
+\section{Calculate cost per hour for a given pricing scheme}
+Results for ccs2.xlarge based on data published for AWS~\cite{AWS:2014} \par
 \csvautotabular{out}
-
+\section{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
 \includegraphics[width=\linewidth]{cph}
-\includegraphics[width=\linewidth]{cph_zoom}
-
-\section{Calculate cost per hour for each pricing scheme: light, medium and heavy}
-
-\section{Graph price per hour on y and number of hours/day on x to find cut points}
-
-\section{Calculate optimality factor based on distance from cut points}
+\section{Collect VM usage data}
+\section{Group usage metrics by machine}
+\section{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
+\section{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}
 
 %%%%%%%%%%%%%%%%%%%%%%
+%%% OPTIONAL!!!
 %%% CAPACITY PLANNING 
@@
 
-\chapter{Capacity Planning Strategies}
+%\chapter{Capacity Planning Strategies}
 
-\section{Peak-to-mean ratio based resource assignment}
+%\section{Peak-to-mean ratio based resource assignment}
 
-\section{Only heavy instances}
+%\section{Only heavy instances}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-%%% BIBLIOGRAPHY 
+%%% REFERENCES 
 %%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{thebibliography}{99}
+\bibitem{spark} https://spark.apache.org
+\bibitem{googleTrace} https://code.google.com/p/googleclusterdata
 \bibitem{AWS:2014} http://forecastcloudy.net/2012/04/02/amazon-web-services-aws-ec2-pricing-data/ https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-light-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-medium-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-heavy-linux.json

### 2015-03-15 paper/paper.tex
small updates

@@
 
-Cost per hour represents an aggregated cost of running a single machine divided by number of hours in a calendar year.
-Aggregate cost is relatively easy to calculate for virtualized environments.
+Cost per hour represents a total cost of running a single machine divided by number of hours in calendar year.
+Aggregated cost is relatively easy to calculate in rented, virtualized environments.
 IaaS providers usually charge an X amount per hour.
 It might vary based on total uptime per month or pricing scheme.
-This number will be considered as partial cost per hour and defined as \textit{pcph}.
+This number will be considered as a partial cost per hour and defined as \textit{pcph}.
 \par
@@
 \par
-Cost per hour will be defined as \textit{cph}.
-Considering the fact that upfront cost is more substantial for light usage than 24/7, cost per hour will be defined as a function of hours/day usage defined as \textit{h} and calculated from equation:
+Cost per hour will be defined as \textit{cph}. It includes upfront cost and IaaS provider per hour charges. It is defined as a function of hours of utilization per day.
+
 \begin{equation}

### 2015-03-17 paper/paper.tex
small refactorings
updates to Design: Intersection points

@@
 \usepackage{csvsimple}
+\usepackage{float}
 
@@
 \def\studentname{Kamil Smuga}
-\def\projecttitle{An approach for Continuous Capacity Planning in Cloud Environments with Uptime-based Pricing Model}
-\def\supervisorname{Liam Murphy}
+\def\projecttitle{An approach for Continuous Capacity Planning in Cloud Environments with an Uptime-based Pricing Model}
+\def\supervisorname{Prof. Liam Murphy}
 \def\moderatorname{Christina Thorpe}
@@
 \tableofcontents\pdfbookmark[0]{Table of Contents}{toc}\newpage
+\listoffigures\newpage
+\par{\textbf{List of tables}}
 
+%%%%
+%The most important parts of your thesis are: abstract, introduction, conclusion and
+%references. These are what get read first and make that vital initial impression.
+%%%
+%%%
+%One of first things your examiners will look at is your literature review. If they see a
+%good number of journal papers, a swathe of conference papers, a few recent workshop
+%papers and not too many web sites, they’ll already be impressed.
+%%%
 %%%%%%%%%%%%%%%%%%%%%%
@@
 
+\section{Description of the problem}
+
+\section{Motivation for solving the problem}
+
 \section{Objectives}
@@
 
-%\section{Thesis statement}
-%\emph{ /* Based on~\cite{proofByConstruction} I have to provide a statement that will be proven by construction. Is it a valid approach? */} \par
+\section{Related work}
 
-%To have a grasp on 'how optimal my infrastructure provisioning is' one should gather VM usage data, split and calculate optimality KPIs
+\section{Challenges}
 
-%To answer price optimality questions asked by IaaS customer, one should split and model usage data and run optimality KPIs based on
+\section{Description of methodology}
+Modelling and experimentation
 
-\section{Structure of the Document}
-
+\section{Structure of the Thesis}
 \textbf{Background}.
@@
 Cost per hour represents a total cost of running a single machine divided by number of hours in calendar year.
-Aggregated cost is relatively easy to calculate in rented, virtualized environments.
+Aggregated cost is relatively easy to calculate in rented, virtualised environments.
 IaaS providers usually charge an X amount per hour.
@@
 \subsection{Intersection points between pricing schemes}
-\emph{/* How to find points when one pricing scheme becomes */}
+
+Cost per hour is calculated for each of available pricing schemes.
+Single result represents a matrix of hours per day and \textit{cph}.
+Graphical representation of results will produce a graph similar to Figure~\ref{fig:cc2_8xlarge}.
+Pricing trends are visible.
+Price drops exponentially for around first 5 hours of usage and almost stabilizes afterwards.
+Light is the cheapest option for around 15 hours per day usage.
+This is better visible on Figure~\ref{fig:cph_cc2_8xlarge_zoom}.
+After 15 hours, it is better to invest in Heavy utilization scheme.
+This is a point that will be called an intersection point between 2 pricing schemes and referred as \textit{ip(scheme1, scheme2)}.
+In this example, we can calculate that for cc2.8xlarge instance \textit{ip(light, heavy) = 16}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{figures/cc2_8xlarge}
+	\caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance for 1 year~\cite{AWS:2014}}
+	\label{fig:cc2_8xlarge}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{figures/cph_cc2_8xlarge_zoom}
+	\caption{Zoomed \textit{cph} results for Amazon AWS cc2.8xlarge reserved instance for 1 year~\cite{AWS:2014}}
+	\label{fig:cph_cc2_8xlarge_zoom}
+\end{figure}
 
@@
 
-\section{Calculate cost per hour for a given pricing scheme}
+\section{Methodology and setup description}
+
+\section{Algorithm walkthrough}
+\subsection{Calculate cost per hour for a given pricing scheme}
 Results for ccs2.xlarge based on data published for AWS~\cite{AWS:2014} \par
 \csvautotabular{out}
-\section{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
-\includegraphics[width=\linewidth]{cph}
-\section{Collect VM usage data}
-\section{Group usage metrics by machine}
-\section{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
-\section{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}
+\subsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+%\includegraphics[width=\linewidth]{cph}
+\subsection{Collect VM usage data}
+\subsection{Group usage metrics by machine}
+\subsection{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
+\subsection{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}
+
+\section{Summary}
 
@@
 
-\chapter{Conclusion and Further Work}
+\chapter{Conclusion and Future Work}
 

### 2015-03-21 paper/paper.tex
couple of algorithm designs

@@
 \usepackage{float}
+\usepackage[]{algorithm2e}
 
@@
 Single result represents a matrix of hours per day and \textit{cph}.
-Graphical representation of results will produce a graph similar to Figure~\ref{fig:cc2_8xlarge}.
-Pricing trends are visible.
-Price drops exponentially for around first 5 hours of usage and almost stabilizes afterwards.
+Graphical representation of results produces a graph similar to Figure~\ref{fig:cc2_8xlarge}.
+\par
+This representation allows to notice pricing trends.
+In the example below, price drops exponentially for around first 5 hours of usage and nearly stabilizes afterwards.
 Light is the cheapest option for around 15 hours per day usage.
@@
 After 15 hours, it is better to invest in Heavy utilization scheme.
-This is a point that will be called an intersection point between 2 pricing schemes and referred as \textit{ip(scheme1, scheme2)}.
-In this example, we can calculate that for cc2.8xlarge instance \textit{ip(light, heavy) = 16}
+This is the point that will be called \textit{an intersection point} between 2 pricing schemes and noted as \textit{ip(scheme1, scheme2)}.
+\par
+For the example below, cc2.8xlarge instance's intersection point between Light and Heavy schemes is \textit{ip(light, heavy) = 16}.
 
@@
 \subsection{Calculate cost per hour for a given pricing scheme}
+
+\begin{algorithm}[H]
+ \KwData{Upfront cost, Per hour charge, Utilisation [hours per day];}
+ \KwResult{Total cost per hour for a given pricing scheme defined as \textit{cph};}
+ read upfront\;
+ read perhour\;
+ read hours\;	
+ return (upfront + (365 * perhour * hours)) / (365 * hours)
+ \caption{Calculate cost per hour for a given pricing scheme}
+\end{algorithm}
+
 \subsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+
+\begin{algorithm}[H]
+ \KwData{Array of cph for 2 pricing schemes;}
+ \KwResult{Intersection point when one pricing scheme becomes cheaper than the other one;}
+ read array1\;
+ read array2\;
+ \ForAll{cph in array} {
+ 	\If{$cph2 <= $cph1} {
+ 		return cph
+ 	}
+ }
+\caption{Calculate intersection point between two pricing schemes}
+\end{algorithm}
+
 \subsection{Collect VM usage data}
+
+        1,machine ID,INTEGER,YES
+
+        2,start time,INTEGER,YES
+
+        3,end time,INTEGER,YES
+
+        4,duration INTEGER,YES
+
+        5,sampled CPU usage,FLOAT,NO
+
+        6,assigned memory usage,FLOAT,NO
+
+        7,maximum memory usage,FLOAT,NO
+
 \subsection{Group usage metrics by machine}
+
+\begin{algorithm}[H]
+ \KwData{VM usage data;}
+ \KwResult{Metrics containing calculated duration and day;}
+ read vmMetrics\;
+ \ForAll{line in vmMetrics} {
+        calculateDuration();
+        calculateDay();
+ }
+\caption{VM metrics transformation}
+
 \subsection{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
+
 \subsection{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}

### 2015-04-05 calculate_cost_per_hour.py
calculate number of tasks per machine per day

@@ -266,7 +266,7 @@ def aggregated_uptime(line):
 
 """
 RRD Spark operations for above ^^
-"""
+
 
 def mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
@@ -276,7 +276,7 @@ def mapping(line):
     return (machine_id, (start, end))
 
 for x in range(0, 30):
-    distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-third-day-" + str(day) + "/part*", use_unicode=False)
+    distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-third-day-" + str(x) + "/part*", use_unicode=False)
 
     min_start = distFile.map(lambda(x): (x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[0], 
                 float(x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[2]))).reduceByKey(lambda a,b: a if a<b else b)
@@ -297,8 +297,29 @@ for x in range(0, 30):
             float(x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[7].strip()))).reduceByKey(lambda a,b: a + b)
 
     destFile = min_start.join(max_end).join(cpu).join(mem).join(total_mem)
-    destFile.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(day))
+    destFile.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x))
     day += 1
+"""
+
+"""
+
+Calculate number of tasks per machine per day
+
+"""
+
+def mapping(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
+    machine_id = splits[0].strip()
+    counter = 1
+    return (machine_id, counter)
+
+for x in range(0, 30):
+    distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-third-day-" + str(x) + "/part*", use_unicode=False)
+
+    task_counter = distFile.map(mapping).reduceByKey(add)
+
+    task_counter.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-forth-tasks-day-" + str(x))
+
 
 
 """

### 2015-04-05 calculate_cost_per_hour.py
add add operator

@@ -4,6 +4,7 @@ import matplotlib.pyplot as plot
 import csv
 from ast import literal_eval
 from pyspark import SparkContext, SparkConf, StorageLevel
+from operator import add
 
 """
 --------------------------------------------------------

### 2015-04-05 calculate_cost_per_hour.py
join capacity and number of tasks info

@@ -306,7 +306,6 @@ for x in range(0, 30):
 
 Calculate number of tasks per machine per day
 
-"""
 
 def mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
@@ -321,7 +320,48 @@ for x in range(0, 30):
 
     task_counter.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-forth-tasks-day-" + str(x))
 
+"""
 
+""" 
+JOIN WITH CPU AND CAPACITY INFO
+
+Schema after transformation (splitted per day)
+        1,machine ID,INTEGER,YES
+        2,uptime (how many hours machine is on),INTEGER,YES
+        3,uptime based on summarized task time,INTEGER,YES
+        4,total CPU usage,INTEGER,YES
+        5,CPU capacity, INTEGER, YES
+        6,total assigned memory usage,FLOAT,NO
+        7,total maximum memory usage,FLOAT,NO
+        8,MEMORY capacity,FLOAT,YES
+        9,NUMBER of tasks per machine per day
+
+"""
+
+def mapping(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
+    machine_id = splits[0].strip()
+    uptime = float(splits[1].strip())
+    uptime_total = float(splits[2].strip())
+    total_cpu = float(splits[3].strip())
+    mem = float(splits[4].strip())
+    total_mem = float(splits[5].strip())
+    return (machine_id, (uptime, uptime_total, total_cpu, mem, total_mem))
+
+for x in range(0, 30):
+    distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
+    basic_forth = distFile.map(mapping)
+
+    distFileTasks = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-tasks-day" + str(x) + "/part*", use_unicode=False)
+    tasks_forth = distFileTasks.map(lambda x: (x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[0], x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[1])) 
+
+    capacity = sc.textFile("/Users/ksmuga/workspace/data/clusterdata-2011-2/machine_events/*")
[…]

### 2015-04-05 calculate_cost_per_hour.py
remove unicode

@@ -355,7 +355,7 @@ for x in range(0, 30):
     distFileTasks = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-tasks-day" + str(x) + "/part*", use_unicode=False)
     tasks_forth = distFileTasks.map(lambda x: (x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[0], x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[1])) 
 
-    capacity = sc.textFile("/Users/ksmuga/workspace/data/clusterdata-2011-2/machine_events/*")
+    capacity = sc.textFile("/Users/ksmuga/workspace/data/clusterdata-2011-2/machine_events/*", use_unicode=False)
     capacity_map = capacity.map(lambda x: (x.split(",")[1], (x.split(",")[4], x.split(",")[5])))
 
     joined_file = basic_forth.join(capacity_map).join(tasks_forth)

### 2015-04-05 calculate_cost_per_hour.py
calc only uptime

@@ -203,8 +203,8 @@ Calculate uptime and real workload (based on uptime)
 Schema after transformation:
 
         1,machine ID,INTEGER,YES
-        2,uptime (how many hours machine is on),INTEGER,YES
-        3,uptime based on summarized task time,INTEGER,YES
+        2,min_start (minimum start hour),INTEGER,YES
+        3,max_end (maximum start hour),INTEGER,YES
         4,total CPU usage,INTEGER,YES
         5,total assigned memory usage,FLOAT,NO
         6,total maximum memory usage,FLOAT,NO
@@ -336,7 +336,7 @@ Schema after transformation (splitted per day)
         8,MEMORY capacity,FLOAT,YES
         9,NUMBER of tasks per machine per day
 
-"""
+
 
 def mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
@@ -362,6 +362,25 @@ for x in range(0, 30):
 
     joined_file.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x))
     day += 1
+"""
+"""
+
+GET ONLY UPTIME
+
+"""
+
+def only_uptime_mapping(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
+    machine_id = splits[0].strip()
+    uptime = float(splits[2].strip()) - float(splits[1].strip())
+    return (machine_id, uptime)
+
+for x in range(0, 30):
+    distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
+    only_uptime = distFile.map(only_uptime_mapping)
+    distinct_uptime = only_uptime.distinct()
+    distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
+
 
 """
 --------------------------------------------------------

### 2015-04-06 calculate_cost_per_hour.py
small uodates

@@ -337,6 +337,7 @@ Schema after transformation (splitted per day)
         9,NUMBER of tasks per machine per day
 
 
+"""
 
 def mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
@@ -352,22 +353,20 @@ for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
     basic_forth = distFile.map(mapping)
 
-    distFileTasks = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-tasks-day" + str(x) + "/part*", use_unicode=False)
+    distFileTasks = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-tasks-day-" + str(x) + "/part*", use_unicode=False)
     tasks_forth = distFileTasks.map(lambda x: (x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[0], x.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")[1])) 
 
     capacity = sc.textFile("/Users/ksmuga/workspace/data/clusterdata-2011-2/machine_events/*", use_unicode=False)
     capacity_map = capacity.map(lambda x: (x.split(",")[1], (x.split(",")[4], x.split(",")[5])))
 
     joined_file = basic_forth.join(capacity_map).join(tasks_forth)
-
-    joined_file.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x))
+    joined_file_distinct = joined_file.distinct()
+    joined_file_distinct.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x))
     day += 1
-"""
 """
 
 GET ONLY UPTIME
 
-"""
 
 def only_uptime_mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
@@ -381,6 +380,7 @@ for x in range(0, 30):
     distinct_uptime = only_uptime.distinct()
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
+"""
 
 """
 --------------------------------------------------------

### 2015-04-08 paper/paper.tex
updates to paper

@@
 
-\subsection{Collect VM usage data}
+\subsection{Metrics collection}
 
@@
 
-\subsection{Group usage metrics by machine}
+        8,CPU capacity,FLOAT,YES
 
-\begin{algorithm}[H]
- \KwData{VM usage data;}
- \KwResult{Metrics containing calculated duration and day;}
- read vmMetrics\;
- \ForAll{line in vmMetrics} {
-        calculateDuration();
-        calculateDay();
- }
-\caption{VM metrics transformation}
+        9,Memory capacity,FLOAT,YES
 
-\subsection{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
+\subsection{Daily metrics aggregation}
+
+\subsubsection{Calculate uptime per machine }
+Input: Job start and end time logged for each machine in a form of \textless K, List\textless V\textgreater\textgreater
+Output: Uptime based on span between min start time and max end time \textless K,V\textgreater
+
+\subsubsection{Aggregate CPU and memory usage}
+Input: Sampled CPU and memory usage logged per job in a form of \textless K, List\textless V\textgreater\textgreater
+Output: Aggregated resource usage \textless K,V\textgreater
+
+\subsubsection{Calculate number of jobs per day}
+Input: Job id per machine \textless K, List\textless V\textgreater\textgreater
+Output: Job count per machine \textless K,V\textgreater
+
+\subsubsection{Calculate cluster averages for aggregations}
+Input:
 

### 2015-04-11 paper/paper.tex
add skeleton of design to the paper

--- 2015-03-21 paper/paper.tex
@@
 \usepackage{amsmath}
-\usepackage{graphicx}
 \usepackage{csvsimple}
@@
 \usepackage[]{algorithm2e}
+\usepackage{titlesec}
+\setcounter{secnumdepth}{4}
+\titleformat{\paragraph}
+{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
+\titlespacing*{\paragraph}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
 
@@
 
-\emph{/* High level design of the algorithm in pseudo-code? */}
+\subsection{Data gathering, transformation and analytics}
 
-\subsection{Calculate cost per hour for a given pricing scheme}
+\subsubsection{Calculate cost per hour for a given pricing scheme}
 
 \begin{algorithm}[H]
- \KwData{Upfront cost, Per hour charge, Utilisation [hours per day];}
+ \KwData{Upfront cost, Per hour charge, Utilization [hours per day];}
  \KwResult{Total cost per hour for a given pricing scheme defined as \textit{cph};}
@@
 
-\subsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+\subsubsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
 
@@
 
-\subsection{Collect VM usage data}
+\subsubsection{Metrics collection}
 
@@
 
-\subsection{Group usage metrics by machine}
+        8,CPU capacity,FLOAT,YES
 
-\begin{algorithm}[H]
- \KwData{VM usage data;}
- \KwResult{Metrics containing calculated duration and day;}
- read vmMetrics\;
- \ForAll{line in vmMetrics} {
-        calculateDuration();
-        calculateDay();
- }
-\caption{VM metrics transformation}
+        9,Memory capacity,FLOAT,YES
 
-\subsection{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
+\subsubsection{Daily metrics aggregation}
 
-\subsection{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}
+\paragraph{Calculate uptime per machine}
+Input: Job start and end time logged for each machine in a form of \textless K, List\textless V\textgreater\textgreater
+
+Output: Uptime based on span between min start time and max end time \textless K,V\textgreater
+
+\paragraph{Aggregate CPU and memory usage}
+Input: Sampled CPU and memory usage logged per job in a form of \textless K, List\textless V\textgreater\textgreater
+
+Output: Aggregated resource usage \textless K,V\textgreater
+
+\paragraph{Calculate number of jobs per day}
+Input: Job id per machine \textless K, List\textless V\textgreater\textgreater
+
+Output: Job count per machine \textless K,V\textgreater
+
+\paragraph{Calculate cluster averages for aggregations}
+Input: Aggregated CPU, memory and number of tasks per machine per day \textless K, List\textless V\textgreater\textgreater
+
+Output: Averages for CPU, memory and number of tasks \textless K, List\textless V\textgreater\textgreater
+
+\paragraph{Calculate distance from optimality points based on uptime}
+Input: Machine uptime per day
+
+Output: Distance from optimality points. Might be positive or negative.
+
+\subsection{Recognize inefficiencies and suggest changes}
+
+\subsubsection{Threshold based report}
+
+Generate report of machines based on distance from optimality points. Use customizable threshold. 
+Results from the report will be used to calculate suggestions.
+
+\subsubsection{Visual recognition}
+
+Visual recognition as a helper method to define thresholds, especially for further incremental optimizations. 
+Useful as static thresholds can miss a whole set of machines that are just below the threshold.
+
+\begin{figure}[H]
+        \includegraphics[width=\linewidth]{figures/cubism}
+        \caption{Visualization of distance from optimality points. Green - positive, Blue - negative values.}
+        \label{fig:cubism}
+\end{figure}
+
+\subsubsection{Squeeze-them-in suggestion strategy}
+
+Input: results of the Threshold based report.
+
+Computation: Calculate spare capacity based on average CPU and memory usage.
+Iterate through a pool of available machines (not included in input report) and advertise available capacity only if CPU and memory usage is lower than average.
+Aggregate spare capacity results.
+
+Output: List of machines that can be shutdown based on computed spare-capacity.
+
+\subsubsection{New capacity configuration suggestion}
+
+Input: results of the Threshold based report.
+
+Computation: Compute a minimal number of VMs for each configuration based on requested capacity
+
+Output: A number and configuration of the cheapest computed configuration.
 
@@
 \emph{/* Evaluation will go through each of Design stages and will perform necessary calculations. Some examples attached below. */}
+
+\subsection{Recognize inefficiencies and suggest changes}
+
+\subsubsection{Threshold based report}
+
+Generate report of machines based on distance from optimality points. Use customizable threshold. 
+Results from the report will be used to calculate suggestions.
+
+\subsubsection{Visual recognition}
 

### 2015-04-11 paper/paper.tex
paper updates

@@
 \usepackage[]{algorithm2e}
-\usepackage{titlesec}
-\setcounter{secnumdepth}{4}
-\titleformat{\paragraph}
-{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
-\titlespacing*{\paragraph}
-{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+\usepackage{url}
+
+\newcommand{\myparagraph}[1]{\paragraph{#1}\mbox{}\\}
 
@@
 
-\chapter{Design}
+\chapter{Design and Implementation}
 
@@
 	\includegraphics[width=\linewidth]{figures/cc2_8xlarge}
-	\caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance for 1 year~\cite{AWS:2014}}
+	\caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:2014}}
 	\label{fig:cc2_8xlarge}
@@
 	\includegraphics[width=\linewidth]{figures/cph_cc2_8xlarge_zoom}
-	\caption{Zoomed \textit{cph} results for Amazon AWS cc2.8xlarge reserved instance for 1 year~\cite{AWS:2014}}
+	\caption{Zoomed \textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:2014}}
 	\label{fig:cph_cc2_8xlarge_zoom}
@@
 
-\subsection{Data gathering, transformation and analytics}
+\subsection{Prerequisites}
+
+In order to make data-driven analysis and suggestions for optimization we need data.
+Below list of metrics was compiled based on algorithm needs to answer critical questions regarding state of infrastructure deployment.
+The list is suitable to guarantee performance SLOs only for compute nodes.
+Machines with significant I/O usage profile are not covered in this paper.
+Metrics can be gathered periodically, per process or per job.
+There are open source tools available to gather this data, e.g.
+collectd\footnote{\url{https://collectd.org}}, munin\footnote{\url{http://munin-monitoring.org}}, Logstash\footnote{\url{http://logstash.net}}.
+
+\myparagraph{Machine Id}
+A unique machine ID. Will be used as a key to compute further analytics.
+
+\myparagraph{Start and End Time}
+Timestamps to indicate start and end of measurement period.
+
+\myparagraph{CPU usage}
+Sampled or averaged CPU usage during measurement period.
+
+\myparagraph{RAM usage}
+Canonical memory usage measurement.
+
+Metrics should be aggregated in \textless K, List\textless V\textgreater\textgreater format where K represents Machine Id and Vs are measurements.
+
+\subsection{Metrics data transformations and analytics}
+
+Metrics gathered in Prerequisites section have limited knowledge about an environment.
+Collection per job or per process produces one log line and is aware only about measurement period.
+Having such structured data would not let to answer machine wide neither cluster wide questions.
+The data has to be post-processed to allow calculation of below listed analytics.
+
+\myparagraph{Uptime per machine}
+
+\begin{algorithm}[H]
+ \KwData{Task start and end time}
+ \KwResult{Number of uptime hours per machine per day}
+\end{algorithm}
+
+Input: Job start and end time logged for each machine in a form of \textless K, List\textless V\textgreater\textgreater
+
+Output: Uptime based on span between min start time and max end time \textless K,V\textgreater
+
+\myparagraph{Aggregate CPU and memory usage}
+Input: Sampled CPU and memory usage logged per job in a form of \textless K, List\textless V\textgreater\textgreater
+
+Output: Aggregated resource usage \textless K,V\textgreater
+
+\myparagraph{Calculate number of jobs per day}
+Input: Job id per machine \textless K, List\textless V\textgreater\textgreater
+
+Output: Job count per machine \textless K,V\textgreater
+
+\myparagraph{Calculate cluster averages for aggregations}
+Input: Aggregated CPU, memory and number of tasks per machine per day \textless K, List\textless V\textgreater\textgreater
+
+Output: Averages for CPU, memory and number of tasks \textless K, List\textless V\textgreater\textgreater
+
+\myparagraph{Calculate distance from optimality points based on uptime}
+Input: Machine uptime per day
+
+Output: Distance from optimality points. Might be positive or negative.
 
@@
 \end{algorithm}
-
-\subsubsection{Metrics collection}
-
-        1,machine ID,INTEGER,YES
-
-        2,start time,INTEGER,YES
-
-        3,end time,INTEGER,YES
-
-        4,duration INTEGER,YES
-
-        5,sampled CPU usage,FLOAT,NO
-
-        6,assigned memory usage,FLOAT,NO
-
-        7,maximum memory usage,FLOAT,NO
-
-        8,CPU capacity,FLOAT,YES
-
-        9,Memory capacity,FLOAT,YES
-
-\subsubsection{Daily metrics aggregation}
-
-\paragraph{Calculate uptime per machine}
-Input: Job start and end time logged for each machine in a form of \textless K, List\textless V\textgreater\textgreater
-
-Output: Uptime based on span between min start time and max end time \textless K,V\textgreater
-
-\paragraph{Aggregate CPU and memory usage}
-Input: Sampled CPU and memory usage logged per job in a form of \textless K, List\textless V\textgreater\textgreater
-
-Output: Aggregated resource usage \textless K,V\textgreater
-
-\paragraph{Calculate number of jobs per day}
-Input: Job id per machine \textless K, List\textless V\textgreater\textgreater
-
-Output: Job count per machine \textless K,V\textgreater
-
-\paragraph{Calculate cluster averages for aggregations}
-Input: Aggregated CPU, memory and number of tasks per machine per day \textless K, List\textless V\textgreater\textgreater
-
-Output: Averages for CPU, memory and number of tasks \textless K, List\textless V\textgreater\textgreater
-
-\paragraph{Calculate distance from optimality points based on uptime}
-Input: Machine uptime per day
-
-Output: Distance from optimality points. Might be positive or negative.
 

### 2015-04-12 paper/paper.tex
paper updates

@@
 \usepackage[]{algorithm2e}
+\usepackage[noend]{algpseudocode}
 \usepackage{url}
+\usepackage{listings}             % Include the listings-package
 
+\RestyleAlgo{boxruled}
+\lstset{language=Python}
 \newcommand{\myparagraph}[1]{\paragraph{#1}\mbox{}\\}
@@
 
-Metrics gathered in Prerequisites section have limited knowledge about an environment.
-Collection per job or per process produces one log line and is aware only about measurement period.
+Metrics gathered in Prerequisites section have limited knowledge about environment.
+Collection per job or per process produces one log line and has scope of a measurement period.
 Having such structured data would not let to answer machine wide neither cluster wide questions.
-The data has to be post-processed to allow calculation of below listed analytics.
+The data has to be post-processed to allow calculation of below listed aggregations and analytics.
+
+\myparagraph{Daily usage}
+
+The data will be analyzed in terms of daily usage. This requires to split metrics into buckets of one day worth of data. 
+It is achieved by applying a simple filter~\ref{daily_usage} in Apache Spark.
+
+\begin{lstlisting}[label={daily_usage},caption={Daily usage filter},frame=single]
+
+def daily_filter(line):
+    day_data = line.split(",")[1]
+    if (day_data == day):
+        return True
+    else:
+        return False
+
+for x in range(0, days_range):
+    split_by_day = distFile.filter(third_filter)
+
+\end{lstlisting}
 
@@
 
-\begin{algorithm}[H]
- \KwData{Task start and end time}
+The aim for this metric is to find out how long a machine is up.
+It is a key metric that will be used to calculate optimality factors later on.
+Algorithm~\ref{alg:uptime_per_machine} represents pseudo code implementation and~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
+
+\begin{algorithm}[h]
+\caption{Uptime per machine}
+\label{alg:uptime_per_machine}
+ \KwData{Task start and end time in a form of \textless K, List\textless V\textgreater\textgreater where K is Machine ID}
  \KwResult{Number of uptime hours per machine per day}
+ \algrenewcommand\algorithmicfunction{\textbf{class}}
+ \algrenewcommand\algorithmicprocedure{\textbf{method}}
+  \begin{algorithmic}[1]
+        \Procedure{map}{$\textrm{Id } key, \textrm{List } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, (startTime, endTime))$
+        \EndProcedure
+        \Procedure{reduce for min start time}{$\textrm{Id } key, \textrm{Tuple } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, (a < b) ? a : b)$
+        \EndProcedure
+        \Procedure{reduce for max end time}{$\textrm{Id } key, \textrm{Tuple } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, (a > b) ? a : b)$
+        \EndProcedure
+  \end{algorithmic}
 \end{algorithm}
 
-Input: Job start and end time logged for each machine in a form of \textless K, List\textless V\textgreater\textgreater
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label={uptime_per_machine_implementation},caption={Uptime per machine implementation in Apache Spark},frame=single] 
+min_start = distFile.map(lambda(line): 
+                (line.split(",")[0], line.split(",")[2]))
+                .reduceByKey(lambda a,b: a if a<b else b)
 
-Output: Uptime based on span between min start time and max end time \textless K,V\textgreater
+max_end = distFile.map(lambda(line): 
+                (line.split(",")[0], line.split(",")[3]))
+                .reduceByKey(lambda a,b: a if a>b else b)
+\end{lstlisting}
+\end{minipage}
 
-\myparagraph{Aggregate CPU and memory usage}
-Input: Sampled CPU and memory usage logged per job in a form of \textless K, List\textless V\textgreater\textgreater
+\myparagraph{Aggregated CPU and memory usage}
 
-Output: Aggregated resource usage \textless K,V\textgreater
+Aggregated values of CPU and memory will help to suggest better utilization strategies that guarantee SLOs.
+Algorithm~\ref{alg:agg_cpu_mem} represents pseudo code implementation and~\ref{agg_cpu_mem_implementation} contains implementation in Python with Apache Spark.
+
+\begin{algorithm}[h]
+\caption{Aggregated CPU and memory}
+\label{alg:agg_cpu_mem}
+ \KwData{CPU and memory usage in a form of \textless K, List\textless V\textgreater\textgreater where K is Machine ID}
+ \KwResult{Aggregated CPU and memory usage}
+ \algrenewcommand\algorithmicfunction{\textbf{class}}
+ \algrenewcommand\algorithmicprocedure{\textbf{method}}
+  \begin{algorithmic}[1]
+        \Procedure{map}{$\textrm{Id } key, \textrm{List } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, (cpuUsage, memoryUsage))$
+        \EndProcedure
+        \Procedure{reduce}{$\textrm{Id } key, \textrm{Int } aggValue}$
+                \State $\textsc{Emit}(\textrm{Id }key, a + b)$
+        \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label={agg_cpu_mem_implementation},caption={Aggregated CPU and memory implementation in Apache Spark},frame=single] 
+cpu = distFile.map(lambda(line): 
+		(line.split(",")[0], line.split(",")[5]))
+            .reduceByKey(lambda a,b: a + b)
+
+mem = distFile.map(lambda(line): 
+		(line.split(",")[0], line.split(",")[6]))
+            .reduceByKey(lambda a,b: a + b)
+\end{lstlisting}
+\end{minipage}
 
 \myparagraph{Calculate number of jobs per day}
-Input: Job id per machine \textless K, List\textless V\textgreater\textgreater
 
-Output: Job count per machine \textless K,V\textgreater
+Algorithm~\ref{alg:jobs_per_day} represents pseudo code implementation and~\ref{jobs_per_day} contains implementation in Python with Apache Spark.
+
+\begin{algorithm}[h]
+\caption{Daily jobs count}
+\label{alg:agg_cpu_mem}
+ \KwData{Job entry in a form of \textless K, List\textless V\textgreater\textgreater where K is Machine ID}
+ \KwResult{Aggregated jobs count}
+ \algrenewcommand\algorithmicfunction{\textbf{class}}
+ \algrenewcommand\algorithmicprocedure{\textbf{method}}
+  \begin{algorithmic}[1]
+        \Procedure{map}{$\textrm{Id } key, \textrm{List } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, 1)$
+        \EndProcedure
+        \Procedure{reduce}{$\textrm{Id } key, \textrm{Int } aggValue}$
+                \State $\textsc{Emit}(\textrm{Id }key, a + b)$
+        \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label={agg_cpu_mem_implementation},caption={Aggregated CPU and memory implementation in Apache Spark},frame=single] 
+def mapping(line):
+    machine_id = line.split(",")[0]
+    counter = 1
+    return (machine_id, counter)
+
+task_counter = distFile.map(mapping).reduceByKey(add)
+\end{lstlisting}
+\end{minipage}
 

### 2015-04-12 calculate_cost_per_hour.py

@@ -377,7 +377,7 @@ def only_uptime_mapping(line):
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
     only_uptime = distFile.map(only_uptime_mapping)
-    distinct_uptime = only_uptime.distinct()
+    distinct_uptime = only_up  time.distinct()
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
 """

### 2015-04-13 paper/paper.tex
paper updates

@@
 \begin{equation}
+\label{eq:cph}
 cph(h) = \frac{u + \sum_{i=1}^{366} pcph \times h}{365 \times h}
@@
 
+\myparagraph{Cost per hour}
+
+This is an implementation of \textit{cph} metric~\ref{eq:cph} defined in Taxonomy.
+Algorithm~\ref{alg:cost_per_hour} represents pseudo code and Listing~\ref{cost_per_hour} shows implementation in Python.
+
+\begin{algorithm}[H]
+ \caption{Cost per hour}
+ \label{alg:cost_per_hour}
+ \KwData{Upfront cost, Charge per hour, Uptime [hours per day];}
+ \KwResult{Total cost per hour for a given pricing scheme defined as \textit{cph};}
+ return (upfront + (365 * perhour * hours)) / (365 * hours)
+\end{algorithm}
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label={cost_per_hour},caption={Uptime per machine implementation in Apache Spark},frame=single] 
+def _calc_cost_per_hour(self, upfront, cost_per_hour, hours_per_day):
+        return (upfront + (365 * cost_per_hour * hours_per_day)) /
+                (365 * hours_per_day)
+\end{lstlisting}
+\end{minipage}
+
+\myparagraph{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+
+\begin{algorithm}[H]
+ \KwData{Array of cph for 2 pricing schemes;}
+ \KwResult{Intersection point when one pricing scheme becomes cheaper than the other one;}
+ read array1\;
+ read array2\;
+ \ForAll{cph in array} {
+ 	\If{$cph2 <= $cph1} {
+ 		return cph
+ 	}
+ }
+\caption{Calculate intersection point between two pricing schemes}
+\end{algorithm}
+
+\myparagraph{Calculate distance from optimality points based on uptime}
+Input: Machine uptime per day
+
+Output: Distance from optimality points. Might be positive or negative.
+
+\subsection{Recognize inefficiencies and suggest changes}
+
+\myparagraph{Threshold based report}
+
+Generate report of machines based on distance from optimality points. Use customizable threshold. 
+Results from the report will be used to calculate suggestions.
+
+\myparagraph{Visual recognition}
+
+Visual recognition as a helper method to define thresholds, especially for further incremental optimizations. 
+Useful as static thresholds can miss a whole set of machines that are just below the threshold.
+
+\begin{figure}[H]
+       \includegraphics[width=\linewidth]{figures/cubism}
+      \caption{Visualization of distance from optimality points. Green - positive, Blue - negative values.}
+        \label{fig:cubism}
+\end{figure}
+
 \myparagraph{Aggregated CPU and memory usage}
@@
 \caption{Daily jobs count}
-\label{alg:agg_cpu_mem}
+\label{alg:jobs_per_day}
  \KwData{Job entry in a form of \textless K, List\textless V\textgreater\textgreater where K is Machine ID}
@@
 \begin{minipage}{\linewidth}
-\begin{lstlisting}[label={agg_cpu_mem_implementation},caption={Aggregated CPU and memory implementation in Apache Spark},frame=single] 
+\begin{lstlisting}[label={jobs_per_day},caption={Aggregated CPU and memory implementation in Apache Spark},frame=single] 
 def mapping(line):
@@
 
-\myparagraph{Calculate cluster averages for aggregations}
-Input: Aggregated CPU, memory and number of tasks per machine per day \textless K, List\textless V\textgreater\textgreater
+\myparagraph{Calculate cluster averages}
 
-Output: Averages for CPU, memory and number of tasks \textless K, List\textless V\textgreater\textgreater
+Algorithm~\ref{alg:avg_cluster} represents pseudo code implementation.
 
-\myparagraph{Calculate distance from optimality points based on uptime}
-Input: Machine uptime per day
-
-Output: Distance from optimality points. Might be positive or negative.
-
-\subsubsection{Calculate cost per hour for a given pricing scheme}
-
-\begin{algorithm}[H]
- \KwData{Upfront cost, Per hour charge, Utilization [hours per day];}
- \KwResult{Total cost per hour for a given pricing scheme defined as \textit{cph};}
- read upfront\;
- read perhour\;
- read hours\;	
- return (upfront + (365 * perhour * hours)) / (365 * hours)
- \caption{Calculate cost per hour for a given pricing scheme}
+\begin{algorithm}[h]
+\caption{}
+\label{alg:avg_cluster}
+ \KwData{Daily CPU, memory usage and task count in a form of \textless K, List\textless V\textgreater\textgreater where K is Machine ID}
+ \KwResult{Average CPU, memory usage and task count}
+ \algrenewcommand\algorithmicfunction{\textbf{class}}
+ \algrenewcommand\algorithmicprocedure{\textbf{method}}
+  \begin{algorithmic}[1]
+        \Procedure{map}{$\textrm{Id } key, \textrm{List } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, (value, 1))$
+        \EndProcedure
+        \Procedure{combine}{$\textrm{Id } key, \textrm{List } values}$
+                \State $\textsc{Emit}(\textrm{Id }key, (currentValue + value, counter += 1))$
+        \EndProcedure
+        \Procedure{reduce}{$\textrm{Id } key, \textrm{Int } aggValue}$
+                \State $\textsc{Emit}(\textrm{Id }key, sum / counter)$
+        \EndProcedure
+  \end{algorithmic}
 \end{algorithm}
 
-\subsubsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label={avg_cluster},caption={Example code for averages in Apache Spark},frame=single] 
+    sc = SparkContext(appName="avgByKey", conf=SparkConf().set("spark.driver.host", "localhost"))
 
-\begin{algorithm}[H]
- \KwData{Array of cph for 2 pricing schemes;}
- \KwResult{Intersection point when one pricing scheme becomes cheaper than the other one;}
- read array1\;
- read array2\;
- \ForAll{cph in array} {
- 	\If{$cph2 <= $cph1} {
- 		return cph
- 	}
- }
-\caption{Calculate intersection point between two pricing schemes}
-\end{algorithm}
+    data = sc.parallelize( [(0, 2.), (0, 4.), (1, 0.), (1, 10.), (1, 20.)] )
 
-\subsection{Recognize inefficiencies and suggest changes}
+    sumCount = data.combineByKey(lambda value: (value, 1),
+                                 lambda x, value: (x[0] + value, x[1] + 1),
+                                 lambda x, y: (x[0] + y[0], x[1] + y[1]))
 
-\subsubsection{Threshold based report}
+    averageByKey = sumCount.map(lambda (label, (value_sum, count)): (label, value_sum / count))
 
-Generate report of machines based on distance from optimality points. Use customizable threshold. 
-Results from the report will be used to calculate suggestions.
-
-\subsubsection{Visual recognition}
-
-Visual recognition as a helper method to define thresholds, especially for further incremental optimizations. 
-Useful as static thresholds can miss a whole set of machines that are just below the threshold.
-
-\begin{figure}[H]
-       \includegraphics[width=\linewidth]{figures/cubism}
-      \caption{Visualization of distance from optimality points. Green - positive, Blue - negative values.}
-        \label{fig:cubism}
-\end{figure}
+    print averageByKey.collectAsMap()
+\end{lstlisting}
+\end{minipage}
 

### 2015-04-14 paper/paper.tex
paper updates

@@
 Collection per job or per process produces one log line and has scope of a measurement period.
-Having such structured data would not let to answer machine wide neither cluster wide questions.
+Such structured data itself would not answer machine wide neither cluster wide questions.
 The data has to be post-processed to allow calculation of below listed aggregations and analytics.
@@
 The data will be analyzed in terms of daily usage. This requires to split metrics into buckets of one day worth of data. 
-It is achieved by applying a simple filter~\ref{daily_usage} in Apache Spark.
+It is achieved by applying a simple filter from Listing~\ref{daily_usage} in Apache Spark.
 
@@
 It is a key metric that will be used to calculate optimality factors later on.
-Algorithm~\ref{alg:uptime_per_machine} represents pseudo code implementation and~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
+Algorithm~\ref{alg:uptime_per_machine} represents pseudo code implementation and Listing~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
 
@@
 
-\myparagraph{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
+\myparagraph{Intersection points}
+
+This is an implementation of \textit{ip(scheme1, scheme2)} defined in Taxonomy.
+This metric will be a reference point that identifies optimality points based on~\textit{cph} for various schemes.
+Algorithm~\ref{alg:intersection points} represents pseudo code and Listing~\ref{intersection_points} shows implementation in Python.
 
 \begin{algorithm}[H]
+ \label{alg:intersection points}
  \KwData{Array of cph for 2 pricing schemes;}
@@
 
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[label={intersection_points},caption={Intersection point between various pricing schemes},frame=single] 
+def _calc_intersection_point(self, first, second):
+   for i in range(len(first)):
+      if (second[i] <= first[i]):
+         return i
+\end{lstlisting}
+\end{minipage}
+
 \myparagraph{Calculate distance from optimality points based on uptime}
+
+This metric represents either positive or negative value to indicate distance from intersection points calculated
+
 Input: Machine uptime per day
@@
 
+Number of jobs per day will be a useful parameter to calculate cluster-wide averages used for suggestions.
 Algorithm~\ref{alg:jobs_per_day} represents pseudo code implementation and~\ref{jobs_per_day} contains implementation in Python with Apache Spark.

### 2015-04-15 paper/paper.tex
paper updates

@@
 \usepackage{listings}             % Include the listings-package
+\usepackage{enumerate}
 
@@
 
-\myparagraph{Calculate distance from optimality points based on uptime}
+\myparagraph{Uptime Based Distance From Intersection Points}
 
-This metric represents either positive or negative value to indicate distance from intersection points calculated
+This metric represents distance from intersection points calculated based on uptime.
+Values can be negative to indicate non-optimal profile or positive otherwise.
+There are cases when it is better to be either below or above a threshold.
+It will be specified by flag parameter calculated together with \textit{ip(scheme1, scheme2)} results.
+Algorithm~\ref{alg:distance_from_optimality} represents pseudo code implementation.
 
-Input: Machine uptime per day
+\begin{algorithm}[H]
+ \label{alg:distance_from_optimality}
+ \KwData{Uptime, Intersection Point, Spin;}
+ \KwResult{Distance from intersection point [-24, 24];}
+  \If{Spin} {
+        return Uptime - Intersection Point
+  } 
+  \Else{
+        return Intersection Point - Uptime
+  }
+\caption{Uptime Based Distance From Intersection Points}
+\end{algorithm}
 
-Output: Distance from optimality points. Might be positive or negative.
+Example algorithm results
+\begin{enumerate}
+\item Input: Uptime: 24, Optimality Point: 18, Spin: True (more optimal for values above threshold). Output: +6
+\item Input: Uptime: 16, Optimality Point: 18, Spin: True (more optimal for values above threshold). Output: -2
+\item Input: Uptime: 15, Optimality Point: 18, Spin: False (more optimal for values below threshold). Output: +3
+\item Input: Uptime: 19, Optimality Point: 18, Spin: False (more optimal for values below threshold). Output: -1
+\end{enumerate}
 
@@
 
-\myparagraph{Threshold based report}
+\myparagraph{Threshold Based Report}
 
@@
 
+The proposed strategy calculates an amount of spare CPU and memory capacity based on cluster-wide averages.
+Aggregated value for CPU and memory represents an amount of capacity gathered from machines that are utilized below average.
+Based on the amount of aggregated capacity, algorithm will attempt to squeeze in a maximum amount of machines gathered from Threshold Based Report.
+As a result, it will propose to turn off a number of machines and guarantee that SLOs will be met.
+This approach performs well when underutilized machines exist.
+For a well-balanced clusters with similar usage patterns it does not find a lot of spare capacity and would propose fairly limited improvements.
+
+\begin{algorithm}[h]
+\caption{test}
+\label{alg:squeeze-them-in}
+ \KwData{Cluster averages for CPU and memory, List of machines from Threshold Based Report, List of other machines}
+ \KwResult{List of machines that can be powered down}
+ \algrenewcommand\algorithmicfunction{\textbf{class}}
+ \algrenewcommand\algorithmicprocedure{\textbf{method}}
+ %\begin{algorithmic}[1]
+
+  %$float spare_cpu;$
+  %$float spare_memory;$
+
+ %  \ForAll{machine in Other Machines} {
+%        \If{$machine.cpu \geq cluster.cpu$} {
+%        \If{a}
+ %              \State test % spare_cpu  cluster.cpu  machine.cpu
+ %        \EndIf
+%        \If{machine.memory < cluster.memory} {
+%                spare_memory += cluster.memory - machine.memory
+%        }
+ %  }
+\ForAll{machine in Other Machines}{
+        \If{$machine.cpu < cluster.cpu$} {
+                 $spare_cpu  clustercpu  machinecpu$
+        }
+        %\EndIf
+}
+%\EndFor
+% \end{algorithmic}
+\end{algorithm}
+
 Input: results of the Threshold based report.

### 2015-04-16 paper/paper.tex
add distance alg

@@
 This metric represents distance from intersection points calculated based on uptime.
-Values can be negative to indicate non-optimal profile or positive otherwise.
-There are cases when it is better to be either below or above a threshold.
-It will be specified by flag parameter calculated together with \textit{ip(scheme1, scheme2)} results.
-Algorithm~\ref{alg:distance_from_optimality} represents pseudo code implementation.
+Result can be negative - indicates non-optimal profile - or positive otherwise.
+Positive and negative values can be yielded in either direction of time axis as presented on Figure~\ref{fig:distance}.
+It will be specified by flag parameter calculated together with \textit{ip(scheme1, scheme2)} results. Algorithm~\ref{alg:distance_from_optimality} represents pseudo code implementation.
+
+\begin{figure}[H]
+       \includegraphics[width=\linewidth]{figures/distance}
+      \caption{Distance from intersection points. Positive (green) values indicate optimal result. Negative (red) represent data for further improvement.}
+        \label{fig:distance}
+\end{figure}
 
@@
  \label{alg:distance_from_optimality}
- \KwData{Uptime, Intersection Point, Spin;}
+ \KwData{Uptime, Intersection Point, Bias;}
  \KwResult{Distance from intersection point [-24, 24];}
-  \If{Spin} {
+  \If{Bias} {
         return Uptime - Intersection Point
@@
 \begin{enumerate}
-\item Input: Uptime: 24, Optimality Point: 18, Spin: True (more optimal for values above threshold). Output: +6
-\item Input: Uptime: 16, Optimality Point: 18, Spin: True (more optimal for values above threshold). Output: -2
-\item Input: Uptime: 15, Optimality Point: 18, Spin: False (more optimal for values below threshold). Output: +3
-\item Input: Uptime: 19, Optimality Point: 18, Spin: False (more optimal for values below threshold). Output: -1
+\item Input: Uptime: 24, Optimality Point: 18, Bias: True (more optimal for values above threshold). Output: +6
+\item Input: Uptime: 16, Optimality Point: 18, Bias: True (more optimal for values above threshold). Output: -2
+\item Input: Uptime: 15, Optimality Point: 18, Bias: False (more optimal for values below threshold). Output: +3
+\item Input: Uptime: 19, Optimality Point: 18, Bias: False (more optimal for values below threshold). Output: -1
 \end{enumerate}
@@
  %  }
-\ForAll{machine in Other Machines}{
+\ForAll{machine in otherMachines}{
         \If{$machine.cpu < cluster.cpu$} {
-                 $spare_cpu  clustercpu  machinecpu$
+                 $spareCpu += clusterCpu - machineCpu$
+        }
+        \If{$machine.memory < cluster.memory$}{
+        	$spareMemory += clusterMemory - machineMemory$
         }

### 2015-04-16 paper/paper.tex
paper updates

@@
 \begin{algorithm}[h]
-\caption{test}
+\caption{Squeeze them in suggestion strategy}
 \label{alg:squeeze-them-in}
@@
  \algrenewcommand\algorithmicprocedure{\textbf{method}}
- %\begin{algorithmic}[1]
 
-  %$float spare_cpu;$
-  %$float spare_memory;$
+  $float spareCpu = 0;$
+  $float spareMemory = 0;$
+  $ID[ ] machinesToExclude;$
 
- %  \ForAll{machine in Other Machines} {
-%        \If{$machine.cpu \geq cluster.cpu$} {
-%        \If{a}
- %              \State test % spare_cpu  cluster.cpu  machine.cpu
- %        \EndIf
-%        \If{machine.memory < cluster.memory} {
-%                spare_memory += cluster.memory - machine.memory
-%        }
- %  }
 \ForAll{machine in otherMachines}{
@@
         }
-        %\EndIf
 }
-%\EndFor
-% \end{algorithmic}
+
+\ForAll{machine in machinesFromReport}{
+        \If{$(spareCpu - machine.cpu > 0) \&\& (spareMemory - machine.memory > 0)$}{
+                add machine to machinesToExclude;
+        }
+}
+
+\Return machinesToExclude;
+
 \end{algorithm}
 
-Input: results of the Threshold based report.
+\subsubsection{New capacity configuration suggestion}
 
-Computation: Calculate spare capacity based on average CPU and memory usage.
-Iterate through a pool of available machines (not included in input report) and advertise available capacity only if CPU and memory usage is lower than average.
-Aggregate spare capacity results.
+\begin{algorithm}[h]
+\caption{New configuration suggestion}
+\label{alg:new_configuration}
+ \KwData{List of machines from Threshold Based Report, List of available schemes;}
+ \KwResult{List of machines that can be powered down}
+ \algrenewcommand\algorithmicfunction{\textbf{class}}
+ \algrenewcommand\algorithmicprocedure{\textbf{method}}
 
-Output: List of machines that can be shutdown based on computed spare-capacity.
+  $float spareCpu = 0;$
+  $float spareMemory = 0;$
+  $ID[ ] machinesToExclude;$
 
-\subsubsection{New capacity configuration suggestion}
+\ForAll{machine in otherMachines}{
+        \If{$machine.cpu < cluster.cpu$} {
+                 $spareCpu += clusterCpu - machineCpu$
+        }
+        \If{$machine.memory < cluster.memory$}{
+                $spareMemory += clusterMemory - machineMemory$
+        }
+}
+
+\ForAll{machine in machinesFromReport}{
+        \If{$(spareCpu - machine.cpu > 0) \&\& (spareMemory - machine.memory > 0)$}{
+                add machine to machinesToExclude;
+        }
+}
+
+\Return machinesToExclude;
+
+\end{algorithm}
 

### 2015-04-16 paper/paper.tex
changes

@@
         }
-        \If{$machine.memory < cluster.memory$}{
+        \If{$machine.memory < cluster.memory$} {
         	$spareMemory += clusterMemory - machineMemory$
@@
 \label{alg:new_configuration}
- \KwData{List of machines from Threshold Based Report, List of available schemes;}
- \KwResult{List of machines that can be powered down}
+ \KwData{List of machines from Threshold Based Report, List of available schemes, Average number tasks per hour;}
+ \KwResult{New configuration suggestion}
  \algrenewcommand\algorithmicfunction{\textbf{class}}
@@
 
-  $float spareCpu = 0;$
-  $float spareMemory = 0;$
-  $ID[ ] machinesToExclude;$
+  $float[ ] totalCost;$
+  $float totalCpu = 0;$
+  $float totalMemory = 0;$
 
-\ForAll{machine in otherMachines}{
+\ForAll{machine in machinesFromReport}{
+        machine.cpu += totalCpu
+        machine.memory += totalMemory
+}
+
+\ForAll{scheme in availableSchemes}{
+		hoursPerDay = scheme.getHours()
         \If{$machine.cpu < cluster.cpu$} {
@@
         }
-        \If{$machine.memory < cluster.memory$}{
+        \If{$machine.memory < cluster.memory$} {
                 $spareMemory += clusterMemory - machineMemory$
@@
 }
-
-\ForAll{machine in machinesFromReport}{
-        \If{$(spareCpu - machine.cpu > 0) \&\& (spareMemory - machine.memory > 0)$}{
-                add machine to machinesToExclude;
-        }
-}
-
-\Return machinesToExclude;
 

### 2015-04-16 paper/paper.tex
updates

@@
 \textbf{Design}. This section explains design and environment conditions for the algorithm to be useful. \par
-\textbf{Implementation}. This sections explains algorithm implementation details and data analytics software that was used - Apache Spark~\cite{spark}. \par
-\textbf{Evaluation}. The proposed algorithmic approach is applied on Google Trace data~\cite{googleTrace}. \par
+\textbf{Implementation}. This section explains algorithm implementation details and data analytics software that was used - Apache Spark~\cite{spark}. \par
+\textbf{Evaluation}. The proposed algorithmic approach is applied to Google Trace data~\cite{googleTrace}. \par
 \textbf{Conclusion and Further Work}. The completed data analysis is discussed. Further work related to algorithm improvements, automation and lessons learned are presented.
@@
 \section{Motivation}
-\emph{ /* Description of the problem - cloud based company runs a mix of software services (real-time streaming, batch, web and database servers) on common hardware.
+\emph{ /* Description of the problem - cloud-based company runs a mix of software services (real-time streaming, batch, web and database servers) on common hardware.
 How do they know whether provisioned VMs run in the most optimal configuration?
@@
 Cost per hour represents a total cost of running a single machine divided by number of hours in calendar year.
-Aggregated cost is relatively easy to calculate in rented, virtualised environments.
+Aggregated cost is relatively easy to calculate in rented, virtualized environments.
 IaaS providers usually charge an X amount per hour.
@@
 This number will be considered as a partial cost per hour and defined as \textit{pcph}.
-\par
-Cost of running of physical hardware is less straightforward to calculate as it includes: power consumption, operations, hardware replacements cost, data center related costs and many others that may vary based on individual cases.
-Although more troublesome to calculate, cost per hour per host for physical infrastructure is possible to calculate.
 \par
@@
 This representation allows to notice pricing trends.
-In the example below, price drops exponentially for around first 5 hours of usage and nearly stabilizes afterwards.
+In the example below, price drops exponentially for around first 5 hours of usage and nearly stabilizes afterward.
 Light is the cheapest option for around 15 hours per day usage.
 This is better visible on Figure~\ref{fig:cph_cc2_8xlarge_zoom}.
-After 15 hours, it is better to invest in Heavy utilization scheme.
+After 15 hours, it is better to invest in the Heavy utilization scheme.
 This is the point that will be called \textit{an intersection point} between 2 pricing schemes and noted as \textit{ip(scheme1, scheme2)}.
@@
 
-In order to make data-driven analysis and suggestions for optimization we need data.
-Below list of metrics was compiled based on algorithm needs to answer critical questions regarding state of infrastructure deployment.
+In order to make data-driven analysis and suggestions for optimization, we need data.
+The below list of metrics was compiled based on algorithm needs to answer critical questions regarding the state of infrastructure deployment.
 The list is suitable to guarantee performance SLOs only for compute nodes.
@@
 \myparagraph{Start and End Time}
-Timestamps to indicate start and end of measurement period.
+Timestamps to indicate start and end of a measurement period.
 
 \myparagraph{CPU usage}
-Sampled or averaged CPU usage during measurement period.
+Sampled or averaged CPU usage during a measurement period.
 
@@
 
-Metrics gathered in Prerequisites section have limited knowledge about environment.
-Collection per job or per process produces one log line and has scope of a measurement period.
-Such structured data itself would not answer machine wide neither cluster wide questions.
+Metrics gathered in Prerequisites section have limited knowledge about the environment.
+Collection per job or per process produces one log line and has a scope of a measurement period.
+Such structured data itself would not answer machine wide neither cluster-wide questions.
 The data has to be post-processed to allow calculation of below listed aggregations and analytics.
@@
 It is a key metric that will be used to calculate optimality factors later on.
-Algorithm~\ref{alg:uptime_per_machine} represents pseudo code implementation and Listing~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
+Algorithm~\ref{alg:uptime_per_machine} represents pseudo-code implementation and Listing~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
 
@@
 This is an implementation of \textit{cph} metric~\ref{eq:cph} defined in Taxonomy.
-Algorithm~\ref{alg:cost_per_hour} represents pseudo code and Listing~\ref{cost_per_hour} shows implementation in Python.
+Algorithm~\ref{alg:cost_per_hour} represents pseudo-code and Listing~\ref{cost_per_hour} shows implementation in Python.
 
@@
 This metric will be a reference point that identifies optimality points based on~\textit{cph} for various schemes.
-Algorithm~\ref{alg:intersection points} represents pseudo code and Listing~\ref{intersection_points} shows implementation in Python.
+Algorithm~\ref{alg:intersection points} represents pseudo-code and Listing~\ref{intersection_points} shows implementation in Python.
 
@@
 
-This metric represents distance from intersection points calculated based on uptime.
-Result can be negative - indicates non-optimal profile - or positive otherwise.
-Positive and negative values can be yielded in either direction of time axis as presented on Figure~\ref{fig:distance}.
-It will be specified by flag parameter calculated together with \textit{ip(scheme1, scheme2)} results. Algorithm~\ref{alg:distance_from_optimality} represents pseudo code implementation.
+This metric represents a distance from intersection points calculated based on uptime.
+The result can be negative - indicates non-optimal profile - or positive otherwise.
+Positive and negative values can be yielded in either direction of the time axis as presented in Figure~\ref{fig:distance}.
+It will be specified by flag parameter calculated together with \textit{ip(scheme1, scheme2)} results. Algorithm~\ref{alg:distance_from_optimality} represents pseudo-code implementation.
 
@@
        \includegraphics[width=\linewidth]{figures/distance}
-      \caption{Distance from intersection points. Positive (green) values indicate optimal result. Negative (red) represent data for further improvement.}
+      \caption{Distance from intersection points. Positive (green) values indicate the optimal result. Negative (red) represent data for further improvement.}
         \label{fig:distance}
@@
 Aggregated values of CPU and memory will help to suggest better utilization strategies that guarantee SLOs.
-Algorithm~\ref{alg:agg_cpu_mem} represents pseudo code implementation and~\ref{agg_cpu_mem_implementation} contains implementation in Python with Apache Spark.
+Algorithm~\ref{alg:agg_cpu_mem} represents pseudo-code implementation and~\ref{agg_cpu_mem_implementation} contains implementation in Python with Apache Spark.
 
@@
 
-Number of jobs per day will be a useful parameter to calculate cluster-wide averages used for suggestions.
-Algorithm~\ref{alg:jobs_per_day} represents pseudo code implementation and~\ref{jobs_per_day} contains implementation in Python with Apache Spark.
+The number of jobs per day will be a useful parameter to calculate cluster-wide averages used for suggestions.
+Algorithm~\ref{alg:jobs_per_day} represents pseudo-code implementation and~\ref{jobs_per_day} contains implementation in Python with Apache Spark.
 
@@
 
-Algorithm~\ref{alg:avg_cluster} represents pseudo code implementation.
+Algorithm~\ref{alg:avg_cluster} represents pseudo-code implementation.
 
@@
 Aggregated value for CPU and memory represents an amount of capacity gathered from machines that are utilized below average.
-Based on the amount of aggregated capacity, algorithm will attempt to squeeze in a maximum amount of machines gathered from Threshold Based Report.
+Based on the amount of aggregated capacity, the algorithm will attempt to squeeze in a maximum amount of machines gathered from Threshold Based Report.
 As a result, it will propose to turn off a number of machines and guarantee that SLOs will be met.
 This approach performs well when underutilized machines exist.
-For a well-balanced clusters with similar usage patterns it does not find a lot of spare capacity and would propose fairly limited improvements.
+For well-balanced clusters with similar usage patterns, it does not find a lot of spare capacities and would propose fairly limited improvements.
 
@@
 \ForAll{machine in machinesFromReport}{
-        machine.cpu += totalCpu
-        machine.memory += totalMemory
+        machine.cpu += totalCpu;
+        machine.memory += totalMemory;
 }
@@
 \ForAll{scheme in availableSchemes}{
-		hoursPerDay = scheme.getHours()
-        \If{$machine.cpu < cluster.cpu$} {
-                 $spareCpu += clusterCpu - machineCpu$
-        }
-        \If{$machine.memory < cluster.memory$} {
-                $spareMemory += clusterMemory - machineMemory$
-        }
+		hoursPerDay = scheme.getHours();
+        cpuCap = hoursPerDay * clusterCpuAvgPerHour;
+        memCap = hoursPerDay * memoryCpuAvgPerHour;
+
+        vmsReqCpu = totalCpu / cpuCap;
+        vmsReqMem = totalMemory / memCap;
+
+        totalCost[scheme] = min(vmsReqCpu, vmsReqMem) * scheme.cph;
 }
 
+\Return{min(totalCost)}
+
 \end{algorithm}
-
-Input: results of the Threshold based report.
-
-Computation: Compute a minimal number of VMs for each configuration based on requested capacity
-
-Output: A number and configuration of the cheapest computed configuration.
 

### 2015-04-17 paper/paper.tex
updates

@@
 %%%%%%%%%%%%%%%%%%%%%%  
-%%% IMPLEMENTATION 
-%%%%%%%%%%%%%%%%%%%%%%
-
-\chapter{Implementation}
-
-\emph{/* Implementation quirks related to make it work. Mostly map/reduce jobs written for Apache Spark. */}
-
-%%%%%%%%%%%%%%%%%%%%%%  
 %%% EVALUATION!!!
@@
 
-\subsection{Recognize inefficiencies and suggest changes}
-
-\subsubsection{Threshold based report}
-
-Generate report of machines based on distance from optimality points. Use customizable threshold. 
-Results from the report will be used to calculate suggestions.
-
-\subsubsection{Visual recognition}
-
 \section{Methodology and setup description}
 
-\section{Algorithm walkthrough}
-\subsection{Calculate cost per hour for a given pricing scheme}
-Results for ccs2.xlarge based on data published for AWS~\cite{AWS:2014} \par
-\csvautotabular{out}
-\subsection{Calculate intersection points to find out the best pricing scheme for a number of hours/day utilization}
-%\includegraphics[width=\linewidth]{cph}
-\subsection{Collect VM usage data}
-\subsection{Group usage metrics by machine}
-\subsection{Calculate number of hours/day, distance from optimality points, CPU and memory utilization}
-\subsection{Suggest changes to the least optimal VMs considering SLAs based on CPU and memory utilization}
+\section{Uptime Based Distance From Intersection Points}
+
+Machines in Google Cluster Trace dataset do not have any pricing data attached. For further evaluation purposes, we have to assign example pricing schemes to each virtual machine.
+
+Total number of machines: 12583
+0.5 and 0.5,0.25 = 10595 = 84.2\% = heavy
+rest = 1988 = 15.8\% = medium
+
+Pricing: cc2.8xlarge
+c_medium = upfront : 4146, perhour : 0.54 
+c_heavy = upfront : 5000, perhour : 0.361
 

### 2015-04-17 calculate_cost_per_hour.py
calc updates

@@ -338,7 +338,7 @@ Schema after transformation (splitted per day)
 
 
 """
-
+"""
 def mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
     machine_id = splits[0].strip()
@@ -364,6 +364,7 @@ for x in range(0, 30):
     joined_file_distinct.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x))
     day += 1
 """
+"""
 
 GET ONLY UPTIME
 
@@ -381,6 +382,27 @@ for x in range(0, 30):
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
 """
+
+
+"""
+
+GET ONLY CPU = 0.5 and MEMORY = (0.5 OR 0.25)
+
+"""
+
+def cpu_and_mem(line):
+    splits = line.split(",")
+    cpu = float(splits[4].strip())
+    mem = float(splits[5].strip())
+
+    if (cpu == 0.5 and (mem == 0.25 or mem == 0.5)):
+        return True
+    else:
+        return False
+
+distFile = sc.textFile("/Users/ksmuga/workspace/data/clusterdata-2011-2/machine_events/*", use_unicode=False)
+heavies = distFile.filter(cpu_and_mem)
+heavies.map(lambda x: 1).reduce(add)
 
 """
 --------------------------------------------------------

### 2015-04-17 calculate_cost_per_hour.py
convertion to float updates

@@ -378,7 +378,7 @@ def only_uptime_mapping(line):
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
     only_uptime = distFile.map(only_uptime_mapping)
-    distinct_uptime = only_up  time.distinct()
+    distinct_uptime = only_up time.distinct()
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
 """
@@ -392,8 +392,14 @@ GET ONLY CPU = 0.5 and MEMORY = (0.5 OR 0.25)
 
 def cpu_and_mem(line):
     splits = line.split(",")
-    cpu = float(splits[4].strip())
-    mem = float(splits[5].strip())
+    try:
+        cpu = float(splits[4])
+    except ValueError:
+        cpu = int(splits[4])
+    try:
+        mem = float(splits[5])
+    except ValueError:
+        mem = int(splits[5])    
 
     if (cpu == 0.5 and (mem == 0.25 or mem == 0.5)):
         return True

### 2015-04-17 calculate_cost_per_hour.py
handle empty strings

@@ -392,14 +392,14 @@ GET ONLY CPU = 0.5 and MEMORY = (0.5 OR 0.25)
 
 def cpu_and_mem(line):
     splits = line.split(",")
-    try:
-        cpu = float(splits[4])
-    except ValueError:
-        cpu = int(splits[4])
-    try:
-        mem = float(splits[5])
-    except ValueError:
-        mem = int(splits[5])    
+    cpu = splits[4]
+    mem = splits[5]
+    
+    if cpu == "" or mem == "":
+        return False
+
+    cpu = float(splits[4])
+    mem = float(splits[5])
 
     if (cpu == 0.5 and (mem == 0.25 or mem == 0.5)):
         return True

### 2015-04-17 calculate_cost_per_hour.py
asdf

@@ -378,7 +378,7 @@ def only_uptime_mapping(line):
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
     only_uptime = distFile.map(only_uptime_mapping)
-    distinct_uptime = only_up time.distinct()
+    distinct_uptime = only_uptime.distinct()
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
 """

### 2015-04-17 calculate_cost_per_hour.py
calc distance

@@ -368,13 +368,15 @@ for x in range(0, 30):
 
 GET ONLY UPTIME
 
+"""
+
 
 def only_uptime_mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
     machine_id = splits[0].strip()
     uptime = float(splits[2].strip()) - float(splits[1].strip())
     return (machine_id, uptime)
-
+"""
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
     only_uptime = distFile.map(only_uptime_mapping)
@@ -382,33 +384,64 @@ for x in range(0, 30):
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
 """
-
-
 """
 
 GET ONLY CPU = 0.5 and MEMORY = (0.5 OR 0.25)
 
+Schema after transformation (splitted per day)
+        1,machine ID
+        2,uptime (how many hours machine is on)
+        3,distance from optimality point
+        4,total cost per day
+        5,cpu usage
+        6,memory usage
+        7,number of tasks
+        8,MEMORY capacity
+        9,CPU capacity
+
 """
 
+vm = Entity(c_name, c_light, c_medium, c_heavy)
+
 def cpu_and_mem(line):
-    splits = line.split(",")
-    cpu = splits[4]
-    mem = splits[5]
-    
-    if cpu == "" or mem == "":
-        return False
+    ip = vm.i_medium_to_high
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
 
-    cpu = float(splits[4])
-    mem = float(splits[5])
+    mach_id = splits[0].strip()
+    start = float(splits[1].strip())
+    end = float(splits[2].strip())
 
-    if (cpu == 0.5 and (mem == 0.25 or mem == 0.5)):
-        return True
+    cpu_usage = float(splits[3].strip())
+    memory_usage = float(splits[4].strip())
+
+    up = end - start
+
+    cpu_cap = splits[6].strip()
+    mem_cap = splits[7].strip()
+    tasks = splits[8].strip()
+
+    if cpu_cap != "":
+        cpu_cap = float(cpu_cap)
+
+    if mem_cap != "":
[…]

### 2015-04-17 calculate_cost_per_hour.py
entity first

@@ -384,6 +384,80 @@ for x in range(0, 30):
     distinct_uptime.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-fifth-uptime-only-day-" + str(x))
 
 """
+""" 
+------------------
+PRICING INPUT DATA
+------------------
+"""
+# values for size: cc2.8xlarge region: us-east
+# 1 year reservation
+c_name = 'cc2.8xlarge'
+c_light = { "upfront" : "1762", "perhour" : "0.904" }
+c_medium = { "upfront" : "4146", "perhour" : "0.54" }
+c_heavy = { "upfront" : "5000", "perhour" : "0.361" }
+
+# values for size: m2.4xlarge region: us-east
+# 1 year reservation
+m_name = 'm2.4xlarge'
+m_light = { "upfront" : "1088", "perhour" : "0.676" }
+m_medium = { "upfront" : "2604", "perhour" : "0.52" }
+m_heavy = { "upfront" : "3156", "perhour" : "0.272" }
+
+# values for size: hs1.8xlarge region: us-east
+# 1 year reservation
+h_name = 'hs1.8xlarge'
+h_light = { "upfront" : "3968", "perhour" : "2.24" }
+h_medium = { "upfront" : "9200", "perhour" : "1.38" }
+h_heavy = { "upfront" : "11213", "perhour" : "0.92" }
+
+""" 
+-----------------------------
+INSTANCE ENTITY UTILITY CLASS
+-----------------------------
+"""
+class Entity:
+    def __init__(self, name, light, medium, heavy):
+        self.name = name
+        self.light = light
+        self.medium = medium
+        self.heavy = heavy
+        self.ligth_cph = [ self._calc_cost_per_hour(self.light, x) for x in range(1,25) ]
+        self.medium_cph = [ self._calc_cost_per_hour(self.medium, x) for x in range(1,25) ]
+        self.heavy_cph = [ self._calc_cost_per_hour(self.heavy, x) for x in range(1,25) ]
+        self.i_light_to_medium = self._calc_intersection_point(self.ligth_cph, self.medium_cph)
+        self.i_medium_to_high = self._calc_intersection_point(self.medium_cph, self.heavy_cph)
+        self.i_light_to_high = self._calc_intersection_point(self.ligth_cph, self.heavy_cph)
+
+    def _calc_cost_per_hour(self, mode, hours_per_day):
+        upfront = int(mode["upfront"])
+        cost_per_hour = float(mode["perhour"])
[…]

### 2015-04-17 calculate_cost_per_hour.py
calc distance points

@@ -508,14 +508,17 @@ def cpu_and_mem(line):
         distance = up - ip
     else:
         cph = vm.medium_cph[int(up)]
-        cost = cph_m * up
+        cost = cph * up
         distance = ip - up
 
     return (mach_id, (up, distance, cost, cpu_usage, memory_usage, tasks, cpu_cap, mem_cap))
 
-distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-1/*", use_unicode=False)
-heavies = distFile.map(cpu_and_mem)
-heavies.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-1")
+for x in range(0, 30):
+    distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x) + "/part*", use_unicode=False)
+    heavies = distFile.map(cpu_and_mem)
+    heavies.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-" + str(x))
+
+
 
 """
 --------------------------------------------------------

### 2015-04-17 calculate_cost_per_hour.py
calc distance points

@@ -503,11 +503,11 @@ def cpu_and_mem(line):
     cost = 0
     distance = 0
     if (cpu_cap == 0.5 and mem_cap > 0.24):
-        cph = vm.heavy_cph[int(up)]
+        cph = vm.heavy_cph[int(up) - 1]
         cost = cph * up
         distance = up - ip
     else:
-        cph = vm.medium_cph[int(up)]
+        cph = vm.medium_cph[int(up) - 1]
         cost = cph * up
         distance = ip - up
 

### 2015-04-17 calculate_cost_per_hour.py
calc agg cost

@@ -338,7 +338,7 @@ Schema after transformation (splitted per day)
 
 
 """
-"""
+
 def mapping(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
     machine_id = splits[0].strip()
@@ -348,7 +348,7 @@ def mapping(line):
     mem = float(splits[4].strip())
     total_mem = float(splits[5].strip())
     return (machine_id, (uptime, uptime_total, total_cpu, mem, total_mem))
-
+"""
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-forth-day-" + str(x) + "/part*", use_unicode=False)
     basic_forth = distFile.map(mapping)
@@ -512,13 +512,52 @@ def cpu_and_mem(line):
         distance = ip - up
 
     return (mach_id, (up, distance, cost, cpu_usage, memory_usage, tasks, cpu_cap, mem_cap))
-
+"""
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x) + "/part*", use_unicode=False)
     heavies = distFile.map(cpu_and_mem)
     heavies.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-" + str(x))
 
+"""
 
+"""" CALC TOTAL COST 
+
+Results:
+1. total cost
+2. total cost for heavy
+3. total cost for medium 
+4. total uptime 
+"""
+def cost(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
+    cost = splits[3].strip()
+    return cost
+
+def cost_for_heavy(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    cpu_cap = splits[4].strip()
+    mem_cap = splits[5].strip()
+    if cpu_cap != "":
+        cpu_cap = float(cpu_cap)
+    if mem_cap != "":
+        mem_cap = float(mem_cap)
+    if (cpu_cap == 0.5 and mem_cap > 0.24):
+        return True
+    else:
+        return False
+
+def total_uptime(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    up = float(splits[1].strip())
+    return up
+
[…]

### 2015-04-17 calculate_cost_per_hour.py
total cost updates

@@ -535,8 +535,8 @@ def cost(line):
 
 def cost_for_heavy(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
-    cpu_cap = splits[4].strip()
-    mem_cap = splits[5].strip()
+    cpu_cap = splits[7].strip()
+    mem_cap = splits[8].strip()
     if cpu_cap != "":
         cpu_cap = float(cpu_cap)
     if mem_cap != "":
@@ -552,11 +552,11 @@ def total_uptime(line):
     return up
 
 distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-1/*", use_unicode=False)
-cost = distFile.map(cost).reduce(add)
+total_cost = distFile.map(cost).reduce(add)
 heavy_cost = distFile.filter(cost_for_heavy).map(cost).reduce(add)
 total_up = distFile.map(total_uptime).reduce(add)
 
-joined = cost.join(heavy_cost).join(total_up)
+joined = total_cost.join(heavy_cost).join(total_up)
 joined.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-day-1")
 
 """

### 2015-04-18 calculate_cost_per_hour.py
join type error

@@ -556,8 +556,9 @@ total_cost = distFile.map(cost).reduce(add)
 heavy_cost = distFile.filter(cost_for_heavy).map(cost).reduce(add)
 total_up = distFile.map(total_uptime).reduce(add)
 
-joined = total_cost.join(heavy_cost).join(total_up)
-joined.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-day-1")
+total_cost.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-cost-day-1")
+heavy_cost.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-heavy-cost-day-1")
+total_up.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-uptime-day-1")
 
 """
 --------------------------------------------------------

### 2015-04-18 calculate_cost_per_hour.py
negatives

@@ -530,7 +530,7 @@ Results:
 """
 def cost(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
-    cost = splits[3].strip()
+    cost = float(splits[3].strip())
     return cost
 
 def cost_for_heavy(line):
@@ -551,14 +551,30 @@ def total_uptime(line):
     up = float(splits[1].strip())
     return up
 
-distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-1/*", use_unicode=False)
+def total_distance(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    distance = float(splits[2].strip())
+    return distance  
+"""
+distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
 total_cost = distFile.map(cost).reduce(add)
 heavy_cost = distFile.filter(cost_for_heavy).map(cost).reduce(add)
 total_up = distFile.map(total_uptime).reduce(add)
+total_dis = distFile.map(total_distance).reduce(add)
+"""
 
-total_cost.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-cost-day-1")
-heavy_cost.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-heavy-cost-day-1")
-total_up.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh-uptime-day-1")
+def filter_negatives(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    distance = float(splits[2].strip())
+    if distance < 0:
+        return True
+    else:
+        return False
+
+distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
+negatives = distFile.filter(filter_negatives)
+negatives.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh")
+
 
 """
 --------------------------------------------------------

### 2015-04-18 calculate_cost_per_hour.py
add negative id mapping

@@ -555,6 +555,22 @@ def total_distance(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
     distance = float(splits[2].strip())
     return distance  
+
+def total_cpu(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    cpu = float(splits[4].strip())
+    return cpu
+
+def total_mem(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    mem = float(splits[5].strip())
+    return mem
+
+def total_tasks(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    tasks = float(splits[6].strip())
+    return tasks
+
 """
 distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
 total_cost = distFile.map(cost).reduce(add)
@@ -563,6 +579,8 @@ total_up = distFile.map(total_uptime).reduce(add)
 total_dis = distFile.map(total_distance).reduce(add)
 """
 
+
+
 def filter_negatives(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
     distance = float(splits[2].strip())
@@ -570,10 +588,33 @@ def filter_negatives(line):
         return True
     else:
         return False
-
+"""
 distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
 negatives = distFile.filter(filter_negatives)
 negatives.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-seventh")
+
+distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-seventh/*", use_unicode=False)
+number_of_machines_to_tune = distFile.map(lambda x: x.split(",")[0]).distinct().collect()
+"""
+# count 
+"""
+distFile.map(lambda x: (x.split(",")[0], 1)).reduceByKey(add).saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-eight")
[…]

### 2015-04-19 paper/paper.tex
updates

@@
 
+Measurement period = 29 days
 Total number of machines: 12583
@@
 c_heavy = upfront : 5000, perhour : 0.361
+
+\begin{center}
+    \begin{tabular}{| l | l | l | l | l | l | l | l |}
+    \hline
+    Total cost & Cost for heavy & Cost for medium & Total Uptime & Total distance & CPU & Memory & Tasks \\ \hline
+    \$8 177 048 & \$7 479 648 & \$697 400 & 8 705 743h & +3 215 231 & 15972217.38 & 30676363.36 & 1234716525 \\ \hline
+    \end{tabular}
+\end{center}
+
+Cost per task = \$8 177 048 / 1234716525 = \$0.00662261161525
+cost per hour = \$8 177 048 / 8 705 743 = 0.93927054818871
+
+\section{Recognize inefficiencies and suggest changes}
+
+\myparagraph{Threshold Based Report}
+
+The report will filter all machines that reported non-optimal - negative distance from intersection point - configuration during measurement period (29 days).
+
+Number of machines to tune: 932
+CPU capacity to accommodate: 1 320 781.36
+Memory capacity to accommodate: 4 286 742.06
+Number of tasks to execute: 118869614
+Cost of running non-optimal machines: \$657 200
+Cost of heavy: \$29.56
+Cost of medium: \$657 170.44
+Total uptime: 648519
+cost per hour = \$1.01338588383687
+Cost per task =  \$657 200 / 118869614 = \$0.00552874681666
+
+\myparagraph{Calculate cluster averages}
+
+Avg cpu usage per machine = 15972217.38/12583 = 1269.35/29 = 43.77
+Avg mem usage per machine = 30676363.36/12583 = 2437.92/29 = 84.06
+Avg tasks number per machine = 1234716525/12583 = 98125.77/29 = 3383.65
+
+\myparagraph{Squeeze-them-in suggestion strategy}
+
+total spare cpu = 3 697 171.44
+total spare mem = 6 685 407.83
+total distance: +3 500 037
+total uptime = 8056601
+cost per hour = \$7 519 848 / 8056601 = 0.93337723935938
+new total cost = \$8 177 048 - \$657 200 = \$7 519 848
+new cost per task = \$7 519 848 / 1234716525 = 0.00609034369245
+
+new averages 
+cpu per machine = 15972217.38/11651 = 1370.88 / 29 = 47.27
+Avg mem usage per machine = 30676363.36/11651 = 2632.94/29 = 90.79
+Avg tasks number per machine = 1234716525/11651 = 105975.15/29 = 3654.31
 

### 2015-04-19 calculate_cost_per_hour.py

@@ -577,6 +577,9 @@ total_cost = distFile.map(cost).reduce(add)
 heavy_cost = distFile.filter(cost_for_heavy).map(cost).reduce(add)
 total_up = distFile.map(total_uptime).reduce(add)
 total_dis = distFile.map(total_distance).reduce(add)
+cpu = distFile.map(total_cpu).reduce(add)
+mem = distFile.map(total_mem).reduce(add)
+tasks = distFile.map(total_tasks).reduce(add)
 """
 
 
@@ -616,6 +619,41 @@ def filter_for_stats(line):
 distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-seventh/*", use_unicode=False)
 stats = distFile.filter(filter_for_stats)
 
+
+"""
+squeeze them in!
+
+"""
+
+def filter_for_other(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    m_id = int(splits[0].strip())
+    if m_id in negatives_ids:
+        return False
+    else:
+        return True
+
+def filter_for_squeeze_cpu(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    cpu = float(splits[4].strip())
+    if cpu < 43.77:
+        return 43.77 - cpu
+    else:
+        return 0
+
+def filter_for_squeeze_mem(line):
+    splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
+    mem = float(splits[5].strip())
+    if mem < 84.06:
+        return 84.06 - mem
+    else:
+        return 0
+
+distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
+total_spare_cpu = distFile.filter(filter_for_other).map(filter_for_squeeze_cpu).reduce(add)
+total_spare_mem = distFile.filter(filter_for_other).map(filter_for_squeeze_mem).reduce(add)
+total_distance = distFile.filter(filter_for_other).map(total_distance).reduce(add)
+total_up = distFile.filter(filter_for_other).map(total_uptime).reduce(add)
 
 """
 --------------------------------------------------------

### 2015-04-19 calculate_cost_per_hour.py
algorithm updates

@@ -582,8 +582,6 @@ mem = distFile.map(total_mem).reduce(add)
 tasks = distFile.map(total_tasks).reduce(add)
 """
 
-
-
 def filter_negatives(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")    
     distance = float(splits[2].strip())
@@ -615,10 +613,10 @@ def filter_for_stats(line):
         return True
     else:
         return False
-
+"""
 distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-seventh/*", use_unicode=False)
 stats = distFile.filter(filter_for_stats)
-
+"""
 
 """
 squeeze them in!
@@ -654,6 +652,39 @@ total_spare_cpu = distFile.filter(filter_for_other).map(filter_for_squeeze_cpu).
 total_spare_mem = distFile.filter(filter_for_other).map(filter_for_squeeze_mem).reduce(add)
 total_distance = distFile.filter(filter_for_other).map(total_distance).reduce(add)
 total_up = distFile.filter(filter_for_other).map(total_uptime).reduce(add)
+
+
+"""
+New configuration suggestion
+
+  $float[ ] totalCost;$
+  $float totalCpu = 0;$
+  $float totalMemory = 0;$
+
+\ForAll{machine in machinesFromReport} {
+    machine.cpu += totalCpu;
+    machine.memory += totalMemory;
+}
+
+\ForAll{scheme in availableSchemes} {
+    hoursPerDay = scheme.getHours();
+    cpuCap = hoursPerDay * clusterCpuAvgPerHour;
+    memCap = hoursPerDay * memoryCpuAvgPerHour;
+
+    vmsReqCpu = totalCpu / cpuCap;
+    vmsReqMem = totalMemory / memCap;
+
+    totalCost[scheme] = min(vmsReqCpu, vmsReqMem) * scheme.cph;
+}
+
+\Return{min(totalCost)}
+
+"""
+distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
+distFile.filter(filter_for_stats).map(total_cpu)
+distFile.filter(filter_for_stats).map(total_mem)
+
+
 
 """
 --------------------------------------------------------

### 2015-04-19 paper/paper.tex
updates

--- 2015-04-17 paper/paper.tex
@@
 \section{Methodology and setup description}
+Tools:
+Apache Spark, Google Data Trace, Workstation
 
@@
 Machines in Google Cluster Trace dataset do not have any pricing data attached. For further evaluation purposes, we have to assign example pricing schemes to each virtual machine. 
+Assumption: we take ip(medium, heavy) and assign heavy scheme to cpu and mem of platform B. The rest is medium.
 
+Measurement period = 29 days
 Total number of machines: 12583
@@
 c_heavy = upfront : 5000, perhour : 0.361
+
+Pricing: m2.4xlarge
+m_medium = upfront : 2604, perhour : 0.52 
+m_heavy = upfront : 3156, perhour : 0.272
+
+\myparagraph{Cluster statistics}
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Indicator & Scheme 1 & Scheme 2 \\
+    \hline
+    Total cost & \$8 177 048 & \$5 637 418 \\
+    \hline
+    KDI & +3 215 231 & +5 431 137 \\
+    \hline
+    Total CPU & 15972217.38 & 15972217.38 \\
+    \hline
+    Total Memory & 30676363.36 & 30676363.36 \\
+    \hline
+    Number of tasks & 1234716525 & 1234716525 \\
+    \hline
+    \end{tabular}
+\end{center}
+
+\section{Recognize inefficiencies and suggest changes}
+
+\myparagraph{Threshold Based Report}
+
+The report will filter all machines that reported non-optimal - negative distance from intersection point - configuration during measurement period (29 days).
+
+\myparagraph{Scheme 1}
+Number of machines to tune: 932
+CPU capacity to accommodate: 1 320 781.36
+Memory capacity to accommodate: 4 286 742.06
+Number of tasks to execute: 118869614
+Cost of running non-optimal machines: \$657 200
+Total uptime: 648519
+cost per hour = \$1.01338588383687
+Cost per task =  \$657 200 / 118869614 = \$0.00552874681666
+
+\myparagraph{Scheme 2}
+Number of machines to tune: 937
+CPU capacity to accommodate: 1 327 202.18
+Memory capacity to accommodate: 4 303 983.31
+Number of tasks to execute: 119367988
+Cost of running non-optimal machines: \$532 840.76
+Total uptime: 651899
+
+\myparagraph{Calculate cluster averages}
+
+Avg cpu usage per machine = 15972217.38/12583 = 1269.35/29 = 43.77
+Avg mem usage per machine = 30676363.36/12583 = 2437.92/29 = 84.06
+Avg tasks number per machine = 1234716525/12583 = 98125.77/29 = 3383.65
+
+\myparagraph{Squeeze-them-in suggestion strategy}
+
+\myparagraph{Scheme 1}
+total spare cpu = 3 697 171.44
+total spare mem = 6 685 407.83
+total distance: +3 500 037
+total uptime = 8056601
+cost per hour = \$7 519 848 / 8056601 = 0.93337723935938
+new total cost = \$8 177 048 - \$657 200 = \$7 519 848
+new cost per task = \$7 519 848 / 1234716525 = 0.00609034369245
+
+new averages 
+cpu per machine = 15972217.38/11651 = 1370.88 / 29 = 47.27
+Avg mem usage per machine = 30676363.36/11651 = 2632.94/29 = 90.79
+Avg tasks number per machine = 1234716525/11651 = 105975.15/29 = 3654.31
+
+\myparagraph{Scheme 2}
+total spare cpu = 3 694 240.56 
+total spare mem = 6 682 956.59
+total distance: +3 501 467
+total uptime = 8053221
+
+\myparagraph{New capacity configuration suggestion}
+
+Input
+
+c_name = cc2.8xlarge
+c_light = "upfront" : "1762", "perhour" : "0.904" 
+c_medium =  "upfront" : "4146", "perhour" : "0.54" 
+c_heavy =  "upfront" : "5000", "perhour" : "0.361"
+
+# values for size: m2.4xlarge region: us-east
+# 1 year reservation
+m_name = 'm2.4xlarge'
+m_light =  "upfront" : "1088", "perhour" : "0.676" 
+m_medium =  "upfront" : "2604", "perhour" : "0.52" 
+m_heavy =  "upfront" : "3156", "perhour" : "0.272"
+
+vm1 = Entity(c_name, c_light, c_medium, c_heavy)
+vm2 = Entity(m_name, m_light, m_medium, m_heavy)
+
+>>> calc_options(uptime, vm2, 29)
+(1597, 3727, 972, 567 273.49, 998 186.23, 410 041.68)
+>>> calc_options(uptime, vm1, 29)
+(1397, 1720, 972, 770 417.58, 876 376.34, 604 274.61)
 

### 2015-04-19 calculate_cost_per_hour.py

@@ -475,10 +475,12 @@ Schema after transformation (splitted per day)
 
 """
 
-vm = Entity(c_name, c_light, c_medium, c_heavy)
+vm1 = Entity(c_name, c_light, c_medium, c_heavy)
+vm2 = Entity(m_name, m_light, m_medium, m_heavy)
+vm3 = Entity(h_name, h_light, h_medium, h_heavy)
 
 def cpu_and_mem(line):
-    ip = vm.i_medium_to_high
+    ip = vm2.i_medium_to_high
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
 
     mach_id = splits[0].strip()
@@ -503,20 +505,21 @@ def cpu_and_mem(line):
     cost = 0
     distance = 0
     if (cpu_cap == 0.5 and mem_cap > 0.24):
-        cph = vm.heavy_cph[int(up) - 1]
+        cph = vm2.heavy_cph[int(up) - 1]
         cost = cph * up
         distance = up - ip
     else:
-        cph = vm.medium_cph[int(up) - 1]
+        cph = vm2.medium_cph[int(up) - 1]
         cost = cph * up
         distance = ip - up
 
     return (mach_id, (up, distance, cost, cpu_usage, memory_usage, tasks, cpu_cap, mem_cap))
+
 """
 for x in range(0, 30):
     distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-fifth-day-" + str(x) + "/part*", use_unicode=False)
     heavies = distFile.map(cpu_and_mem)
-    heavies.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-" + str(x))
+    heavies.saveAsTextFile("/Users/ksmuga/workspace/data/out/transformation-sixth-beta-day-" + str(x))
 
 """
 
@@ -528,6 +531,7 @@ Results:
 3. total cost for medium 
 4. total uptime 
 """
+
 def cost(line):
     splits = line.replace("\"","").replace("(", "").replace(")", "").replace("\'","").split(",")
     cost = float(splits[3].strip())
@@ -589,27 +593,28 @@ def filter_negatives(line):
         return True
     else:
         return False
-"""
-distFile = sc.textFile("/Users/ksmuga/workspace/data/out/transformation-sixth-day-*/*", use_unicode=False)
-negatives = distFile.filter(filter_negatives)
[…]

### 2015-04-20 paper/paper.tex
evaluation updates

@@
 
-\emph{/* Evaluation will go through each of Design stages and will perform necessary calculations. Some examples attached below. */}
+\section{Methodology}
 
-\section{Methodology and setup description}
-Tools:
-Apache Spark, Google Data Trace, Workstation
+Evaluation process can be splitted into four phases:
+\begin{enumerate}
+\item collection and transformation of pricing and usage data,
+\item calculation of key indicators (\textit{cph, ip(scheme1, scheme2), KDI}) and cluster characteristics (\textit{total cost, performance averages}), 
+\item suggestion of changes based on Algorithm~\ref{alg:squeeze-them-in} and~\ref{alg:new_configuration},
+\item validation of positive correlation between \textit{KDI} and \textit{total cost} changes.
+\end{enumerate}
 
-\section{Uptime Based Distance From Intersection Points}
+\section{Setup description}
+
+\myparagraph{Workstation}
+Hardware - Intel Xeon 6-core x86\_64 2.4GHz, L3 cache 24MB, RAM 28GB. \\
+Operating System - OS X Yosemite 10.10.2
+
+\myparagraph{Platform}
+Apache Spark 1.2.1\footnote{\url{https://spark.apache.org}} - an engine for large-scale data processing.
+All of map reduce jobs were written using Python API.
+Jobs were executed using standalone cluster mode with settings specified in Table~\ref{tab:spark_conf}.
+Job run times varied widely based on input data size and complexity of computations - from 20hs+ for transformations on 100G+ data to 1 minute aggregations on 100-200M input.
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    \textbf{Setting name} & \textbf{Value} \\
+    \hline
+    spark.master & local \\
+    \hline
+    spark.driver.memory & 8G \\
+    \hline
+    spark.executor.memory & 20G \\
+    \hline
+    spark.default.parallelism & 500 \\
+    \hline
+    \end{tabular}
+\end{center}
+\label{tab:spark_conf}
+\caption{Apache Spark standalone cluster configuration settings}
+\end{table}
+
+Settings were tweaked during computation mostly due to frequent Out Of Memory errors.
+Default driver and executor settings - 256M and 512M respectively - were not enough to handle the volume of input data.
+Further OOMs occurred during \textit{RRD.reduceByKey}\footnote{\url{https://spark.apache.org/docs/1.2.1/api/python/pyspark.html#pyspark.RDD.reduceByKey}} operations as it required to compute a fairly large hash map.
+Tweaks of \textit{spark.default.parallelism} setting worked the best as it enforced more granular input data chunking.
+
+\section{Input data}
+
+\myparagraph{Pricing data}
+
+Amazon Web Services\footnote{\url{http://aws.amazon.com/}} offered variable pricing schemes based on uptime.
+Customer could choose an instance configuration and total cost would be calculated based on a number of hours of uptime per day.
+The offering was splitted into three options:
+\begin{enumerate}
+\item light - low upfront and high per hour cost,
+\item medium - medium upfront and medium per hour cost,
+\item heavy - high upfront and low per hour cost.
+\end{enumerate}
+
+For evaluation purposes we choose two pricing schemes for two different VM configurations~\cite{AWS:2014}. Pricing details are specified in Table~\ref{tab:aws_pricing}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l | l |}
+    \hline
+    \textbf{Name} & \textbf{Scheme} & \textbf{Upfront (\$)} & \textbf{Per hour (\$)} \\
+    \hline
+    cc2.8xlarge & light & 1762 & 0.904 \\
+    \hline
+    cc2.8xlarge & medium & 4146 & 0.54 \\
+    \hline
+    cc2.8xlarge & heavy & 5000 & 0.361 \\
+    \hline
+    m2.4xlarge & light & 1088 & 0.676 \\
+    \hline
+    m2.4xlarge & medium & 2604 & 0.52 \\
+    \hline
+    m2.4xlarge & heavy & 3156 & 0.272 \\
+    \hline
+    \end{tabular}
+\end{center}
+\label{tab:aws_pricing}
+\caption{AWS pricing for reserved cc2.8xlarge and m2.4xlarge instances}
+\end{table}
+
+Pricing: m2.4xlarge
+m_light =  upfront : 1088, perhour : 0.676
+m_medium = upfront : 2604, perhour : 0.52 
+m_heavy = upfront : 3156, perhour : 0.272
+
+\myparagraph{Usage data}
+
+\section{Calculate cph and ip}
+
+\section{Transform initial Google data schema to below per day}
+
+Schema after transformation (splitted per day)
+        1,machine ID,INTEGER,YES
+        2,uptime (how many hours machine is on),INTEGER,YES
+        3,uptime based on summarized task time,INTEGER,YES
+        4,total CPU usage,INTEGER,YES
+        5,CPU capacity, INTEGER, YES
+        6,total assigned memory usage,FLOAT,NO
+        7,total maximum memory usage,FLOAT,NO
+        8,MEMORY capacity,FLOAT,YES
+        9,NUMBER of tasks per machine per day
+
+\section{Calculate KDI and total cost on google trace data }
 
@@
 rest = 1988 = 15.8\% = medium
-
-Pricing: cc2.8xlarge
-c_medium = upfront : 4146, perhour : 0.54 
-c_heavy = upfront : 5000, perhour : 0.361
-
-Pricing: m2.4xlarge
-m_medium = upfront : 2604, perhour : 0.52 
-m_heavy = upfront : 3156, perhour : 0.272
 
@@
 
-\section{Recognize inefficiencies and suggest changes}
+\myparagraph{Cluster SLO averages}
 
-\myparagraph{Threshold Based Report}
+Avg cpu usage per machine = 15972217.38/12583 = 1269.35/29 = 43.77
+Avg mem usage per machine = 30676363.36/12583 = 2437.92/29 = 84.06
+Avg tasks number per machine = 1234716525/12583 = 98125.77/29 = 3383.65
+
+\section{Recognize inefficiencies with Threshold-based Report}
 
@@
 
-\myparagraph{Calculate cluster averages}
-
-Avg cpu usage per machine = 15972217.38/12583 = 1269.35/29 = 43.77
-Avg mem usage per machine = 30676363.36/12583 = 2437.92/29 = 84.06
-Avg tasks number per machine = 1234716525/12583 = 98125.77/29 = 3383.65
-
-\myparagraph{Squeeze-them-in suggestion strategy}
+\section{Squeeze-them-in suggestion strategy}
 
@@
 total spare mem = 6 685 407.83
-total distance: +3 500 037
+new KDI: +3 500 037
+old KDI: +3 215 231
+KDI change: +8.14\%
 total uptime = 8056601
+new total cost = \$8 177 048 - \$657 200 = \$7 519 848
+cost change: -8.04\%
 cost per hour = \$7 519 848 / 8056601 = 0.93337723935938
-new total cost = \$8 177 048 - \$657 200 = \$7 519 848
-new cost per task = \$7 519 848 / 1234716525 = 0.00609034369245
+new cost per task = \$7 519 848 / 118869614 = 0.06326131419927
 
 new averages 
-cpu per machine = 15972217.38/11651 = 1370.88 / 29 = 47.27
-Avg mem usage per machine = 30676363.36/11651 = 2632.94/29 = 90.79
-Avg tasks number per machine = 1234716525/11651 = 105975.15/29 = 3654.31
+cpu per machine = 15972217.38/11651 = 1370.88 / 29 = 47.27 +7.4\%
+Avg mem usage per machine = 30676363.36/11651 = 2632.94/29 = 90.79 +7.41\%
+Avg tasks number per machine = 1234716525/11651 = 105975.15/29 = +7.4\%
 
@@
 total spare mem = 6 682 956.59
-total distance: +3 501 467
+new KDI: +5 913 933
+old KDI: +5 431 137
+KDI change: +8.16\%
 total uptime = 8053221
+new total cost = \$5 637 418 - \$532 840.76 = \$5 104 577.24
+cost change: -9.45\%
+cost per hour = \$5 104 577.24 / 8053221 = 0.63385535303204
+new cost per task = \$5 104 577.24 / 119367988 = 0.04276336834964
 
-\myparagraph{New capacity configuration suggestion}
+new averages 
+cpu per machine = 15972217.38/11646 = 1371.48 / 29 = 47.29 +7.44\%
+Avg mem usage per machine = 30676363.36/11646 = 2634.07 / 29 = 90.83 +7.45\%
+Avg tasks number per machine = 1234716525/11646 = 106020.65 / 29 = 3655.88 +7.45\%
 
-Input
-
-c_name = cc2.8xlarge
-c_light = "upfront" : "1762", "perhour" : "0.904" 
-c_medium =  "upfront" : "4146", "perhour" : "0.54" 
-c_heavy =  "upfront" : "5000", "perhour" : "0.361"
-
-# values for size: m2.4xlarge region: us-east
-# 1 year reservation
-m_name = 'm2.4xlarge'
-m_light =  "upfront" : "1088", "perhour" : "0.676" 
-m_medium =  "upfront" : "2604", "perhour" : "0.52" 
-m_heavy =  "upfront" : "3156", "perhour" : "0.272"
+\section{New capacity configuration suggestion}
 
@@
 (1397, 1720, 972, 770 417.58, 876 376.34, 604 274.61)
+
+\section{Validate influence of suggestions and correlation with KDI}
 
@@
 \bibitem{googleTrace} https://code.google.com/p/googleclusterdata
-\bibitem{AWS:2014} http://forecastcloudy.net/2012/04/02/amazon-web-services-aws-ec2-pricing-data/ https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-light-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-medium-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-heavy-linux.json
+\bibitem{AWS:2014} Access: Apr 20th, 2015, https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-light-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-medium-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-heavy-linux.json
 \end{thebibliography}

### 2015-04-20 paper/paper.tex
bibtex updates

@@
 \usepackage{enumerate}
+\usepackage{cite}
 
@@
 \textbf{Design}. This section explains design and environment conditions for the algorithm to be useful. \par
-\textbf{Implementation}. This section explains algorithm implementation details and data analytics software that was used - Apache Spark~\cite{spark}. \par
-\textbf{Evaluation}. The proposed algorithmic approach is applied to Google Trace data~\cite{googleTrace}. \par
+\textbf{Implementation}. This section explains algorithm implementation details and data analytics software that was used - Apache Spark\footnote{\url{https://spark.apache.org}}. \par
+\textbf{Evaluation}. The proposed algorithmic approach is applied to Google Trace data~\cite{clusterdata:Reiss2011}. \par
 \textbf{Conclusion and Further Work}. The completed data analysis is discussed. Further work related to algorithm improvements, automation and lessons learned are presented.
@@
 	\includegraphics[width=\linewidth]{figures/cc2_8xlarge}
-	\caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:2014}}
+	\caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}}
 	\label{fig:cc2_8xlarge}
@@
 	\includegraphics[width=\linewidth]{figures/cph_cc2_8xlarge_zoom}
-	\caption{Zoomed \textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:2014}}
+	\caption{Zoomed \textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}}
 	\label{fig:cph_cc2_8xlarge_zoom}
@@
 
-For evaluation purposes we choose two pricing schemes for two different VM configurations~\cite{AWS:2014}. Pricing details are specified in Table~\ref{tab:aws_pricing}
+For evaluation purposes we choose two pricing schemes for two different VM configurations~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}.
+Pricing details are specified in Table~\ref{tab:aws_pricing}
 
@@
 
-Pricing: m2.4xlarge
-m_light =  upfront : 1088, perhour : 0.676
-m_medium = upfront : 2604, perhour : 0.52 
-m_heavy = upfront : 3156, perhour : 0.272
+\myparagraph{Usage data}
+Google released 29 day cluster usage traces of more than 12k machines running over 650k jobs~\cite{clusterdata:Reiss2011}. As analyzed in ~\cite{clusterdata:Reiss2012b}
 
-\myparagraph{Usage data}
+\myparagraph{Pricing assignment to usage data}
 
@@
 \newpage
-\begin{thebibliography}{99}
-\bibitem{spark} https://spark.apache.org
-\bibitem{googleTrace} https://code.google.com/p/googleclusterdata
-\bibitem{AWS:2014} Access: Apr 20th, 2015, https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-light-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-medium-linux.json https://a0.awsstatic.com/pricing/1/deprecated/ec2/ri-heavy-linux.json
-\end{thebibliography}
+
 \label{endpage}
-
+\bibliography{bib}{}
+\bibliographystyle{plain}
+\end{document}
 \end{document}

### 2015-04-20 paper/paper.tex
more evaluation

@@
 \textbf{Evaluation}. The proposed algorithmic approach is applied to Google Trace data~\cite{clusterdata:Reiss2011}. \par
-\textbf{Conclusion and Further Work}. The completed data analysis is discussed. Further work related to algorithm improvements, automation and lessons learned are presented.
+\textbf{Conclusion and Future Work}. The completed data analysis is discussed. Further work related to algorithm improvements, automation and lessons learned are presented.
 
@@
 \end{center}
+\caption{Apache Spark standalone cluster configuration settings} 
 \label{tab:spark_conf}
-\caption{Apache Spark standalone cluster configuration settings}
 \end{table}
@@
 \end{center}
+\caption{AWS pricing for reserved cc2.8xlarge and m2.4xlarge instances}
 \label{tab:aws_pricing}
-\caption{AWS pricing for reserved cc2.8xlarge and m2.4xlarge instances}
 \end{table}
@@
 \myparagraph{Usage data}
-Google released 29 day cluster usage traces of more than 12k machines running over 650k jobs~\cite{clusterdata:Reiss2011}. As analyzed in ~\cite{clusterdata:Reiss2012b}
+
+Google released 29-day cluster usage traces of more than 12k machines running over 650k jobs~\cite{clusterdata:Reiss2011}.
+An analysis~\cite{clusterdata:Reiss2012b} revealed that the workload is heterogeneous, which is somewhat unique compared to other traces and makes a good fit for testing.
+The trace contains data about jobs submitted by users.
+Each job is made of tasks.
+Each task reports metrics about resource consumption - e.g.
+CPU, memory - and timestamps.
+The trace does not contain information about the purpose of the jobs, which limits SLO assumptions to resource utilization.
+Resource utilization data is provided in normalized units - the strongest configuration is 1 and the rest of configurations are described in relation to that value.
 
@@
 
-\section{Calculate cph and ip}
+Google trace does not provide any information about pricing schemes of included machines.
+For evaluation purposes, we will produce example configurations of different pricing schemes that could exist within a real cluster.
+
+The majority of machines within Google cluster i.e.
+82\%, have following configuration: 0.5 CPU, 0.5 RAM or 0.5 CPU, 0.25 RAM.
+We will assign heavy pricing scheme for these machines.
+For the rest of the cluster, we will assign medium pricing scheme.
+
+The light scheme is a good fit for relatively short periods of time, e.g.
+during peak.
+An analysis of the trace~\cite{clusterdata:Reiss2012b} shows that jobs with the highest peak-to-mean ratio are production priority jobs and they constitute only 7\% of all jobs.
+Therefore, we will not assign any light pricing schemes for this evaluation purposes.
+
+We will refer to cc2.8xlarge and m2.4xlarge reserved instances pricing data as \textit{Scheme 1} and \textit{Scheme 2}.
+
+\section{Pricing data indicators}
+
+\subsection{Cost per hour}
+\myparagraph{Scheme1}
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l | l |}
+    \hline
+    \textbf{Hour} & \textbf{Light (price/hour)} & \textbf{Medium (price/hour)} & \textbf{Heavy (price/hour)} \\
+    \hline
+	1&5.7313972602739724&11.898904109589042&14.059630136986302 \\
+    \hline
+	2&3.317698630136986&6.2194520547945205&7.21031506849315 \\
+    \hline
+	3&2.513132420091324&4.326301369863014&4.9272100456621 \\
+    \hline
+	4&2.110849315068493&3.3797260273972602&3.785657534246575 \\    
+	\hline
+5&1.8694794520547946&2.811780821917808&3.1007260273972603 \\
+	\hline
+6&1.7085662100456622&2.433150684931507&2.64410502283105 \\
+	\hline 
+7&1.593628180039139&2.162700587084149&2.317947162426614 \\
+\hline 
+8&1.5074246575342467&1.9598630136986301&2.0733287671232876 \\
+\hline
+9&1.4403774733637749&1.8021004566210044&1.8830700152207003 \\
+\hline
+10&1.3867397260273973&1.6758904109589041&1.73086301369863 \\ 
+\hline
+11&1.342854296388543&1.5726276463262765&1.6063300124533002 \\ 
+\hline
+12&1.3062831050228312&1.4865753424657535&1.5025525114155251 \\
+\hline
+13&1.2753382507903057&1.4137618545837725&1.414740779768177 \\
+\hline
+14&1.2488140900195697&1.3513502935420745&1.3394735812133072 \\
+\hline
+15&1.2258264840182649&1.2972602739726027&1.2742420091324203 \\
+\hline
+16&1.2057123287671234&1.249931506849315&1.2171643835616437 \\
+\hline
+17&1.1879645447219984&1.208170829975826&1.1668017727639 \\
+\hline
+18&1.1721887366818875&1.1710502283105022&1.12203500761035 \\ 
+\hline
+19&1.1580735400144198&1.1378370583994233&1.0819805335255948 \\
+\hline
+20&1.1453698630136988&1.107945205479452&1.045931506849315 \\
+\hline
+21&1.1338760600130462&1.0809001956947162&1.0133157208088714 \\
+\hline
+22&1.1234271481942715&1.0563138231631384&0.9836650062266501 \\
+\hline
+23&1.1138868374032165&1.0338653960690887&0.9565926146515782 \\ 
+\hline
+24&1.1051415525114157&1.0132876712328769&0.9317762557077626 \\ 
+    \hline
+    \end{tabular}
+\end{center}
+\caption{cph results for cc2.8xlarge reserved instance}
+\label{tab:cph:scheme1}
+\end{table}
+
+\myparagraph{Scheme2}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l | l |}
+    \hline
+    \textbf{Hour} & \textbf{Light (price/hour)} & \textbf{Medium (price/hour)} & \textbf{Heavy (price/hour)} \\
+    \hline
+1&3.656821917808219&7.654246575342467&8.918575342465754\\ 
+ \hline 
+2&2.1664109589041094&4.087123287671233&4.595287671232876\\ 
+ \hline 
+3&1.669607305936073&2.898082191780822&3.1541917808219178\\ 
+ \hline 
+4&1.4212054794520548&2.303561643835616&2.4336438356164383\\ 
+ \hline 
+5&1.2721643835616436&1.9468493150684931&2.0013150684931507\\ 
+ \hline 
+6&1.1728036529680366&1.709041095890411&1.713095890410959\\ 
+ \hline 
+7&1.1018317025440314&1.5391780821917809&1.5072250489236791\\ 
+ \hline 
+8&1.0486027397260274&1.411780821917808&1.352821917808219\\ 
+ \hline 
+9&1.0072024353120244&1.3126940639269407&1.2327305936073059\\ 
+ \hline 
+10&0.9740821917808219&1.2334246575342467&1.1366575342465755\\ 
+ \hline 
+11&0.9469838107098382&1.1685678704856788&1.058052303860523\\ 
+ \hline 
+12&0.9244018264840183&1.1145205479452056&0.9925479452054796\\ 
+ \hline 
+13&0.9052939936775553&1.0687881981032665&0.9371211801896734\\ 
+ \hline 
+14&0.8889158512720158&1.0295890410958906&0.8896125244618396\\ 
+ \hline 
+15&0.8747214611872147&0.9956164383561644&0.8484383561643836\\ 
+ \hline 
+16&0.8623013698630138&0.9658904109589042&0.8124109589041095\\ 
+ \hline 
+17&0.8513424657534246&0.9396615632554393&0.7806220789685737\\ 
+ \hline 
+18&0.8416012176560121&0.9163470319634702&0.7523652968036529\\ 
+ \hline 
+19&0.8328853640951694&0.8954866618601299&0.7270829127613554\\ 
+ \hline 
+20&0.825041095890411&0.8767123287671232&0.7043287671232877\\ 
+ \hline 
+21&0.8179439008480104&0.8597260273972603&0.6837416829745597\\ 
+ \hline 
+22&0.8114919053549191&0.8442839352428394&0.6650261519302615\\ 
+ \hline 
+23&0.8056009529481835&0.8301846337105421&0.6479380583680763\\ 
+ \hline 
+24&0.8002009132420091&0.8172602739726028&0.6322739726027398 \\
+ \hline
+
+    \end{tabular}
+\end{center}
+\caption{cph results for m2.4xlarge reserved instance}
+\label{tab:cph:scheme1}
+\end{table}
+
+\subsection{Intersection points}
+\myparagraph{Scheme1}
+Intersection points:
+Light/medium : 17
+Medium/heavy : 13
+Light / heavy : 16
+\myparagraph{Scheme2}
+Intersection points:
+Light/medium : Never (surprisingly medium is never cheaper than light)
+Medium/heavy : 6
+Light / heavy : 14
 
@@
 %%%%%%%%%%%%%%%%%%%%%%
-%%% OPTIONAL!!!
-%%% CAPACITY PLANNING 
-%%% STRATEGIES    
-%%%%%%%%%%%%%%%%%%%%%%
-
-%\chapter{Capacity Planning Strategies}
-
-%\section{Peak-to-mean ratio based resource assignment}
-
-%\section{Only heavy instances}
-
-%%%%%%%%%%%%%%%%%%%%%%
-%%% CONCLUSION 
+%%% CONCLUSION AND 
+%%% FUTURE WORK
 %%%%%%%%%%%%%%%%%%%%%%
@@
 \chapter{Conclusion and Future Work}
+
+\myparagraph{Conclusion}
+It works and is great!
+
+\myparagraph{Future Work}
+\begin{enumerate}
+\item Visual recognition of underperforming machines
+\item Prediction trend to generate suggestions
+\end{enumerate}
 

### 2015-04-21 paper/paper.tex
updates

@@
 
+TO BE REWRITTEN!
+
 Cost per hour is calculated for each of available pricing schemes.
@@
 For the example below, cc2.8xlarge instance's intersection point between Light and Heavy schemes is \textit{ip(light, heavy) = 16}.
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{figures/cc2_8xlarge}
-	\caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}}
-	\label{fig:cc2_8xlarge}
-\end{figure}
 
@@
 \subsection{Intersection points}
-\myparagraph{Scheme1}
-Intersection points:
-Light/medium : 17
-Medium/heavy : 13
-Light / heavy : 16
-\myparagraph{Scheme2}
-Intersection points:
-Light/medium : Never (surprisingly medium is never cheaper than light)
-Medium/heavy : 6
-Light / heavy : 14
-
-\section{Transform initial Google data schema to below per day}
-
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    & \textbf{Scheme 1} & \textbf{Scheme 2} \\
+    \hline
+    light/medium & 17 & never \\
+    \hline
+    light/heavy & 16 & 6 \\
+    \hline
+    medium/heavy & 13 & 14 \\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Intersection points for Scheme 1 and Scheme 2}
+\label{tab:intersection_points}
+\end{table}
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{figures/cc2_8xlarge}
+  \caption{\textit{cph} results for Amazon AWS cc2.8xlarge reserved instance rented for 1 year~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}}
+  \label{fig:cc2_8xlarge}
+\end{figure}
+
+\begin{figure}[H]
+  \includegraphics[width=\linewidth]{figures/m2_4xlarge}
+  \caption{\textit{cph} results for Amazon AWS m2.4xlarge reserved instance rented for 1 year~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}}
+  \label{fig:m2_4xlarge}
+\end{figure}
+
+\section{Google trace data schema transformation}
+
+We used machine and tasks data for further evaluation. Machine information regarding ID and capacity can be found in machine events  .
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    & \textbf{Column name} \\
+    \hline
+    1 & timestamp\\
+    \hline
+    2 & machine id\\
+     \hline
+    3 & event type\\
+     \hline
+    4 & platform id\\
+     \hline
+    5 & capacity: CPU\\
+     \hline
+    6 & capacity: memory\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Machine events table schema from Google trace}
+\label{tab:machine_events}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    & \textbf{Column name} \\
+    \hline
+    1 & start time of the measurement period\\
+    \hline
+    2 & end time of the measurement period\\
+     \hline
+    3 & job ID\\
+     \hline
+    4 & task index\\
+     \hline
+    5 & machine ID \\
+     \hline
+    6 & mean CPU usage rate \\
+    \hline
+    7 & canonical memory usage\\
+    \hline
+    8 & assigned memory usage\\
+    \hline
+    9 & unmapped page cache memory usage\\
+    \hline
+    10 & total page cache memory usage\\
+    \hline
+    11 & maximum memory usage\\
+    \hline
+    12 & mean disk I/O time\\
+    \hline
+    13 & mean local disk space used\\
+    \hline
+    14 & maximum CPU usage\\
+    \hline
+    15 & maximum disk IO time\\
+    \hline
+    16 & cycles per instruction (CPI)\\
+    \hline
+    17 & memory accesses per instruction\\
+    \hline
+    18 & sample portion\\
+    \hline
+    19 & aggregation type \\
+    \hline    
+    20 & sampled CPU usage\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Machine events table schema from Google trace}
+\label{tab:task_events}
+\end{table}
+
+Transformation to below schema. 
+- uptime per day
+- total CPU usage per day
+- total memory usage per day
+- total number of tasks per day
+
+\begin{lstlisting}
 Schema after transformation (splitted per day)
         1,machine ID,INTEGER,YES
-        2,uptime (how many hours machine is on),INTEGER,YES
-        3,uptime based on summarized task time,INTEGER,YES
+        2,uptime
         4,total CPU usage,INTEGER,YES
@@
         6,total assigned memory usage,FLOAT,NO
-        7,total maximum memory usage,FLOAT,NO
         8,MEMORY capacity,FLOAT,YES
         9,NUMBER of tasks per machine per day
+\end{lstlisting}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    & \textbf{Column name} \\
+    \hline
+    1 & timestamp\\
+    \hline
+    2 & machine id\\
+     \hline
+    3 & event type\\
+     \hline
+    4 & platform id\\
+     \hline
+    5 & capacity: CPU\\
+     \hline
+    6 & capacity: memory\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Machine events table schema from Google trace}
+\label{tab:new_schema}
+\end{table}
+
+~\cite{test}
 

### 2015-04-21 paper/paper.tex
updates

@@
 \begin{minipage}{\linewidth}
-\begin{lstlisting}[label={cost_per_hour},caption={Uptime per machine implementation in Apache Spark},frame=single] 
+\begin{lstlisting}[label={cost_per_hour},caption={Cost per hour implementation in Python},frame=single] 
 def _calc_cost_per_hour(self, upfront, cost_per_hour, hours_per_day):
@@
 
-\myparagraph{Calculate cluster averages}
-
-Algorithm~\ref{alg:avg_cluster} represents pseudo-code implementation.
-
-\begin{algorithm}[h]
-\caption{}
-\label{alg:avg_cluster}
- \KwData{Daily CPU, memory usage and task count in a form of \textless K, List\textless V\textgreater\textgreater where K is Machine ID}
- \KwResult{Average CPU, memory usage and task count}
- \algrenewcommand\algorithmicfunction{\textbf{class}}
- \algrenewcommand\algorithmicprocedure{\textbf{method}}
-  \begin{algorithmic}[1]
-        \Procedure{map}{$\textrm{Id } key, \textrm{List } values}$
-                \State $\textsc{Emit}(\textrm{Id }key, (value, 1))$
-        \EndProcedure
-        \Procedure{combine}{$\textrm{Id } key, \textrm{List } values}$
-                \State $\textsc{Emit}(\textrm{Id }key, (currentValue + value, counter += 1))$
-        \EndProcedure
-        \Procedure{reduce}{$\textrm{Id } key, \textrm{Int } aggValue}$
-                \State $\textsc{Emit}(\textrm{Id }key, sum / counter)$
-        \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\begin{minipage}{\linewidth}
-\begin{lstlisting}[label={avg_cluster},caption={Example code for averages in Apache Spark},frame=single] 
-    sc = SparkContext(appName="avgByKey", conf=SparkConf().set("spark.driver.host", "localhost"))
-
-    data = sc.parallelize( [(0, 2.), (0, 4.), (1, 0.), (1, 10.), (1, 20.)] )
-
-    sumCount = data.combineByKey(lambda value: (value, 1),
-                                 lambda x, value: (x[0] + value, x[1] + 1),
-                                 lambda x, y: (x[0] + y[0], x[1] + y[1]))
-
-    averageByKey = sumCount.map(lambda (label, (value_sum, count)): (label, value_sum / count))
-
-    print averageByKey.collectAsMap()
-\end{lstlisting}
-\end{minipage}
-
 \subsubsection{Squeeze-them-in suggestion strategy}
@@
 
+\section{Input data schema transformation and aggregations}
+
+Google trace data contains a lot of detailed information.
+This paper focuses on uptime and we extracted only relevant data for further evaluation.
+Machine and task events original schema is presented on Tables~\ref{tab:machine_events} and~\ref{tab:task_events} and fields selected for further transformation and aggregation are marked in bold text.
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    & \textbf{Column name} \\
+    \hline
+    1 & \textbf{timestamp}\\
+    \hline
+    2 & \textbf{machine id} \\
+     \hline
+    3 & event type\\
+     \hline
+    4 & platform id\\
+     \hline
+    5 & \textbf{capacity: CPU}\\
+     \hline
+    6 & \textbf{capacity: memory}\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Machine events table schema from Google trace}
+\label{tab:machine_events}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    & \textbf{Column name} \\
+    \hline
+    1 & \textbf{start time of the measurement period}\\
+    \hline
+    2 & \textbf{end time of the measurement period}\\
+     \hline
+    3 & job ID\\
+     \hline
+    4 & task index\\
+     \hline
+    5 & \textbf{machine ID} \\
+     \hline
+    6 & mean CPU usage rate \\
+    \hline
+    7 & \textbf{canonical memory usage}\\
+    \hline
+    8 & assigned memory usage\\
+    \hline
+    9 & unmapped page cache memory usage\\
+    \hline
+    10 & total page cache memory usage\\
+    \hline
+    11 & maximum memory usage\\
+    \hline
+    12 & mean disk I/O time\\
+    \hline
+    13 & mean local disk space used\\
+    \hline
+    14 & maximum CPU usage\\
+    \hline
+    15 & maximum disk IO time\\
+    \hline
+    16 & cycles per instruction (CPI)\\
+    \hline
+    17 & memory accesses per instruction\\
+    \hline
+    18 & sample portion\\
+    \hline
+    19 & aggregation type \\
+    \hline    
+    20 & \textbf{sampled CPU usage}\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Task events table schema from Google trace}
+\label{tab:task_events}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l |}
+    \hline
+    & \textbf{Column name} \\
+    \hline
+    1& machine ID\\
+    \hline
+    2 & total CPU usage\\
+     \hline
+    3 & CPU capacity\\
+     \hline
+    4 & total assigned memory usage\\
+     \hline
+    5 & memory capacity\\
+     \hline
+    6 & number of tasks\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Aggregated statistics from Google trace data}
+\label{tab:new_schema}
+\end{table}
+
+Table~\ref{tab:new_schema} is a new schema produced from join of capacity metrics from machine events and aggregations computed from tasks events data.
+
+\myparagraph{Daily usage}
+As a first step, the task usage data is grouped into separate buckets to represent 1 day worth of data based on timestamps.
+Listing~\ref{daily_usage} shows Apache Spark filter implementation in Python.
+This transformation created 29 folders of data.
+
+\myparagraph{Uptime}
+Uptime per day per machine was calculated based on aggregation of task duration times for a given day.
+Algorithm~\ref{alg:uptime_per_machine} shows design and Listing~\ref{uptime_per_machine_implementation} represents implementation of this aggregation.
+Total number of hours in cluster is 8705743, which gives 23.36 hours daily per machine on average.
+
+Transformation to below schema. 
+- total number of tasks per day
+
+\myparagraph{Total CPU and memory usage}
+CPU and memory aggregation logic is presented in Algorithm~\ref{alg:agg_cpu_mem} and implemented in Listing~\ref{agg_cpu_mem_implementation}.
+The total CPU consumption of the cluster is 15972217.38, which gives 43.77 daily per machine on average.
+The total memory usage is 30676363.36, which is 84.06 on average.
+
+\myparagraph{Total number of tasks}
+An aggregated number of tasks is calculated based on Algorithm~\ref{alg:jobs_per_day} and implementation presented in Listing~\ref{jobs_per_day}.
+The total number of tasks is 1234716525, which gives 3383.65 tasks per day per machine on average.
+
 \section{Pricing data indicators}
@@
 \subsection{Cost per hour}
+
+Based on Formula~\ref{eq:cph} implemented in Python and presented on Listing~\ref{cost_per_hour}, we calculated \textit{cost per hour} for both analyzed schemes.
+Results are presented on Table~\ref{tab:cph:scheme1} and~\ref{tab:cph:scheme2} in respect of number of hours per day a machine would run.
+The cost per hour drops when uptime per day rises.
+The same trend exhibits for both schemes.
+
 \myparagraph{Scheme1}
@@
  \hline 
-
     \end{tabular}
@@
 \caption{cph results for m2.4xlarge reserved instance}
-\label{tab:cph:scheme1}
+\label{tab:cph:scheme2}
 \end{table}
@@
 \subsection{Intersection points}
+
+Based on Formula~\ref{alg:intersection points} implemented in Python and presented on Listing~\ref{intersection_points} we calculated \textit{intersection points} between pricing schemes.
+Table~\ref{tab:intersection_points} shows results of the calculation.
+Intuitively, heavy schemes is the cheapest option for highly utilized machines.
+An interesting finding is that medium is never cheaper than light in case of Scheme 2.
+Figures~\ref{fig:cc2_8xlarge} and~\ref{fig:m2_4xlarge} represent graphically \textit{cph} results and \textit{intersection points} can be noticed.
 
@@
 
-\section{Google trace data schema transformation}
-
-We used machine and tasks data for further evaluation. Machine information regarding ID and capacity can be found in machine events  .
+\subsection{Key distance indicator and total cost}
+
+Key distance indicator got calculated for both schemes based on Algorithm~\ref{alg:distance_from_optimality}.
 
@@
 \begin{center}
-    \begin{tabular}{| l | l |}
-    \hline
-    & \textbf{Column name} \\
-    \hline
-    1 & timestamp\\
-    \hline
-    2 & machine id\\
-     \hline
-    3 & event type\\
-     \hline
-    4 & platform id\\
-     \hline
-    5 & capacity: CPU\\
-     \hline
-    6 & capacity: memory\\
-    \hline
+    \begin{tabular}{| l | l | l |}
+    \hline
+    Indicator & Scheme 1 & Scheme 2 \\
+    \hline
+    Total cost & \$8 177 048 & \$5 637 418 \\
+    \hline
+    KDI & +3 215 231 & +5 431 137 \\
+    \hline
+%    Total CPU & 15972217.38 & 15972217.38 \\
+%    \hline
+%    Total Memory & 30676363.36 & 30676363.36 \\
+%    \hline
+%    Number of tasks & 1234716525 & 1234716525 \\
+%    \hline
     \end{tabular}
 \end{center}
-\caption{Machine events table schema from Google trace}
-\label{tab:machine_events}
+\caption{KDI and total cost for cluster}
+\label{tab:kdi_and_cost}
 \end{table}
-
-\begin{table}[h]
-\begin{center}
-    \begin{tabular}{| l | l |}
-    \hline
-    & \textbf{Column name} \\
-    \hline
-    1 & start time of the measurement period\\
-    \hline
-    2 & end time of the measurement period\\
-     \hline
-    3 & job ID\\
-     \hline
-    4 & task index\\
-     \hline
-    5 & machine ID \\
-     \hline
-    6 & mean CPU usage rate \\
-    \hline
-    7 & canonical memory usage\\
-    \hline
-    8 & assigned memory usage\\
-    \hline
-    9 & unmapped page cache memory usage\\
-    \hline
-    10 & total page cache memory usage\\
-    \hline
-    11 & maximum memory usage\\
-    \hline
-    12 & mean disk I/O time\\
-    \hline
-    13 & mean local disk space used\\
-    \hline
-    14 & maximum CPU usage\\
-    \hline
-    15 & maximum disk IO time\\
-    \hline
-    16 & cycles per instruction (CPI)\\
-    \hline
-    17 & memory accesses per instruction\\
-    \hline
-    18 & sample portion\\
-    \hline
-    19 & aggregation type \\
-    \hline    
-    20 & sampled CPU usage\\
-    \hline
-    \end{tabular}
-\end{center}
-\caption{Machine events table schema from Google trace}
-\label{tab:task_events}
-\end{table}
-
-Transformation to below schema. 
-- uptime per day
-- total CPU usage per day
-- total memory usage per day
-- total number of tasks per day
-
-\begin{lstlisting}
-Schema after transformation (splitted per day)
-        1,machine ID,INTEGER,YES
-        2,uptime
-        4,total CPU usage,INTEGER,YES
-        5,CPU capacity, INTEGER, YES
-        6,total assigned memory usage,FLOAT,NO
-        8,MEMORY capacity,FLOAT,YES
-        9,NUMBER of tasks per machine per day
-\end{lstlisting}
-
-\begin{table}[h]
-\begin{center}
-    \begin{tabular}{| l | l |}
-    \hline
-    & \textbf{Column name} \\
-    \hline
-    1 & timestamp\\
-    \hline
-    2 & machine id\\
-     \hline
-    3 & event type\\
-     \hline
-    4 & platform id\\
-     \hline
-    5 & capacity: CPU\\
-     \hline
-    6 & capacity: memory\\
-    \hline
-    \end{tabular}
-\end{center}
-\caption{Machine events table schema from Google trace}
-\label{tab:new_schema}
-\end{table}
-
-~\cite{test}
-
-\section{Calculate KDI and total cost on google trace data }
-
-Machines in Google Cluster Trace dataset do not have any pricing data attached. For further evaluation purposes, we have to assign example pricing schemes to each virtual machine. 
-Assumption: we take ip(medium, heavy) and assign heavy scheme to cpu and mem of platform B. The rest is medium.
-
-Measurement period = 29 days
-Total number of machines: 12583
-0.5 and 0.5,0.25 = 10595 = 84.2\% = heavy
-rest = 1988 = 15.8\% = medium
-
-\myparagraph{Cluster statistics}
-\begin{center}
-    \begin{tabular}{| l | l | l |}
-    \hline
-    Indicator & Scheme 1 & Scheme 2 \\
-    \hline
-    Total cost & \$8 177 048 & \$5 637 418 \\
-    \hline
-    KDI & +3 215 231 & +5 431 137 \\
-    \hline
-    Total CPU & 15972217.38 & 15972217.38 \\
-    \hline
-    Total Memory & 30676363.36 & 30676363.36 \\
-    \hline
-    Number of tasks & 1234716525 & 1234716525 \\
-    \hline
-    \end{tabular}
-\end{center}
 

### 2015-04-22 paper/paper.tex
updates

@@
 
-\myparagraph{Visual recognition}
-
-Visual recognition as a helper method to define thresholds, especially for further incremental optimizations. 
-Useful as static thresholds can miss a whole set of machines that are just below the threshold.
-
-\begin{figure}[H]
-       \includegraphics[width=\linewidth]{figures/cubism}
-      \caption{Visualization of distance from optimality points. Green - positive, Blue - negative values.}
-        \label{fig:cubism}
-\end{figure}
-
 \myparagraph{Aggregated CPU and memory usage}
@@
 Default driver and executor settings - 256M and 512M respectively - were not enough to handle the volume of input data.
-Further OOMs occurred during \textit{RRD.reduceByKey}\footnote{\url{https://spark.apache.org/docs/1.2.1/api/python/pyspark.html#pyspark.RDD.reduceByKey}} operations as it required to compute a fairly large hash map.
+Further OOMs occurred during \textit{RRD.reduceByKey}\footnote{\url{https://spark.apache.org/docs/1.2.1/api/python/pyspark.html\#pyspark.RDD.reduceByKey}} operations as it required to compute a fairly large hash map.
 Tweaks of \textit{spark.default.parallelism} setting worked the best as it enforced more granular input data chunking.
@@
 For the rest of the cluster, we will assign medium pricing scheme.
+As a result of the assignment cluster machines are grouped into 10595 machines of heavy and 2258 of medium usage.
 
@@
 
-Transformation to below schema. 
-- total number of tasks per day
-
 \myparagraph{Total CPU and memory usage}
@@
 
-Key distance indicator got calculated for both schemes based on Algorithm~\ref{alg:distance_from_optimality}.
+Key distance indicator was calculated based on Algorithm~\ref{alg:distance_from_optimality}. 
+Total cost was calculated based on total uptime and pricing scheme assignment. Results are presented on Table~\ref{tab:kdi_and_cost}.
 
@@
     \hline
-    Indicator & Scheme 1 & Scheme 2 \\
+    \textbf{Indicator} & \textbf{Scheme 1} & \textbf{Scheme 2} \\
     \hline
@@
 
-\myparagraph{Cluster SLO averages}
-
-Avg cpu usage per machine = 15972217.38/12583 = 1269.35/29 = 43.77
-Avg mem usage per machine = 30676363.36/12583 = 2437.92/29 = 84.06
-Avg tasks number per machine = 1234716525/12583 = 98125.77/29 = 3383.65
-
-\section{Recognize inefficiencies with Threshold-based Report}
-
-The report will filter all machines that reported non-optimal - negative distance from intersection point - configuration during measurement period (29 days).
-
-\myparagraph{Scheme 1}
-Number of machines to tune: 932
-CPU capacity to accommodate: 1 320 781.36
-Memory capacity to accommodate: 4 286 742.06
-Number of tasks to execute: 118869614
-Cost of running non-optimal machines: \$657 200
-Total uptime: 648519
-cost per hour = \$1.01338588383687
-Cost per task =  \$657 200 / 118869614 = \$0.00552874681666
-
-\myparagraph{Scheme 2}
-Number of machines to tune: 937
-CPU capacity to accommodate: 1 327 202.18
-Memory capacity to accommodate: 4 303 983.31
-Number of tasks to execute: 119367988
-Cost of running non-optimal machines: \$532 840.76
-Total uptime: 651899
+KDI for Scheme 1 and 2 is a positive number, which is expected based on 82\% of heavy scheme assignment and 23.36 hours of uptime per day on average.
+Scheme 1 is 31\% more expensive than Scheme 2 and this is expected as well as \textit{cph} values are higher for Scheme 1.
+
+\section{Recognize inefficiencies using threshold-based report}
+
+KDI can be used to find inefficiencies.  Any value below 0 indicates non-optimal usage for a given scheme. 
+A user can specify a threshold to find non-optimal cases. We will assume that non-optimal usage within the cluster is when machine reports negative KDI values for the whole trace period.
+ These criteria identified 932 and 937 machines for Scheme 1 and 2 respectively. Detailed results splitted by cost and resource utilization are presented in Table~\ref{tab:threshold_based_report}.
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    \textbf{Property} & \textbf{Scheme 1} & \textbf{Scheme 2} \\
+    \hline
+    Number of machines to tune & 932 & 937 \\
+    \hline
+    Total cost & \$657 200 & \$532 840.76 \\
+    \hline
+    CPU capacity to accommodate & 1 320 781.36 & 1 327 202.18 \\
+    \hline
+    Memory capacity to accommodate & 4 286 742.06 & 4 303 983.31 \\
+    \hline
+    Number of tasks to execute & 118869614 & 119367988 \\
+    \hline
+    Total uptime & 648519 & 651899 \\
+    \hline
+    Average cost per hour & \$1.0134 & \$0.817 \\
+    \hline
+    Average cost per task & \$0.00553  & \$0.00446 \\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Results of report run on cluster with criteria of KDI below 0 for 29 days}
+\label{tab:threshold_based_report}
+\end{table}
+
+The cost of running such identified non-optimal machines is around 8\% and 10\% of total cost for Scheme 1 and 2 respectively.
 
 \section{Squeeze-them-in suggestion strategy}
+
+Once inefficiencies are identified, we will attempt to find a more optimal solution.
 
@@
 
-vm1 = Entity(c_name, c_light, c_medium, c_heavy)
-vm2 = Entity(m_name, m_light, m_medium, m_heavy)
-
->>> calc_options(uptime, vm2, 29)
+$vm1 = Entity(c_name, c_light, c_medium, c_heavy)$
+$vm2 = Entity(m_name, m_light, m_medium, m_heavy)$
+
+$calc_options(uptime, vm2, 29)$
 (1597, 3727, 972, 567 273.49, 998 186.23, 410 041.68)
->>> calc_options(uptime, vm1, 29)
+$calc_options(uptime, vm1, 29)$
 (1397, 1720, 972, 770 417.58, 876 376.34, 604 274.61)

### 2015-04-23 paper/paper.tex
updates

@@
 
-\subsubsection{New capacity configuration suggestion}
+\subsubsection{New Capacity configuration suggestion}
 
@@
 \item suggestion of changes based on Algorithm~\ref{alg:squeeze-them-in} and~\ref{alg:new_configuration},
-\item validation of positive correlation between \textit{KDI} and \textit{total cost} changes.
+\item validation of a positive correlation between \textit{KDI} and \textit{total cost} changes.
 \end{enumerate}
@@
 Apache Spark 1.2.1\footnote{\url{https://spark.apache.org}} - an engine for large-scale data processing.
-All of map reduce jobs were written using Python API.
+All of the map-reduce jobs were written using Python API.
 Jobs were executed using standalone cluster mode with settings specified in Table~\ref{tab:spark_conf}.
-Job run times varied widely based on input data size and complexity of computations - from 20hs+ for transformations on 100G+ data to 1 minute aggregations on 100-200M input.
+Job run times varied widely based on input data size and complexity of computations - from 20hs+ for transformations on 100G+ data to 1-minute aggregations on 100-200M input.
 
@@
 
-For evaluation purposes we choose two pricing schemes for two different VM configurations~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}.
+For evaluation purposes, we choose two pricing schemes for two different VM configurations~\cite{AWS:light}~\cite{AWS:medium}~\cite{AWS:heavy}.
 Pricing details are specified in Table~\ref{tab:aws_pricing}
@@
 
-Google released 29-day cluster usage traces of more than 12k machines running over 650k jobs~\cite{clusterdata:Reiss2011}.
+Google released 29-day cluster usage traces of more than 12k machines that run over 650k jobs~\cite{clusterdata:Reiss2011}.
 An analysis~\cite{clusterdata:Reiss2012b} revealed that the workload is heterogeneous, which is somewhat unique compared to other traces and makes a good fit for testing.
@@
 This paper focuses on uptime and we extracted only relevant data for further evaluation.
-Machine and task events original schema is presented on Tables~\ref{tab:machine_events} and~\ref{tab:task_events} and fields selected for further transformation and aggregation are marked in bold text.
+Machine and task events original schema is presented in Tables~\ref{tab:machine_events} and~\ref{tab:task_events} and fields selected for further transformation and aggregation are marked in bold text.
 
@@
 \myparagraph{Uptime}
-Uptime per day per machine was calculated based on aggregation of task duration times for a given day.
-Algorithm~\ref{alg:uptime_per_machine} shows design and Listing~\ref{uptime_per_machine_implementation} represents implementation of this aggregation.
-Total number of hours in cluster is 8705743, which gives 23.36 hours daily per machine on average.
+Uptime per day per machine was calculated based on an aggregation of task duration times for a given day.
+Algorithm~\ref{alg:uptime_per_machine} shows design and Listing~\ref{uptime_per_machine_implementation} represents the implementation of this aggregation.
+The total number of hours in the cluster is 8705743, which gives 23.36 hours daily per machine on average.
 
@@
 
-Based on Formula~\ref{eq:cph} implemented in Python and presented on Listing~\ref{cost_per_hour}, we calculated \textit{cost per hour} for both analyzed schemes.
-Results are presented on Table~\ref{tab:cph:scheme1} and~\ref{tab:cph:scheme2} in respect of number of hours per day a machine would run.
+Based on Formula~\ref{eq:cph} implemented in Python and presented in Listing~\ref{cost_per_hour}, we calculated a \textit{cost per hour} for both analyzed schemes.
+Results are presented in Table~\ref{tab:cph:scheme1} and~\ref{tab:cph:scheme2} in respect of number of hours per day a machine would run.
 The cost per hour drops when uptime per day rises.
@@
 
-Based on Formula~\ref{alg:intersection points} implemented in Python and presented on Listing~\ref{intersection_points} we calculated \textit{intersection points} between pricing schemes.
+Based on Formula~\ref{alg:intersection points} implemented in Python and presented in Listing~\ref{intersection_points} we calculated \textit{intersection points} between pricing schemes.
 Table~\ref{tab:intersection_points} shows results of the calculation.
-Intuitively, heavy schemes is the cheapest option for highly utilized machines.
-An interesting finding is that medium is never cheaper than light in case of Scheme 2.
+Intuitively, a heavy scheme is the cheapest option for highly utilized machines.
+An interesting finding is that a medium is never cheaper than light in case of Scheme 2.
 Figures~\ref{fig:cc2_8xlarge} and~\ref{fig:m2_4xlarge} represent graphically \textit{cph} results and \textit{intersection points} can be noticed.
@@
 
-Key distance indicator was calculated based on Algorithm~\ref{alg:distance_from_optimality}. 
-Total cost was calculated based on total uptime and pricing scheme assignment. Results are presented on Table~\ref{tab:kdi_and_cost}.
+The key distance indicator was calculated based on Algorithm~\ref{alg:distance_from_optimality}. 
+Total cost was calculated based on total uptime and pricing scheme assignment. Results are presented in Table~\ref{tab:kdi_and_cost}.
 
@@
     \hline
-    Total cost & \$657 200 & \$532 840.76 \\
+    Total cost & \$657 200 & \$532 841 \\
     \hline
@@
 
-The cost of running such identified non-optimal machines is around 8\% and 10\% of total cost for Scheme 1 and 2 respectively.
+The cost of running identified non-optimal machines is around 8\% and 10\% of total cost for Scheme 1 and 2 respectively.
 
@@
 
-Once inefficiencies are identified, we will attempt to find a more optimal solution.
-
-\myparagraph{Scheme 1}
-total spare cpu = 3 697 171.44
-total spare mem = 6 685 407.83
-new KDI: +3 500 037
-old KDI: +3 215 231
-KDI change: +8.14\%
-total uptime = 8056601
-new total cost = \$8 177 048 - \$657 200 = \$7 519 848
-cost change: -8.04\%
-cost per hour = \$7 519 848 / 8056601 = 0.93337723935938
-new cost per task = \$7 519 848 / 118869614 = 0.06326131419927
-
-new averages 
-cpu per machine = 15972217.38/11651 = 1370.88 / 29 = 47.27 +7.4\%
-Avg mem usage per machine = 30676363.36/11651 = 2632.94/29 = 90.79 +7.41\%
-Avg tasks number per machine = 1234716525/11651 = 105975.15/29 = +7.4\%
-
-\myparagraph{Scheme 2}
-total spare cpu = 3 694 240.56 
-total spare mem = 6 682 956.59
-new KDI: +5 913 933
-old KDI: +5 431 137
-KDI change: +8.16\%
-total uptime = 8053221
-new total cost = \$5 637 418 - \$532 840.76 = \$5 104 577.24
-cost change: -9.45\%
-cost per hour = \$5 104 577.24 / 8053221 = 0.63385535303204
-new cost per task = \$5 104 577.24 / 119367988 = 0.04276336834964
-
-new averages 
-cpu per machine = 15972217.38/11646 = 1371.48 / 29 = 47.29 +7.44\%
-Avg mem usage per machine = 30676363.36/11646 = 2634.07 / 29 = 90.83 +7.45\%
-Avg tasks number per machine = 1234716525/11646 = 106020.65 / 29 = 3655.88 +7.45\%
+Based on Algorithm~\ref{alg:squeeze-them-in} we calculated spare capacity for both schemes. Results are presented in Table~\ref{tab:squeeze-them-in-capacity}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    \textbf{Property} & \textbf{Scheme 1} & \textbf{Scheme 2} \\
+    \hline
+    Total spare CPU & 3 697 171.44 & 6 3 694 240.56  \\
+    \hline
+    Total spare memory & 6 685 407.83 & 6 682 956.59 \\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Results of spare capacity calculation on the cluster}
+\label{tab:squeeze-them-in-capacity}
+\end{table}
+
+In both cases, there is more spare capacity than results from Table~\ref{tab:threshold_based_report} require.
+It indicates we can power off all of the non-optimal machines and investigate an impact on workload and pricing indicators.
+Results of such an experiment are presented in Table~\ref{tab:squeeze-them-in}.
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    \textbf{Property} & \textbf{Scheme 1} & \textbf{Scheme 2} \\
+    \hline
+    New KDI & +3 500 037 & +5 913 933 \\
+    \hline
+    KDI change & +8.14\% & +8.16\% \\
+    \hline
+    New total cost & \$7 519 848 & \$5 104 577.24 \\
+    \hline
+    Total cost change & -8.04\% & -9.45\% \\
+    \hline
+    New cost per hour & \$0.9334 & \$0.6339 \\
+    \hline
+    New cost per task & \$0.0633  & \$0.0428 \\
+    \hline
+    New CPU average per machine & 47.27 & 47.29 \\
+    \hline
+    CPU average change & +7.4\% & +7.44\% \\
+    \hline
+    New memory usage average & 90.79 & 90.83 \\
+    \hline
+    Memory average change & +7.41\% & +7.45\% \\
+    \hline
+    New task number average & 3654 & 3655 \\
+    \hline
+    Task average change & +7.4\% & +7.45\% \\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{Cluster statistics with non-optimal machines powered off}
+\label{tab:squeeze-them-in}
+\end{table}
+
+The KDI increased by above 8\% for both schemes.
+This is expected as we removed machines with negative KDI values.
+The total cost decreased by 8.04\% and 9.45\% for Scheme 1 and 2 respectively.
+This is also expected.
+We noticed a correlation between the increase of KDI and decrease of total cost.
+However, there is a trade-off in terms of resource utilization, which increased around 7.4\% on average.
+A user has to decide whether this is acceptable for his specific workload SLOs.
+It might work well for under-utilized clusters with a low peak-to-mean ratio.
+For high peak-to-mean workloads, the real difference in utilization patterns can be higher than detected by averages.
+To measure SLOs in these environments, we would have to use more specific metrics like requests per second or queries per second during peak.
 
@@
 
-$vm1 = Entity(c_name, c_light, c_medium, c_heavy)$
-$vm2 = Entity(m_name, m_light, m_medium, m_heavy)$
-
-$calc_options(uptime, vm2, 29)$
-(1597, 3727, 972, 567 273.49, 998 186.23, 410 041.68)
-$calc_options(uptime, vm1, 29)$
-(1397, 1720, 972, 770 417.58, 876 376.34, 604 274.61)
-
-\section{Validate influence of suggestions and correlation with KDI}
+We used Algorithm~\ref{alg:new_configuration} to come up with new configuration suggestions based on the uptime of machines from Table~\ref{tab:threshold_based_report} and Scheme 1 and 2 pricing details.
+\\
+The algorithm analyzed all three \textit{intersection points} available for each scheme and returned total cost and number of machines required.
+Results are presented in Table~\ref{tab:new-cap-scheme1} and~\ref{tab:new-cap-scheme2}.
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    \textbf{Scheme} & \textbf{Machines} & \textbf{Total cost} \\
+    \hline
+    Light & 1397 & \$770 417.58 \\
+    \hline
+    Medium & 1720 & \$876 376.34 \\
+    \hline
+    Heavy & 972 & \$604 274.61\\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{New capacity configuration suggestions for Scheme 1}
+\label{tab:new-cap-scheme1}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+    \begin{tabular}{| l | l | l |}
+    \hline
+    \textbf{Scheme} & \textbf{Machines} & \textbf{Total cost} \\
+    \hline
+    Light & 1597 & \$567 273.49 \\
+    \hline
+    Medium & 3727 & \$998 186.23 \\
+    \hline
+    Heavy & 972 & \$410 041.68 \\
+    \hline
+    \end{tabular}
+\end{center}
+\caption{New capacity configuration suggestions for Scheme 2}
+\label{tab:new-cap-scheme2}
+\end{table}
+
+An interesting pattern raised - a medium is always the least optimal choice and heavy is the best.
+The suggestion to run heavy configuration for Scheme 1 would save 8.05\% compared to non-optimal machines set and 0.0065\% from total cost.
+Scheme 2 would save 23\% compared to non-optimal machines set and 0.022\% from total cost.
+KDI increased for both schemes.
+Scheme 1 noted 8.42\% change and Scheme 2 8.31\%.
+Light and medium suggestions turned out to be more expensive and were excluded from the rest of considerations.
+\\
+
+As opposed to the previous algorithm, this one does not violate SLOs. The new suggested capacity handles the same amount of uptime using the same resource configuration as the rest of the cluster.
+
+\section{Validate correlation of total cost and KDI}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{figures/kdi_to_cost}
+	\caption{KDI to total cost correlation}
+	\label{fig:kdi_to_cost}
+\end{figure}
 

### 2015-04-25 paper/paper.tex
related work

@@
 
+Thanks to my wife Karen for her constant support and patience during the research and writing
+of this Thesis.
+Also, thanks to Prof. John Murphy and Viliam Holub for their support, help and assistance.
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%        *Some discussion to provide context
+%%        *A description of the problem
+%%        *Motivation for solving the problem
+%%        *A statement of the objectives
+%%        *Related work (mention how your work is different/better)
+%%        *Describe the proposed solution
+%%        *Mention some of the challenges
+%%        *Describe the methodology (modelling and experimentation) and key overall result (e.g., Results show proposed approach can achieve a significant reduction in the cost by enabling a more optimal selection of VM types based on resource utilisation)
+%%        *Structure of the thesis
+
 \chapter{Introduction}
 
-\emph{ /* Type of the project: The 'Big Idea' / proof by construction. */}
+Cloud computing receives significant attention recently ~\cite{5741288}~\cite{5704303}.
+
+The popularity of cloud service spurs the increasing demands of cloud resources to the cloud service providers.
+Along with the new business opportunities, the pay-as-you-go model drastically changes the usage pattern and brings technology challenges to effective capacity planning.~\cite{6274129}
+
+In the cloud context, pricing and capacity planning are two important factors to the profit of the Infrastructure-as-a-Service (IaaS) providers~\cite{6963393}
+
+. We assume that the IaaS provider offers a pay only what you use scheme similar to the Amazon EC2 service, comprising flat, on demand, and spot virtual machine instances.
+
+Cloud computing has emerged as a new IT delivery model in which an organization or individual can rent remote compute and storage resources dynamically, using a credit card, to host networked applications “in the cloud.”.
+The appeal of cloud computing lies in its usage-based pricing model - organizations only pay for the resources that are actually used, and can flexibly increase or decrease the resource capacity allocated to them at any time.
+This elasticity provided by Cloud computing can yield significant cost savings when compared to the traditional approach of maintaining an expensive IT infrastructure that is provisioned for peak usage - organizations can instead simply rent capacity, and grow and shrink it as the workload changes.~\cite{5961733}
 
@@
 \section{Literature review}
-\emph{/* Haven't found anything that would tackle this specific problem so far.
-Might mention the most research is done from IaaS provider perspective?
-This includes resource scheduling.
-My work can lead to cost-aware scheduler research */}
-
-%good number of journal papers, a swathe of conference papers, a few recent workshop
-%papers and not too many web sites
+
+Work related to IaaS market is focused either on provider's perspective~\cite{6274129}, provider-consumer business interactions~\cite{6676685}~\cite{6963393} or consumer point of view~\cite{5961733}~\cite{6295066}.
+
+In this paper~\cite{6274129}, authors focused on an analysis of pay-as-you-go flexible charging impact on IaaS provider's capacity planning.
+Their contribution was focused on cloud capacity estimation and management.
+Specifically, they developed a method to predict provisioning demand based on selected time series metrics.
+Although prediction is not included in this paper, it would be a next logical step for future work to better address real-life use cases.
+
+This paper~\cite{6963393} studied IaaS consumer's optimal decisions in the monopoly and multiple IaaS providers market.
+Based on rational customer decisions, authors derived an optimal solution to provider's pricing problem.
+Furthermore, they modeled the pricing and capacity planning competition among the IaaS providers as a three-stage Stackelberg game and studied existence and conditions to achieve the Nash equilibrium.
+
+In the work conducted in paper~\cite{6676685}, authors proposed a two-stage provisioning scheme.
+At the first stage, they proposed a solution that works with flat and on-demand instances.
+Secondly, they studied bidding for the spot instances as a Stackelberg game.
+Although not focused on reserved instances, authors developed the system model that might serve as an inspiration for future work to provide suggestions based on application specific QoS metrics.
+
+Paper~\cite{5961733} described a cost-aware system that computed both a cost optimized configuration for the desired capacity as well as plan to migrate current setup to the proposed one.
+The cost-aware provisioning decisions are based on resource utilization and focused on finding the best resource configuration.
+Transition cost-aware provisioning included costs of migration and studied options to minimize it.
 
@@
 
-Investment in compute resources comes with upfront cost.
-In case of physical machine it is hardware or lease cost.
-For virtual resources, IaaS providers offer per hour discounts for upfront charged schemes.
+Investment in compute resources comes with an upfront cost.
+In case of a physical machines, it is hardware or lease cost.
+For virtual resources, a client can use IaaS provider services.
+These usually come with some fixed upfront cost and further per hour charges.
 \par
-Upfront cost is considered to be one time payment and will be represented by \textit{u}.
+
+The upfront cost is considered to be a one-time payment and will be represented by \textit{u}.
 
@@
 
-Cost per hour represents a total cost of running a single machine divided by number of hours in calendar year.
-Aggregated cost is relatively easy to calculate in rented, virtualized environments.
 IaaS providers usually charge an X amount per hour.
-It might vary based on total uptime per month or pricing scheme.
+The charge might vary based on monthly uptime utilization or assigned pricing scheme.
 This number will be considered as a partial cost per hour and defined as \textit{pcph}.
-\par
-Cost per hour will be defined as \textit{cph}. It includes upfront cost and IaaS provider per hour charges. It is defined as a function of hours of utilization per day.
+\\
+Cost per hour represents a \textit{total cost} of running a single machine divided by number of hours in a calendar year.
+It includes upfront cost and per hour charges.
+It is defined as \textit{cph} and is a function of uptime hours per day summarized by equation~\ref{eq:cph}
 
@@
 
-TO BE REWRITTEN!
-
 Cost per hour is calculated for each of available pricing schemes.
-Single result represents a matrix of hours per day and \textit{cph}.
+The single result represents a matrix of hours per day and \textit{cph}.
 Graphical representation of results produces a graph similar to Figure~\ref{fig:cc2_8xlarge}.
 \par
-This representation allows to notice pricing trends.
-In the example below, price drops exponentially for around first 5 hours of usage and nearly stabilizes afterward.
-Light is the cheapest option for around 15 hours per day usage.
-This is better visible on Figure~\ref{fig:cph_cc2_8xlarge_zoom}.
-After 15 hours, it is better to invest in the Heavy utilization scheme.
-This is the point that will be called \textit{an intersection point} between 2 pricing schemes and noted as \textit{ip(scheme1, scheme2)}.
-\par
-For the example below, cc2.8xlarge instance's intersection point between Light and Heavy schemes is \textit{ip(light, heavy) = 16}.
-
 \begin{figure}[H]
@@
 
+On the graph, there are visible intersections between pricing scheme's~\textit{cph}s.
+These points are an area of interest as they indicate when one scheme becomes more expensive than the other.
+For example, a light is the cheapest option to run for the first 16 hours a day.
+After this point, it is better to run a heavy utilization scheme.
+However, if we were to choose between medium and heavy, it would be better to run a medium for the first 13 hours.
+And between a medium and light, it makes sense to utilize light scheme for the first 18 hours and then switch to medium.
+
 \section{Algorithm}
@@
 The below list of metrics was compiled based on algorithm needs to answer critical questions regarding the state of infrastructure deployment.
-The list is suitable to guarantee performance SLOs only for compute nodes.
+The list is suitable to guarantee performance SLAs for compute nodes.
 Machines with significant I/O usage profile are not covered in this paper.
@@
 
+The following list of metrics is required.
 \myparagraph{Machine Id}
-A unique machine ID. Will be used as a key to compute further analytics.
+A unique machine ID. Used as a key to compute further analytics.
 
@@
 
-Metrics should be aggregated in \textless K, List\textless V\textgreater\textgreater format where K represents Machine Id and Vs are measurements.
+Metrics should be aggregated in \textless K, List\textless V\textgreater\textgreater format where K represents Machine Id and List of V is a list of measurements in a given period.
 
@@
 Collection per job or per process produces one log line and has a scope of a measurement period.
-Such structured data itself would not answer machine wide neither cluster-wide questions.
-The data has to be post-processed to allow calculation of below listed aggregations and analytics.
+Such structured data would not help to answer machine-wide neither cluster-wide questions.
+The data has to be post-processed to allow calculation of below-listed aggregations and analytics.
 
@@
 
-The data will be analyzed in terms of daily usage. This requires to split metrics into buckets of one day worth of data. 
-It is achieved by applying a simple filter from Listing~\ref{daily_usage} in Apache Spark.
+To comply with IaaS pricing strategies, the data is analyzed in terms of a daily usage. This requires splitting metrics into buckets of one day worth of data. 
+It is achieved by applying a simple Apache Spark filter presented in Listing~\ref{daily_usage} that calculates day number based on metric timestamps.
 
@@
 
-The aim for this metric is to find out how long a machine is up.
-It is a key metric that will be used to calculate optimality factors later on.
+This metric indicates a daily number of hours of machine being up.
+It is a key metric that is be used to calculate optimality indicators discussed later on.
 Algorithm~\ref{alg:uptime_per_machine} represents pseudo-code implementation and Listing~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
+The data is extracted from timestamp values previously grouped into daily data buckets.
 
@@
 Algorithm~\ref{alg:cost_per_hour} represents pseudo-code and Listing~\ref{cost_per_hour} shows implementation in Python.
+The algorithm returns a mapping of number of uptime hours per day and corresponding~\textit{cph} values.
 
@@
 
-This is an implementation of \textit{ip(scheme1, scheme2)} defined in Taxonomy.
-This metric will be a reference point that identifies optimality points based on~\textit{cph} for various schemes.
+This is an implementation of \textit{ip(scheme1, scheme2)} metric defined in Taxonomy.
+This represents a reference point that finds intersections of~\textit{cph}s for various schemes.
 Algorithm~\ref{alg:intersection points} represents pseudo-code and Listing~\ref{intersection_points} shows implementation in Python.
@@
 
-\myparagraph{Uptime Based Distance From Intersection Points}
+\myparagraph{Key Distance Indicator}
 
 This metric represents a distance from intersection points calculated based on uptime.
-The result can be negative - indicates non-optimal profile - or positive otherwise.
+The result can be negative - to indicate a non-optimal profile - or positive otherwise.
 Positive and negative values can be yielded in either direction of the time axis as presented in Figure~\ref{fig:distance}.
-It will be specified by flag parameter calculated together with \textit{ip(scheme1, scheme2)} results. Algorithm~\ref{alg:distance_from_optimality} represents pseudo-code implementation.
 
@@
 \end{figure}
+
+At 18 hours, the point of scheme cost intersection (highlighted as 1), it does not matter which scheme is chosen as cost is the same. This is point 0.
+
+At each side of the intersection point either scheme1 or scheme 2 is more preferable.
+
+For example, a machine running for 20 hours on scheme1 is running 2 hours outside our optimum 18 hour scheme cost intersection.
+This gives us a score of -2 for this machine compared to the same machine running on scheme 2 where the benefits continue to increase the longer the machine is running.
+
+The same is true for a machine running for 10 hours on scheme 2.
+We are running at -8 below our optimum.
+In this case we would be more cost beneficial to run this machine on scheme1.
+Giving us a score of +8.
+
+Algorithm~\ref{alg:distance_from_optimality} represents pseudo-code implementation.
 
@@
 
-Example algorithm results
-\begin{enumerate}
-\item Input: Uptime: 24, Optimality Point: 18, Bias: True (more optimal for values above threshold). Output: +6
-\item Input: Uptime: 16, Optimality Point: 18, Bias: True (more optimal for values above threshold). Output: -2
-\item Input: Uptime: 15, Optimality Point: 18, Bias: False (more optimal for values below threshold). Output: +3
-\item Input: Uptime: 19, Optimality Point: 18, Bias: False (more optimal for values below threshold). Output: -1
-\end{enumerate}
-
-\subsection{Recognize inefficiencies and suggest changes}
+\subsection{Recognition of inefficiencies and suggestions for changes}
 
@@
 
-Generate report of machines based on distance from optimality points. Use customizable threshold. 
-Results from the report will be used to calculate suggestions.
+A user can define which KDI value is non-optimal and run a report to find machines below this point.
+For example, every negative KDI value signalizes there is a better pricing scheme available.
+A user can catch this situation by having a threshold value set to 0.
+However, there might be a situation when machine is taken for maintenance and reports negative values for one or couple of days.
+To avoid this kind of noise, a user can add an another criteria to specify a minimum number of days of reported values below the threshold.
 
@@
 
-Aggregated values of CPU and memory will help to suggest better utilization strategies that guarantee SLOs.
+An aggregated daily values of CPU and memory will help to suggest better utilization strategies that guarantee SLAs.
 Algorithm~\ref{alg:agg_cpu_mem} represents pseudo-code implementation and~\ref{agg_cpu_mem_implementation} contains implementation in Python with Apache Spark.
@@
 The proposed strategy calculates an amount of spare CPU and memory capacity based on cluster-wide averages.
-Aggregated value for CPU and memory represents an amount of capacity gathered from machines that are utilized below average.
-Based on the amount of aggregated capacity, the algorithm will attempt to squeeze in a maximum amount of machines gathered from Threshold Based Report.
-As a result, it will propose to turn off a number of machines and guarantee that SLOs will be met.
-This approach performs well when underutilized machines exist.
-For well-balanced clusters with similar usage patterns, it does not find a lot of spare capacities and would propose fairly limited improvements.
+Aggregated value for CPU and memory represents an amount of capacity gathered from machines that are utilized below the average.
+Based on the amount of aggregated capacity, the algorithm will attempt to squeeze in a maximum amount of machines returned by threshold-based report.
+As a result, it proposes to turn off a number of machines and that can fit in calculated below-average capacity.
+\\
+This approach has a potential to perform well in under-utilized clusters.
+For well-balanced environments with similar usage patterns, it does not find a lot of spare capacity and would propose fairly limited improvements.
+\\
+A pseudo-code implementation of this idea is presented in Algorith~\ref{alg:squeeze-them-in}.
 
@@
 
-\subsubsection{New Capacity configuration suggestion}
+\subsubsection{New capacity configuration suggestion}
+
+New configuration is calculated based on input from the report, list of available schemes and total uptime to handle.
+Pseudo-code implementation is presented in Algorithm~\ref{alg:new_configuration}.
+The algorithm iterates through all of available schemes, reads \textit{intersection points} and gets \textit{cph} for each of the \textit{ip}s.
+It returns a total cost and number of machines required for each proposed configuration.
+\\
+Suggestion results are limited to one type of pricing scheme. In reality, there might be a more optimal solution yielded from mixing pricing schemes.
 
@@
 \label{alg:new_configuration}
- \KwData{List of machines from Threshold Based Report, List of available schemes, Average number tasks per hour;}
+ \KwData{List of machines from Threshold Based Report, List of available schemes, Total uptime;}
  \KwResult{New configuration suggestion}
@@
   $float[ ] totalCost;$
-  $float totalCpu = 0;$
-  $float totalMemory = 0;$
-
-\ForAll{machine in machinesFromReport}{
-        machine.cpu += totalCpu;
-        machine.memory += totalMemory;
+
+\ForAll{scheme in availableSchemes}{
+		\ForAll{ip in scheme.intersectionPoints}{
+			vms = Uptime / ip.getHours() \\
+			cost = ip.getCphForHours(ip.getHours()) \\
+			totalCost[ip] = (vms, cost)
 		}
-
-\ForAll{scheme in availableSchemes}{
-		hoursPerDay = scheme.getHours();
-        cpuCap = hoursPerDay * clusterCpuAvgPerHour;
-        memCap = hoursPerDay * memoryCpuAvgPerHour;
-
-        vmsReqCpu = totalCpu / cpuCap;
-        vmsReqMem = totalMemory / memCap;
-
-        totalCost[scheme] = min(vmsReqCpu, vmsReqMem) * scheme.cph;
 }
 
-\Return{min(totalCost)}
-
+\Return{totalCost}
 \end{algorithm}
@@
 
+We aggregated all of the generated suggestions and graphed values of KDI and total cost changes on Figure~\ref{fig:kdi_to_cost}. 
+By design, KDI reports higher positive number when usage is more optimal.
+We can see that for each suggestion generated by both algorithms, KDI increased and the total cost dropped.
+This suggests a direct correlation between these values.
+However, higher KDI increase did not directly decreases the total cost further.
+
 \begin{figure}[H]
@@
 
-\section{Summary}
-
 %%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{enumerate}
-\item Visual recognition of underperforming machines
-\item Prediction trend to generate suggestions
+\item Visual recognition of underperforming machines using horizon charts (http://vis.berkeley.edu/papers/horizon/)
+\item Create an optimality model and adjust new capacity suggestions to find the best combination
+\item Predict growth trend and adjust configuration suggestions
+
+%%
+%%Also, just to reiterate a couple of things I mentioned
+%%before: while a realistic algorithm would have to
+%%account for predicted prices, predicted discounts,
+%%etc I think you should initially focus on a scenario
+%%where the pricing scheme is fixed (though the prices
+%%themselves will vary in time, naturally). And it's very
+%%important that you produce 'something' working e.g.
+%%an implementation of your algorithm, tied into the
+%%data and so on.
+
+%%
+
 \end{enumerate}

### 2015-04-25 paper/paper.tex
updates

@@
 \section{The Problem Domain}
-\emph{/* Description of IaaS world - providers and consumers. */}
-\subsection{IaaS consumer point of view}
-\subsection{IaaS provider uptime-based pricing schemes}
-
-\section{Motivation}
-\emph{ /* Description of the problem - cloud-based company runs a mix of software services (real-time streaming, batch, web and database servers) on common hardware.
-How do they know whether provisioned VMs run in the most optimal configuration?
-*/ }
+
+In this thesis, we look from the perspective of Software as a Service (SaaS) provider that uses an Infrastructure as a Service (IaaS) provider's services to host a product. \\
+SaaS provider seeks i) predictable cost estimates and ii) the best value for money.
+The latter for obvious reasons - every healthy business is interested in minimizing costs.
+The former is needed to set a right product pricing.
+Not too expensive and not too cheap.
+Failure to do so may ruin a business's budget.
+\\
+IaaS provider pricing is flexible as described in below sections.
+One instance configuration can have many prices based on daily uptime.
+Cost-aware uptime-based provisioning becomes a new consideration and adds one more parameter to standard resource-aware provisioning techniques~\cite{Bartolini:2014:AFC:2658949.2637480}.
+
+\section{Amazon EC2 pricing schemes}
+
+A price of an instance in Amazon EC2 depends on its resources and pricing scheme it belongs to.
+As a result of the latter, one resource configuration can have few different prices.
+Amazon offers on-demand, spot and reserved instances billed hourly.
+
+\paragraph{On-demand instances.} Allow to pay by the hour with no other commitments. The most expensive option.
+\paragraph{Spot instances.} A real-time bidding platform. A customer can bid on price. Instance can run only when the bid exceeds current Spot Price.
+\paragraph{Reserved instances.} Discounted price in comparison to On-demand in exchange of 1 or 3 years commitment. Guaranteed availability of reserved capacity.
+
+This paper focuses around the concept of Reserved Instances.
+The scheme is further split into Light, Medium and Heavy options.
+They differ in upfront and hourly costs.
+Based on pricing data, Light serves well for short periods of daily usage, Heavy is optimized for long workloads and Medium is in between.
+\\
+This scheme got simplified on December 2014~\cite{AWS:pricing_change} and naming used in this paper does not match the latest offering anymore.
+However, the idea of having the same instance configuration priced differently based on uptime is still valid and is used by other IaaS providers, e.g.
+Google Compute Engine.
+
+\section{Google Compute Engine pricing}
+
+Google had greatly simplified the Compute Engine pricing.
+There is no concept of reservation nor bidding.
+All of the options are priced within pay-as-you-go model with 1 minute granularity.
+However, Google offers increasing discounts for sustained usage - 20\% for 25-50\% of month usage, 50-75\% of usage yields 40\% and everything above 75\% is priced with 60\% of discount rate.
+This is similar to the concept of Reserved Instances in a sense that uptime value influences a final price for the instance.
 
@@
 
-Work related to IaaS market is focused either on provider's perspective~\cite{6274129}, provider-consumer business interactions~\cite{6676685}~\cite{6963393} or consumer point of view~\cite{5961733}~\cite{6295066}.
+Work related to IaaS market is focused either on provider's perspective~\cite{6274129}~\cite{6253563}, provider-consumer business interactions~\cite{6676685}~\cite{6963393} or consumer point of view~\cite{5961733}~\cite{6295066}.
+
+Paper~\cite{6253563} proposed a pricing framework for cloud services using game theory and data mining techniques.
+The framework determines the price based on recent usage data and available resources.
+It also takes into consideration various economic models and attempts to find the best price for provider and consumer at the same time.
 
@@
 
-Evaluation process can be splitted into four phases:
+Evaluation process can be split into four phases:
 \begin{enumerate}
@@
 Customer could choose an instance configuration and total cost would be calculated based on a number of hours of uptime per day.
-The offering was splitted into three options:
+The offering was split into three options:
 \begin{enumerate}
@@
 CPU, memory - and timestamps.
-The trace does not contain information about the purpose of the jobs, which limits SLO assumptions to resource utilization.
+The trace does not contain information about the purpose of the jobs, which limits SLA assumptions to resource utilization.
 Resource utilization data is provided in normalized units - the strongest configuration is 1 and the rest of configurations are described in relation to that value.
@@
 A user can specify a threshold to find non-optimal cases. We will assume that non-optimal usage within the cluster is when machine reports negative KDI values for the whole trace period.
- These criteria identified 932 and 937 machines for Scheme 1 and 2 respectively. Detailed results splitted by cost and resource utilization are presented in Table~\ref{tab:threshold_based_report}.
+ These criteria identified 932 and 937 machines for Scheme 1 and 2 respectively. Detailed results split by cost and resource utilization are presented in Table~\ref{tab:threshold_based_report}.
 
@@
 However, there is a trade-off in terms of resource utilization, which increased around 7.4\% on average.
-A user has to decide whether this is acceptable for his specific workload SLOs.
+A user has to decide whether this is acceptable for his specific workload SLAs.
 It might work well for under-utilized clusters with a low peak-to-mean ratio.
 For high peak-to-mean workloads, the real difference in utilization patterns can be higher than detected by averages.
-To measure SLOs in these environments, we would have to use more specific metrics like requests per second or queries per second during peak.
+To measure SLAs in these environments, we would have to use more specific metrics like requests per second or queries per second during peak.
 
@@
 
-As opposed to the previous algorithm, this one does not violate SLOs. The new suggested capacity handles the same amount of uptime using the same resource configuration as the rest of the cluster.
+As opposed to the previous algorithm, this one does not violate SLAs. The new suggested capacity handles the same amount of uptime using the same resource configuration as the rest of the cluster.
 

### 2015-04-25 paper/paper.tex
update

@@
 
+\chapter{Introduction}
+
 %%        *Some discussion to provide context
-%%        *A description of the problem
-%%        *Motivation for solving the problem
-%%        *A statement of the objectives
-%%        *Related work (mention how your work is different/better)
-%%        *Describe the proposed solution
-%%        *Mention some of the challenges
-%%        *Describe the methodology (modelling and experimentation) and key overall result (e.g., Results show proposed approach can achieve a significant reduction in the cost by enabling a more optimal selection of VM types based on resource utilisation)
-%%        *Structure of the thesis
-
-\chapter{Introduction}
-
-Cloud computing receives significant attention recently ~\cite{5741288}~\cite{5704303}.
+%%  Cloud computing stal sie popularny. Firmy zaczynaja budowac cale modele biznesowe w oparciu o Cloud, 
+%% 	Wczuwamy sie w  SaaS. Uzywamy IaaS zeby hostowac na tym produkt.
+
+In recent years, Cloud Computing receives significant attention~\cite{6185529}~\cite{5741288}~\cite{5704303} and delivers infrastructure, platform and/or software as services over the Internet~\cite{7073248}.
+
+The popularity of cloud service pushes cloud service providers to provide
 
@@
 Along with the new business opportunities, the pay-as-you-go model drastically changes the usage pattern and brings technology challenges to effective capacity planning.~\cite{6274129}
+
+%%        *A description of the problem
+%% Znamy wymagania sprzetowe. Stajemy przed pytaniami: 1) jak duzo VMow 2) jakiego rodzaju 3) na jak dlugo ma kazda dzialac
+
+%%        *Motivation for solving the problem
+%% rozwiazanie problemu ma 2 glowne aspekty: 1) optymalniejsze ustawienie istniejacej architektury - oszczednosci (jaki procent?), ktore mozna zainwestowac gdzies indziej 2) to decyduje o byc lub nie byc dla oplacalnosci biznesu, ktory dziala na marginesach
+
+%%        *A statement of the objectives
+%% 1) chcemy opracowac liste niezbednych metryk, ktore umozliwiaja podejmowanie decyzji na podstawie danych 2) chcemy zaprojektowac wspolczynniki, ktore wyznaczaja i) rzeczywista cene uzywania pricing scheme ii) momenty stawania sie jednej scheme bardziej oplacalna od drugiej iii) wspolczynnika okreslajacego odleglosc od optymalnych punktow 3) chcemy zaprojektowac algorytmy, ktore zaproponuja zmiany w architekturze ze wgledu na uptime i cene i podadza nowa cene
+
+%%        *Related work (mention how your work is different/better)
+%% prace czesto operieraja sie na IaaS albo relacji pomiedzy Saas i IaaS. Te, ktore sa z perspektywy SaaS nie uwzgledniaja tego, ze uptime wplywa na pricing.
+
+%%        *Describe the proposed solution
+%% 1) wybralismy pewne metryki 2) policzylismy agregacje metryk w stosunku per maszyna dziennie 3) obliczylismy wspolczynniki - cph, ip(scheme1, scheme2), KDI 5) zaprojektowalismy 2 algorytmy do sugestii - krotki opis 5) uzylismy wspolczynnikow i metryk jako danych wejsciowych to algorytmow sugestii zmian
+
+%%        *Mention some of the challenges
+%% 1) zmieniajace sie ceny i zasady u providerow - coraz taniej (dowod?) 2) dostosowanie sie do sugestii wymaga migracji z istniejacego srodowiska do nowego - koszt tego nie zostal wliczony, a powinien 3) obliczenia agregacji musza byc dobrze zaimplementowane, zeby byly mozliwe na duzych ilosciach danych
+
+%%        *Describe the methodology (modelling and experimentation) and key overall result (e.g., Results show proposed approach can achieve a significant reduction in the cost by enabling a more optimal selection of VM types based on resource utilisation)
+%% metodologia - zamodelowanie sytuacji i rozwiazania - eksperymentacja z algorytmami - 
+%%        *Structure of the thesis
 
@@
 This elasticity provided by Cloud computing can yield significant cost savings when compared to the traditional approach of maintaining an expensive IT infrastructure that is provisioned for peak usage - organizations can instead simply rent capacity, and grow and shrink it as the workload changes.~\cite{5961733}
-
-\section{Description of the problem}
-
-\section{Motivation for solving the problem}
-
-\section{Objectives}
-This work aims to design an algorithm that answer IaaS consumer's price optimality questions and help to make adjustments based on uptime based charging per machine.
-
-\section{Related work}
-
-\section{Challenges}
-
-\section{Description of methodology}
-Modelling and experimentation
 

### 2015-04-26 paper/paper.tex
conclusion

--- 2015-04-25 paper/paper.tex
@@
 
+\chapter{Introduction}
+
 %%        *Some discussion to provide context
+
+In recent years, Cloud Computing receives significant attention~\cite{6185529}~\cite{5741288}~\cite{5704303} and delivers infrastructure, platform and/or software as services over the Internet~\cite{7073248}.
+The popularity of cloud service encouraged many cloud service providers, like Amazon, Google, Microsoft, IBM and VMWare to come up with a range of provisioning and pricing schemes.
+The popular pay-as-you-go pricing model changes the usage patterns and influences the way capacity planning is done~\cite{6274129}.
+For example, one virtual machine configuration can have few different prices based on its uptime and pricing scheme.
+The dependency on uptime value is the central point of consideration in this thesis.
+We look from the perspective of Software as a Service (SaaS) provider that uses an Infrastructure as a Service (IaaS) provider's services to host a product.
+In this context, pricing and capacity planning are two important constituents of SaaS provider's success~\cite{6963393}.
+
 %%        *A description of the problem
+
+IaaS consumer can face a situation when resource-based capacity is known and it is time to rent an optimal machine configuration from the provider.
+It raises few open ended questions, including i) how many VMs is needed?
+ii) what kind of VMs and which pricing scheme?
+iii) how long should each VM run?
+In this thesis, we attempt to answer these questions in a cost-aware fashion considering uptime-based pricing models.
+
 %%        *Motivation for solving the problem
+
+Answers to above questions have influence in two main cases 1) to decide on pricing of a product, 2) to optimize an existing configuration and bring savings.
+The former is crucial for SaaS providers as they derive profits from the margin between the operational costs of infrastructure and revenue generated from customers~\cite{efficient_resource_allocation}.
+In other words, having this set right, gives a competitive advantage and greatly contributes to a healthy business.
+The latter proved to yield savings up to 9.45\% that can be invested in other aspects of a business.
+
 %%        *A statement of the objectives
+
+Motivated by a possibility to influence above factors, we propose a data-driven approach to solving this problem. Our objectives include:
+\begin{enumerate}
+\item Selection, collection and post-processing of crucial metrics that contribute to a final price and Service Level Agreements (SLAs).
+\item Design and calculation of key indicators related to: i) total cost per hour (\textit{cph}), ii) exact points when one scheme becomes more optimal than the other - called intersection points (\textit{ip(scheme1, scheme2)}), iii) distance from intersection points (\textit{KDI}) used as a key indicator to categorize a single VM usage pattern.
+\item Design and evaluation of two suggestion algorithms that take above data as an input and returns a new configuration suggestion and associated total cost.
+\end{enumerate}
+
 %%        *Related work (mention how your work is different/better)
+
+Most of the related work is conducted from an IaaS provider perspective~\cite{6274129}~\cite{6253563} or aimed to find an optimal pricing state for providers and consumers to collaborate~\cite{6676685}~\cite{6963393}.
+This thesis is different in a way that it analyses a SaaS provider capacity planning in an environment when one resource have different prices tightly coupled with an uptime.
+
 %%        *Describe the proposed solution
+
+Our proposed solution implements listed objectives using mainly map-reduce jobs executed on Apache Spark\footnote{\url{https://spark.apache.org}} platform.
+Aggregations, key indicators and algorithms were designed, implemented in pseudo-code and Python and evaluated on Google cluster-usage trace~\cite{clusterdata:Reiss2011}.
+
 %%        *Mention some of the challenges
+
+Main challenges include dealing with dynamic changes of pricing schemes offered by IaaS providers.
+Even though the thesis uses Amazon's related taxonomy, it should be read as an attempt to solve a general case for uptime-based variable pricing capacity planning.
+Other challenging aspect is a cost of migration to suggested environment.
+It was not considered in the results, however, it created an opportunity for future work to investigate an impact of the delta difference between an old and new configuration cost.
+Lastly, we hit a technical issues related to implementation and execution of aggregations and algorithms on large data sets.
+Jobs were failing due to not enough RAM capacity on a single machine.
+The issues were resolved by a combination of code improvements, like changing the way aggregations build hash-maps and an increase of parallelism and chunking granularity on the platform.
+
 %%        *Describe the methodology (modelling and experimentation) and key overall result (e.g., Results show proposed approach can achieve a significant reduction in the cost by enabling a more optimal selection of VM types based on resource utilisation)
-%%        *Structure of the thesis
-
-\chapter{Introduction}
-
-Cloud computing receives significant attention recently ~\cite{5741288}~\cite{5704303}.
-
-The popularity of cloud service spurs the increasing demands of cloud resources to the cloud service providers.
-Along with the new business opportunities, the pay-as-you-go model drastically changes the usage pattern and brings technology challenges to effective capacity planning.~\cite{6274129}
-
-In the cloud context, pricing and capacity planning are two important factors to the profit of the Infrastructure-as-a-Service (IaaS) providers~\cite{6963393}
-
-. We assume that the IaaS provider offers a pay only what you use scheme similar to the Amazon EC2 service, comprising flat, on demand, and spot virtual machine instances.
-
-Cloud computing has emerged as a new IT delivery model in which an organization or individual can rent remote compute and storage resources dynamically, using a credit card, to host networked applications “in the cloud.”.
-The appeal of cloud computing lies in its usage-based pricing model - organizations only pay for the resources that are actually used, and can flexibly increase or decrease the resource capacity allocated to them at any time.
-This elasticity provided by Cloud computing can yield significant cost savings when compared to the traditional approach of maintaining an expensive IT infrastructure that is provisioned for peak usage - organizations can instead simply rent capacity, and grow and shrink it as the workload changes.~\cite{5961733}
-
-\section{Description of the problem}
-
-\section{Motivation for solving the problem}
-
-\section{Objectives}
-This work aims to design an algorithm that answer IaaS consumer's price optimality questions and help to make adjustments based on uptime based charging per machine.
-
-\section{Related work}
-
-\section{Challenges}
-
-\section{Description of methodology}
-Modelling and experimentation
-
-\section{Structure of the Thesis}
-\textbf{Background}.
-This section will outline the background information related to IaaS consumer challenges related to understanding the whole picture for price optimality of rented infrastructure.
+
+We used modelling and experimentation methodology to build and evaluate the result.
+The SaaS provider perspective got represented by an availability of per task and per machine metrics.
+We came up with aggregations and key indicators that helped to characterise an existing state of the configuration.
+Then, we experimented with algorithms to design ones that can yield positive improvements in the total cost.
+Results show proposed approach can achieve a meaningful reduction in the cost by enabling a more optimal assignment of VMs based on uptime and resource utilization.
+
+\myparagraph{Structure of the Thesis}\\
+\textbf{Background}. This section outlines the background information related to IaaS consumer point of view and challenges related to finding the best configuration for the money. \par
+\textbf{Design and Implementation}.
+This section explains design and environmental conditions for the aggregations, key indicators and algorithms.
+It also dives into implementation details on the Apache Spark platform.
 \par
-\textbf{Design}. This section explains design and environment conditions for the algorithm to be useful. \par
-\textbf{Implementation}. This section explains algorithm implementation details and data analytics software that was used - Apache Spark\footnote{\url{https://spark.apache.org}}. \par
-\textbf{Evaluation}. The proposed algorithmic approach is applied to Google Trace data~\cite{clusterdata:Reiss2011}. \par
-\textbf{Conclusion and Future Work}. The completed data analysis is discussed. Further work related to algorithm improvements, automation and lessons learned are presented.
+\textbf{Evaluation}. The proposed map-reduce jobs and algorithms are evaluated on Google cluster-usage trace data. \par
+\textbf{Conclusion and Future Work}. The suggestion results data and future work related to algorithm improvements are presented.
 
@@
 \myparagraph{Conclusion}
-It works and is great!
+We proposed an uptime-based approach for capacity planning in Cloud environments.
+
+At first, we selected metrics required to make data-driven decisions.
+We applied transformations and aggregations to extract relevant data on a daily, per machine basis.
+Based on that, we calculated and applied our proposed key indicators -~\textit{cph, ip(scheme1, scheme2), KDI}- and assessed the state of Google trace data's optimality.
+It turned out that based purely on~\textit{KDI} values, around 7-8\% of machines could benefit from reconfiguration.
+We used this data as an input to our two suggestion algorithms.
+
+Squeeze-them-in algorithm suggested that there is enough spare capacity to power off non-optimal machines.
+As a result, the total cost dropped by 8.04\% and 9.45\% for both analyzed pricing configurations.
+Obviously, this came with a trade-off related to a higher utilization of the cluster.
+CPU and memory averages increased by around 7\%.
+There is not enough data to state whether this is acceptable or not.
+We would have to make assumptions, therefore the final evaluation of the algorithm's usefulness is delegated to the user and his specific requirements.
+
+New capacity configuration algorithm suggested alternative VM arrangements.
+Some of them turned out to be worse and got excluded from further analysis.
+It is worth to note that suggestions that were actually better, in this context - cheaper, contained only Heavy instances for both analyzed schemes.
+They yielded improvements in a range from 8 to 23\% compared to the previous non-optimal configuration and from 0.0065 to 0.022\% in relation to the total cost.
+
+It is somewhat safe to assume that when~\textit{KDI} is negative, a configuration is non-optimal.
+It indicates that there are more machines that report negative~\textit{KDI} values than positive and it is something worth to investigate.
+However, when cluster-wide~\textit{KDI} equals 0 or is positive, the number itself does not represent a lot of insight.
+There might be a situation when half of the machines run in a quite optimal way and yield high~\textit{KDI} values and, at the same time, the other half is non-optimal.
+As a result of that, we cannot tie a positive~\textit{KDI} value to indicate an optimal configuration.
+A positive value is better than negative, however, even this statement was not formally proved and can yield surprising results.
+Nevertheless, we discovered a correlation between~\textit{KDI} increase and total cost decrease.
+It is not a linear correlation - higher~\textit{KDI} increases does not seem to cause bigger cost decreases.
+This is an exciting discovery that can be validated further in future work using bigger data sets and different environments.
+The fact these two metrics correlate can be used to validate new suggestion algorithms.
 
 \myparagraph{Future Work}
+
+Each aggregation, key indicator and algorithm was designed as a proof of concept and can be developed and evaluated further.
+However, there are few key contributions that would make this thesis more valuable in real-life scenarios.
+
 \begin{enumerate}
-\item Visual recognition of underperforming machines using horizon charts (http://vis.berkeley.edu/papers/horizon/)
-\item Create an optimality model and adjust new capacity suggestions to find the best combination
-\item Predict growth trend and adjust configuration suggestions
+\item Capacity growth trend calculation. Update algorithms to suggest optimal configurations using the trend. 
+\item Predict price changes and discounts. Moore's Law still works~\cite{7057609} and optimizations that affect future should take this into account.
+\item Use horizon charts\footnote{\url{http://vis.berkeley.edu/papers/horizon}} to visualize~\textit{KDI} values.
+It would be an improvement to the threshold-based report as there might be a set of machines just below the threshold that are not recognized by the report.
+\item Research whether~\textit{KDI} correlates to a total cost in bigger sets of data and various pricing environments. If this was true, KDI could be used to validate suggestion algorithms results.
 

### 2015-04-26 paper/paper.tex
abstract

@@
 
-\emph{ /* To be revised. */} \par
 \textsl{New Infrastructure as a Service solutions are becoming available with a growing number of supported pricing models.
-More often than not, a hosted Cloud environment is used to design and build an infrastructure for a product.
-The recent availability of different pricing schemes based on resource utilization and uptime reveals new challenges in already unpredictable capacity planning process.
-There is a choice between ad-hoc provisioning and upfront payments with reduced hourly rates.
-Reserved instances charged upfront are categorized into three groups: light, medium and heavy.
-Which one is better for a given utilization model?
-When exactly does one pricing scheme becomes more cost effective?
-Determining which machine type is better for a given utilization model, or at which point the cost effectiveness of a pricing scheme changes, is vital for the companies subscribing to the IaaS.
+A hosted Cloud environment can be a choice to design and build an infrastructure for a product.
+The recent availability of different pricing schemes based on resource utilization and uptime reveals new challenges in an already complex capacity planning process.
+There is a choice between ad-hoc provisioning and time commitment that comes with reduced hourly rates.
+When one decides for a time commitment, there are still further pricing options available.
+Few different prices for the same resource that depend on a daily uptime.
+Which one is the best one?
+When exactly does one pricing scheme becomes more cost effective that the other?
+We attempted to answer these critical questions by i) aggregating and transforming metrics data, ii) designing key optimality indicators, iii) using the indicators to build algorithms that suggest better configuration options, iv) using an existing large dataset to evaluate the algorithms.
 }
@@
 
-We used modelling and experimentation methodology to build and evaluate the result.
+We used modelling and experimentation methodology to build and evaluate results.
 The SaaS provider perspective got represented by an availability of per task and per machine metrics.
@@
 
-\section{Input data schema transformation and aggregations}
+\section{Schema transformations and aggregations}
 
@@
 \item Predict price changes and discounts. Moore's Law still works~\cite{7057609} and optimizations that affect future should take this into account.
+\item Calculate a cost of migration to a newly suggested configuration. This cost should be added to a total cost that is calculated.
 \item Use horizon charts\footnote{\url{http://vis.berkeley.edu/papers/horizon}} to visualize~\textit{KDI} values.

### 2015-04-26 paper/paper.tex
review changes

@@
 %One of first things your examiners will look at is your literature review. If they see a
-%good number of journal papers, a swathe of conference papers, a few recent workshop
-%papers and not too many web sites, they’ll already be impressed.
+%good number of journal thesiss, a swathe of conference thesiss, a few recent workshop
+%thesiss and not too many web sites, they’ll already be impressed.
 %%%
@@
 In this thesis, we attempt to answer these questions in a cost-aware fashion considering uptime-based pricing models.
+It is a new consideration that adds uptime parameter to the standard resource-aware provisioning techniques~\cite{Bartolini:2014:AFC:2658949.2637480}.
 
@@
 The former is crucial for SaaS providers as they derive profits from the margin between the operational costs of infrastructure and revenue generated from customers~\cite{efficient_resource_allocation}.
-In other words, having this set right, gives a competitive advantage and greatly contributes to a healthy business.
+In other words, having this set right gives a competitive advantage and greatly contributes to a healthy business.
 The latter proved to yield savings up to 9.45\% that can be invested in other aspects of a business.
@@
 Main challenges include dealing with dynamic changes of pricing schemes offered by IaaS providers.
-Even though the thesis uses Amazon's related taxonomy, it should be read as an attempt to solve a general case for uptime-based variable pricing capacity planning.
+Even though this thesis uses Amazon's related taxonomy, it should be read as an attempt to solve a general case for uptime-based variable pricing capacity planning.
 Other challenging aspect is a cost of migration to suggested environment.
@@
 
-In this thesis, we look from the perspective of Software as a Service (SaaS) provider that uses an Infrastructure as a Service (IaaS) provider's services to host a product. \\
+We look from the perspective of Software as a Service (SaaS) provider that uses an Infrastructure as a Service (IaaS) provider's services to host a product. \\
 SaaS provider seeks i) predictable cost estimates and ii) the best value for money.
-The latter for obvious reasons - every healthy business is interested in minimizing costs.
+The latter for obvious reasons - every healthy business is interested in minimizing the costs.
 The former is needed to set a right product pricing.
@@
 \\
-IaaS provider pricing is flexible as described in below sections.
-One instance configuration can have many prices based on daily uptime.
-Cost-aware uptime-based provisioning becomes a new consideration and adds one more parameter to standard resource-aware provisioning techniques~\cite{Bartolini:2014:AFC:2658949.2637480}.
+IaaS provider pricing can be flexible as described in below sections. One instance configuration can have many prices based on a daily uptime. SaaS provider faces the challenge to pick the best one.
 
@@
 
-This paper focuses around the concept of Reserved Instances.
+This thesis focuses around the concept of Reserved Instances.
 The scheme is further split into Light, Medium and Heavy options.
@@
 \\
-This scheme got simplified on December 2014~\cite{AWS:pricing_change} and naming used in this paper does not match the latest offering anymore.
+This scheme got simplified on December 2014~\cite{AWS:pricing_change} and naming used in this thesis does not match the latest offering anymore.
 However, the idea of having the same instance configuration priced differently based on uptime is still valid and is used by other IaaS providers, e.g.
@@
 
-Paper~\cite{6253563} proposed a pricing framework for cloud services using game theory and data mining techniques.
+Authors of the paper~\cite{6253563} proposed a pricing framework for cloud services using game theory and data mining techniques.
 The framework determines the price based on recent usage data and available resources.
@@
 Specifically, they developed a method to predict provisioning demand based on selected time series metrics.
-Although prediction is not included in this paper, it would be a next logical step for future work to better address real-life use cases.
-
-This paper~\cite{6963393} studied IaaS consumer's optimal decisions in the monopoly and multiple IaaS providers market.
+Although prediction is not included in this thesis, it would be a next logical step for future work to better address real-life use cases.
+
+IaaS consumer's optimal decisions were studied in the monopoly and multiple IaaS providers market~\cite{6963393}.
 Based on rational customer decisions, authors derived an optimal solution to provider's pricing problem.
@@
 
+A new algorithm was proposed to generate a resource provisioning scheme which minimizes the cost for customers and satisfies the application needs using branch and bound method~\cite{6295066}.
+Their main focus is to avoid over-provisioning and find the best resource configuration for given requirements.
+This thesis assumes that this work is already done and SaaS provider looks for the best pricing scheme configuration for an already known resource configuration.
+
 %%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Taxonomy}
+This section explains a terminology being used for further design and evaluation stages.
+It is important to re-define some terms, like cost per hour, which is sometimes presented by IaaS providers in a non-complete way and may not include all costs.
+We also define new terminology that will be used throughout this thesis.
 
@@
 
+The section contains design and implementation details for an end-to-end approach that analyzes machine configuration and proposes changes.
+We start with metrics data prerequisites and guide through data transformations.
+We further define key indicators and formalize their implementations.
+The indicators can be used to filter non-optimal configurations.
+Algorithms that take the filter data as an input and generate configuration suggestions are presented later.
+
 \subsection{Prerequisites}
@@
 In order to make data-driven analysis and suggestions for optimization, we need data.
-The below list of metrics was compiled based on algorithm needs to answer critical questions regarding the state of infrastructure deployment.
-The list is suitable to guarantee performance SLAs for compute nodes.
-Machines with significant I/O usage profile are not covered in this paper.
+We compiled a list of metrics that will be used as a source of truth to determine configuration's state.
+The list is suitable to measure performance SLAs for compute nodes.
+Machines with significant I/O usage profile are not covered in this thesis.
 Metrics can be gathered periodically, per process or per job.
@@
 Collection per job or per process produces one log line and has a scope of a measurement period.
-Such structured data would not help to answer machine-wide neither cluster-wide questions.
+Such structured data would not help to answer machine-wide nor cluster-wide questions.
 The data has to be post-processed to allow calculation of below-listed aggregations and analytics.
@@
 To comply with IaaS pricing strategies, the data is analyzed in terms of a daily usage. This requires splitting metrics into buckets of one day worth of data. 
-It is achieved by applying a simple Apache Spark filter presented in Listing~\ref{daily_usage} that calculates day number based on metric timestamps.
-
+It is achieved by applying a simple Apache Spark filter presented in Listing~\ref{daily_usage} that calculates an integer to indicate a day number based on a metric timestamp.
+The implementation can vary for various timestamp formats.
+Results can be saved in separated folders or tables and all other aggregations and transformations are executed in scope of 1 day worth of data.
+
+\begin{minipage}{\textwidth}
 \begin{lstlisting}[label={daily_usage},caption={Daily usage filter},frame=single]
@@
 for x in range(0, days_range):
-    split_by_day = distFile.filter(third_filter)
+    split_by_day = distFile.filter(daily_filter)
 
 \end{lstlisting} 
+\end{minipage}
 
@@
 This metric indicates a daily number of hours of machine being up.
-It is a key metric that is be used to calculate optimality indicators discussed later on.
+It is a key metric that is be used to calculate optimality indicators discussed in the next three following paragraphs.
 Algorithm~\ref{alg:uptime_per_machine} represents pseudo-code implementation and Listing~\ref{uptime_per_machine_implementation} contains implementation in Python with Apache Spark.
 The data is extracted from timestamp values previously grouped into daily data buckets.
+By design, this approach extracts start and end time of jobs recorded in metrics using map function (lines 1-2).
+Then, two reduce jobs are executed that find the earliest start time (lines 3-4) and the latest end time (lines 5-6).
+As a result, we get two timestamps that indicate a range of a daily uptime.
+This design is flawed in a sense that if there are only two jobs logged - one in the morning and one in the evening - it will assume that machine was up no matter what happened in between.
 
@@
 Algorithm~\ref{alg:cost_per_hour} represents pseudo-code and Listing~\ref{cost_per_hour} shows implementation in Python.
-The algorithm returns a mapping of number of uptime hours per day and corresponding~\textit{cph} values.
+The algorithm takes upfront cost, charge per hour and daily uptime as input parameters.
+It aggregates both costs and calculates a real (total) cost per hour~\textit{if} machine were to run every day for a number of previously specified hours.
+We use this algorithm to build lookup tables for a 24 hour usage period.
+Essentially, we pass the same pricing scheme parameters and specify daily uptime from 1 to 24.
+This creates a table with hours per day to~\textit{cph} mapping.
 
@@
 At 18 hours, the point of scheme cost intersection (highlighted as 1), it does not matter which scheme is chosen as cost is the same. This is point 0.
-
 At each side of the intersection point either scheme1 or scheme 2 is more preferable.
-
 For example, a machine running for 20 hours on scheme1 is running 2 hours outside our optimum 18 hour scheme cost intersection.
 This gives us a score of -2 for this machine compared to the same machine running on scheme 2 where the benefits continue to increase the longer the machine is running.
-
 The same is true for a machine running for 10 hours on scheme 2.
@@
 A user can catch this situation by having a threshold value set to 0.
-However, there might be a situation when machine is taken for maintenance and reports negative values for one or couple of days.
+However, there might be a situation when machine is taken for maintenance and reports negative values for one or a couple of days.
 To avoid this kind of noise, a user can add an another criteria to specify a minimum number of days of reported values below the threshold.
@@
 Algorithm~\ref{alg:agg_cpu_mem} represents pseudo-code implementation and~\ref{agg_cpu_mem_implementation} contains implementation in Python with Apache Spark.
+The algorithm takes metrics data grouped by keys that represent machine IDs and list of values as an input.
+It extracts CPU and memory usage using map function (lines 1-2).
+Then, uses the same reduce function to calculate a sum for each of the metrics (lines 3-4) per machine.
 
@@
 
-The number of jobs per day will be a useful parameter to calculate cluster-wide averages used for suggestions.
+The number of jobs per day is a useful parameter to calculate cluster-wide averages used for suggestions.
 Algorithm~\ref{alg:jobs_per_day} represents pseudo-code implementation and~\ref{jobs_per_day} contains implementation in Python with Apache Spark.
+Similarly to the previous algorithms, this one takes metrics data grouped by keys that represent machine IDs and list of values as an input.
+It extracts a task usage information using map function (lines 1-2).
+Then, it uses a reduce function to calculate an aggregated cluster value (lines 3-4) per machine.
 
@@
 The proposed strategy calculates an amount of spare CPU and memory capacity based on cluster-wide averages.
-Aggregated value for CPU and memory represents an amount of capacity gathered from machines that are utilized below the average.
-Based on the amount of aggregated capacity, the algorithm will attempt to squeeze in a maximum amount of machines returned by threshold-based report.
-As a result, it proposes to turn off a number of machines and that can fit in calculated below-average capacity.
+An aggregated value for CPU and memory represents an amount of capacity gathered from machines that are utilized below the average.
+Based on the amount of an aggregated capacity, the algorithm attempts to squeeze in a maximum amount of machines returned by a threshold-based report.
+As a result, it proposes to turn off some number of machines that can fit in the previously calculated below-average capacity.
 \\
@@
 \\
-A pseudo-code implementation of this idea is presented in Algorith~\ref{alg:squeeze-them-in}.
+A pseudo-code implementation of this idea is presented in Algorithm~\ref{alg:squeeze-them-in}.
 
@@
  \algrenewcommand\algorithmicprocedure{\textbf{method}}
-
+ \begin{algorithmic}[1]
   $float spareCpu = 0;$
@@
 \Return machinesToExclude;
-
+\end{algorithmic}
 \end{algorithm}
@@
 Google trace data contains a lot of detailed information.
-This paper focuses on uptime and we extracted only relevant data for further evaluation.
+This thesis focuses on uptime and we extracted only relevant data for further evaluation.
 Machine and task events original schema is presented in Tables~\ref{tab:machine_events} and~\ref{tab:task_events} and fields selected for further transformation and aggregation are marked in bold text.
@@
 \item Calculate a cost of migration to a newly suggested configuration. This cost should be added to a total cost that is calculated.
-\item Use horizon charts\footnote{\url{http://vis.berkeley.edu/papers/horizon}} to visualize~\textit{KDI} values.
+\item Use horizon charts\footnote{\url{http://vis.berkeley.edu/thesiss/horizon}} to visualize~\textit{KDI} values.
 It would be an improvement to the threshold-based report as there might be a set of machines just below the threshold that are not recognized by the report.

### 2015-04-26 paper/paper.tex
final version

@@
 \listoffigures\newpage
-\par{\textbf{List of tables}}
+\listoftables
 
@@
 \textsl{New Infrastructure as a Service solutions are becoming available with a growing number of supported pricing models.
-A hosted Cloud environment can be a choice to design and build an infrastructure for a product.
+A hosted Cloud environment is a strong choice to design and build an infrastructure for a product.
 The recent availability of different pricing schemes based on resource utilization and uptime reveals new challenges in an already complex capacity planning process.
 There is a choice between ad-hoc provisioning and time commitment that comes with reduced hourly rates.
-When one decides for a time commitment, there are still further pricing options available.
+When one decides for a time commitment, there are still more pricing options available.
 Few different prices for the same resource that depend on a daily uptime.
@@
 \newpage
-
-Thanks to my wife Karen for her constant support and patience during the research and writing
-of this Thesis.
-Also, thanks to Prof. John Murphy and Viliam Holub for their support, help and assistance.
+\chapter*{Acknowledgements}
+
+I would like to thank Prof. Liam Murphy for help and guidance.
+
+I thank Christina Thorpe for many insightful conversations during the development of the ideas in this thesis. And for her time that I should use more.
+
+I also wish to express my gratitude to Alan for inspiration to do this research.
+
+I owe more than thanks to my lovely Klaudia for her support and constant encouragement during the entire process.
+
+And finally, but not least, thanks goes to my parents who have always motivated me to achieve more.
 
@@
 IaaS consumer can face a situation when resource-based capacity is known and it is time to rent an optimal machine configuration from the provider.
-It raises few open ended questions, including i) how many VMs is needed?
+It raises few open-ended questions, including i) how many VMs is needed?
 ii) what kind of VMs and which pricing scheme?
@@
 Most of the related work is conducted from an IaaS provider perspective~\cite{6274129}~\cite{6253563} or aimed to find an optimal pricing state for providers and consumers to collaborate~\cite{6676685}~\cite{6963393}.
-This thesis is different in a way that it analyses a SaaS provider capacity planning in an environment when one resource have different prices tightly coupled with an uptime.
+This thesis is different in a way that it analyses a SaaS provider capacity planning in an environment when one resource has different prices tightly coupled with an uptime.
 
@@
 
-Our proposed solution implements listed objectives using mainly map-reduce jobs executed on Apache Spark\footnote{\url{https://spark.apache.org}} platform.
+Our proposed solution implements the objectives using mainly map-reduce jobs executed on Apache Spark\footnote{\url{https://spark.apache.org}} platform.
 Aggregations, key indicators and algorithms were designed, implemented in pseudo-code and Python and evaluated on Google cluster-usage trace~\cite{clusterdata:Reiss2011}.
@@
 Even though this thesis uses Amazon's related taxonomy, it should be read as an attempt to solve a general case for uptime-based variable pricing capacity planning.
-Other challenging aspect is a cost of migration to suggested environment.
+Another challenging aspect is a cost of migration to suggested environment.
 It was not considered in the results, however, it created an opportunity for future work to investigate an impact of the delta difference between an old and new configuration cost.
-Lastly, we hit a technical issues related to implementation and execution of aggregations and algorithms on large data sets.
-Jobs were failing due to not enough RAM capacity on a single machine.
-The issues were resolved by a combination of code improvements, like changing the way aggregations build hash-maps and an increase of parallelism and chunking granularity on the platform.
+Lastly, we encountered technical issues related to implementation and execution of aggregations and algorithms on large data sets.
+Jobs were failing due to not enough of RAM capacity on a single machine.
+The issues were resolved by a combination of code improvements, like changing the way aggregations build hashmaps and an increase of parallelism and chunking granularity on the platform.
 
@@
 \myparagraph{Structure of the Thesis}\\
-\textbf{Background}. This section outlines the background information related to IaaS consumer point of view and challenges related to finding the best configuration for the money. \par
+\textbf{Background}. This section outlines the background information related to IaaS consumer's point of view and challenges related to finding the best configuration for the money. \par
 \textbf{Design and Implementation}.
-This section explains design and environmental conditions for the aggregations, key indicators and algorithms.
+This section explains the design and environmental conditions for the aggregations, key indicators and algorithms.
 It also dives into implementation details on the Apache Spark platform.
@@
 \\
-IaaS provider pricing can be flexible as described in below sections. One instance configuration can have many prices based on a daily uptime. SaaS provider faces the challenge to pick the best one.
+IaaS provider's pricing can be flexible as described in below sections. One instance configuration can have many prices based on a daily uptime. SaaS provider faces the challenge to pick the best one.
 
@@
 \paragraph{On-demand instances.} Allow to pay by the hour with no other commitments. The most expensive option.
-\paragraph{Spot instances.} A real-time bidding platform. A customer can bid on price. Instance can run only when the bid exceeds current Spot Price.
-\paragraph{Reserved instances.} Discounted price in comparison to On-demand in exchange of 1 or 3 years commitment. Guaranteed availability of reserved capacity.
+\paragraph{Spot instances.} A real-time bidding platform. A customer can bid on the price. An instance can run only when the bid exceeds current Spot Price.
+\paragraph{Reserved instances.} Discounted price in comparison to On-demand in exchange for 1 or 3 years of commitment. Guaranteed availability of reserved capacity.
 
 This thesis focuses around the concept of Reserved Instances.
-The scheme is further split into Light, Medium and Heavy options.
+The scheme is split further into Light, Medium and Heavy options.
 They differ in upfront and hourly costs.
@@
 There is no concept of reservation nor bidding.
-All of the options are priced within pay-as-you-go model with 1 minute granularity.
+All of the options are priced within pay-as-you-go model with 1-minute granularity.
 However, Google offers increasing discounts for sustained usage - 20\% for 25-50\% of month usage, 50-75\% of usage yields 40\% and everything above 75\% is priced with 60\% of discount rate.
@@
 Secondly, they studied bidding for the spot instances as a Stackelberg game.
-Although not focused on reserved instances, authors developed the system model that might serve as an inspiration for future work to provide suggestions based on application specific QoS metrics.
+Although not focused on reserved instances, authors developed the system model that might serve as an inspiration for future work to provide suggestions based on application-specific QoS metrics.
 
@@
 This section explains a terminology being used for further design and evaluation stages.
-It is important to re-define some terms, like cost per hour, which is sometimes presented by IaaS providers in a non-complete way and may not include all costs.
+It is important to redefine some terms, like cost per hour, which is sometimes presented by IaaS providers in a non-complete way and may not include all costs.
 We also define new terminology that will be used throughout this thesis.
@@
 
-The section contains design and implementation details for an end-to-end approach that analyzes machine configuration and proposes changes.
+This section contains design and implementation details for an end-to-end approach that analyzes machine configuration and proposes changes.
 We start with metrics data prerequisites and guide through data transformations.
@@
 The indicators can be used to filter non-optimal configurations.
-Algorithms that take the filter data as an input and generate configuration suggestions are presented later.
+Finally, we present algorithms that take the filter data as an input and generate configuration suggestions.
 
@@
 
-The following list of metrics is required.
+The following list presents required metrics.
 \myparagraph{Machine Id}
@@
 
-Metrics should be aggregated in \textless K, List\textless V\textgreater\textgreater format where K represents Machine Id and List of V is a list of measurements in a given period.
+Metrics should be aggregated in \textless K, List\textless V\textgreater\textgreater format where K represents Machine Id and List of V is a list of measurements for a given period.
 
@@
 
-Metrics gathered in Prerequisites section have limited knowledge about the environment.
+Metrics gathered in Prerequisites Section have limited knowledge about the environment.
 Collection per job or per process produces one log line and has a scope of a measurement period.
@@
 The implementation can vary for various timestamp formats.
-Results can be saved in separated folders or tables and all other aggregations and transformations are executed in scope of 1 day worth of data.
+Results can be saved in separated folders or tables and all other aggregations and transformations are executed in the scope of 1 day worth of data.
 
@@
 
-This metric indicates a daily number of hours of machine being up.
+This metric indicates a daily number of hours of a machine being up.
 It is a key metric that is be used to calculate optimality indicators discussed in the next three following paragraphs.
@@
 As a result, we get two timestamps that indicate a range of a daily uptime.
-This design is flawed in a sense that if there are only two jobs logged - one in the morning and one in the evening - it will assume that machine was up no matter what happened in between.
+This design is flawed in a sense that if there are only two jobs logged - one in the morning and one in the evening - it will assume that a machine was up no matter what happened in between.
 
@@
 The algorithm takes upfront cost, charge per hour and daily uptime as input parameters.
-It aggregates both costs and calculates a real (total) cost per hour~\textit{if} machine were to run every day for a number of previously specified hours.
-We use this algorithm to build lookup tables for a 24 hour usage period.
+It aggregates both costs and calculates a real (total) cost per hour~\textit{if} a machine were to run every day for a number of previously specified hours.
+We use this algorithm to build lookup tables for a 24-hour usage period.
 Essentially, we pass the same pricing scheme parameters and specify daily uptime from 1 to 24.
-This creates a table with hours per day to~\textit{cph} mapping.
+This creates a set with hours per day to~\textit{cph} mapping.
 
@@
 
-At 18 hours, the point of scheme cost intersection (highlighted as 1), it does not matter which scheme is chosen as cost is the same. This is point 0.
+At 18 hours, the point of scheme cost intersection (highlighted as 1), it does not matter which scheme is chosen as the cost is the same. This is point 0.
 At each side of the intersection point either scheme1 or scheme 2 is more preferable.
-For example, a machine running for 20 hours on scheme1 is running 2 hours outside our optimum 18 hour scheme cost intersection.
+For example, a machine running for 20 hours on scheme1 is running 2 hours outside our optimum 18-hour scheme cost intersection.
 This gives us a score of -2 for this machine compared to the same machine running on scheme 2 where the benefits continue to increase the longer the machine is running.
@@
 We are running at -8 below our optimum.
-In this case we would be more cost beneficial to run this machine on scheme1.
+In this case, we would be more cost beneficial to run this machine on scheme1.
 Giving us a score of +8.
@@
 A user can catch this situation by having a threshold value set to 0.
-However, there might be a situation when machine is taken for maintenance and reports negative values for one or a couple of days.
-To avoid this kind of noise, a user can add an another criteria to specify a minimum number of days of reported values below the threshold.
+However, there might be a situation when a machine is taken for maintenance and reports negative values for one or a couple of days.
+To avoid this kind of background noise, a user can add an another criterion to specify a minimum number of days of reported values below the threshold.
 
@@
 This approach has a potential to perform well in under-utilized clusters.
-For well-balanced environments with similar usage patterns, it does not find a lot of spare capacity and would propose fairly limited improvements.
+For well-balanced environments with similar usage patterns, it does not find a lot of spare capacities and would propose fairly limited improvements.
 \\
@@
  \KwResult{List of machines that can be powered down}
- \algrenewcommand\algorithmicfunction{\textbf{class}}
- \algrenewcommand\algorithmicprocedure{\textbf{method}}
  \begin{algorithmic}[1]
-  $float spareCpu = 0;$
-  $float spareMemory = 0;$
-  $ID[ ] machinesToExclude;$
-
-\ForAll{machine in otherMachines}{
-        \If{$machine.cpu < cluster.cpu$} {
+  \State $float spareCpu = 0;$
+  \State $float spareMemory = 0;$
+  \State $ID[ ] machinesToExclude;$
+  \State \ForAll{machine in otherMachines}{
+   \If {$machine.cpu < cluster.cpu$}{
   	 $spareCpu += clusterCpu - machineCpu$
@@
 	}
-
-\ForAll{machine in machinesFromReport}{
+  \State \ForAll{machine in machinesFromReport}{
 	  \If{$(spareCpu - machine.cpu > 0) \&\& (spareMemory - machine.memory > 0)$}{
-                add machine to machinesToExclude;
+	                $add machine to machinesToExclude;$
 	        }
 	}
-
-\Return machinesToExclude;
+\State\Return machinesToExclude;
 \end{algorithmic}
@@
 
+The algorithm takes CPU and memory cluster averages, a list of machines to optimize and list of the other machines as an input.
+First, it calculates the amount of spare capacity based on other machines data (code block 4).
+The results are saved in memory using variables listed in line 1 and 2.
+Then, it iterates through all of the machines that were identified to be non-optimal (code block 5) and if there is enough spare capacity it adds them to the exclusion list (line 3).
+Finally, it returns a list of machines that can be powered off.
+
 \subsubsection{New capacity configuration suggestion}
 
-New configuration is calculated based on input from the report, list of available schemes and total uptime to handle.
+The new configuration is calculated based on input from the report, list of available schemes and total uptime to handle.
 Pseudo-code implementation is presented in Algorithm~\ref{alg:new_configuration}.
-The algorithm iterates through all of available schemes, reads \textit{intersection points} and gets \textit{cph} for each of the \textit{ip}s.
+The algorithm iterates through all of the available schemes, reads \textit{intersection points} and gets \textit{cph} for each of the \textit{ip}s.
 It returns a total cost and number of machines required for each proposed configuration.
@@
  \KwResult{New configuration suggestion}
- \algrenewcommand\algorithmicfunction{\textbf{class}}
- \algrenewcommand\algorithmicprocedure{\textbf{method}}
-
-  $float[ ] totalCost;$
-
-\ForAll{scheme in availableSchemes}{
+ \begin{algorithmic}[1]
+  \State $float\{\} totalCost;$
+  \State\ForAll{scheme in availableSchemes}{
 		\ForAll{ip in scheme.intersectionPoints}{
-			vms = Uptime / ip.getHours() \\
-			cost = ip.getCphForHours(ip.getHours()) \\
-			totalCost[ip] = (vms, cost)
+			$vms = Uptime / ip.getHours()$ \par
+			$cost = ip.getCphForHours(ip.getHours())$ \par
+			$totalCost[ip] = (vms, cost)$\par
 		}
 }
-
-\Return{totalCost}
+\State \Return{totalCost}
+\end{algorithmic}
 \end{algorithm}
+
+The algorithm takes a list of machines from the report, list of available schemes and total uptime to accommodate.
+It iterates through available schemes (code block 2) and for each of them attempts to suggest a configuration for a given uptime.
+The suggestion process is split into phases 1) calculate a number of VMs required to handle requested uptime -  VM's uptime time depends on \textit{intersection point} value, 2) calculate cost of running the VM based on a corresponding \textit{cph} value, 3) add results to totalCost map (line 1).
+Finally, the algorithm returns a map of total cost and VMs required for each analyzed scheme.
 
@@
 \item suggestion of changes based on Algorithm~\ref{alg:squeeze-them-in} and~\ref{alg:new_configuration},
-\item validation of a positive correlation between \textit{KDI} and \textit{total cost} changes.
+\item validation of a correlation between \textit{KDI} and \textit{total cost} changes.
 \end{enumerate}
@@
 Amazon Web Services\footnote{\url{http://aws.amazon.com/}} offered variable pricing schemes based on uptime.
-Customer could choose an instance configuration and total cost would be calculated based on a number of hours of uptime per day.
+Customers could choose an instance configuration and the total cost would be calculated based on a number of hours of uptime per day.
 The offering was split into three options:
@@
 
-Table~\ref{tab:new_schema} is a new schema produced from join of capacity metrics from machine events and aggregations computed from tasks events data.
+Table~\ref{tab:new_schema} is a new schema produced from a join of capacity metrics from machine events and aggregations computed from tasks events data.
 
@@
 
-\myparagraph{Scheme1}
-\begin{table}[h]
+\begin{table}[H]
 \begin{center}
@@
 
-\myparagraph{Scheme2}
-
-\begin{table}[h]
+\begin{table}[H]
 \begin{center}
@@
 \end{center}
-\caption{Results of report run on cluster with criteria of KDI below 0 for 29 days}
+\caption{Results of the report run on the cluster with criteria of KDI below 0 for 29 days}
 \label{tab:threshold_based_report}
@@
 An interesting pattern raised - a medium is always the least optimal choice and heavy is the best.
-The suggestion to run heavy configuration for Scheme 1 would save 8.05\% compared to non-optimal machines set and 0.0065\% from total cost.
+The suggestion to run heavy configuration for Scheme 1 would save 8.05\% compared to non-optimal machines set and 0.0065\% to the total cost.
 Scheme 2 would save 23\% compared to non-optimal machines set and 0.022\% from total cost.
@@
 
-At first, we selected metrics required to make data-driven decisions.
-We applied transformations and aggregations to extract relevant data on a daily, per machine basis.
-Based on that, we calculated and applied our proposed key indicators -~\textit{cph, ip(scheme1, scheme2), KDI}- and assessed the state of Google trace data's optimality.
+First, we selected metrics required to make data-driven decisions.
+Then, we applied transformations and aggregations to extract relevant data on a daily, per machine basis.
+Finally, we calculated and applied our proposed key indicators -~\textit{cph, ip(scheme1, scheme2), KDI}- and assessed the state of Google trace data's optimality.
 It turned out that based purely on~\textit{KDI} values, around 7-8\% of machines could benefit from reconfiguration.
@@
 It indicates that there are more machines that report negative~\textit{KDI} values than positive and it is something worth to investigate.
-However, when cluster-wide~\textit{KDI} equals 0 or is positive, the number itself does not represent a lot of insight.
+However, when cluster-wide~\textit{KDI} equals 0 or is positive, the number itself does not represent a lot of insights.
 There might be a situation when half of the machines run in a quite optimal way and yield high~\textit{KDI} values and, at the same time, the other half is non-optimal.
@@
 
-Each aggregation, key indicator and algorithm was designed as a proof of concept and can be developed and evaluated further.
+Each aggregation, a key indicator and an algorithm was designed as a proof of concept and can be developed and evaluated further.
 However, there are few key contributions that would make this thesis more valuable in real-life scenarios.
@@
 \item Calculate a cost of migration to a newly suggested configuration. This cost should be added to a total cost that is calculated.
-\item Use horizon charts\footnote{\url{http://vis.berkeley.edu/thesiss/horizon}} to visualize~\textit{KDI} values.
+\item Use horizon charts\footnote{\url{http://vis.berkeley.edu/papers/horizon}} to visualize~\textit{KDI} values.
 It would be an improvement to the threshold-based report as there might be a set of machines just below the threshold that are not recognized by the report.
@@
 
-\label{endpage}
 \bibliography{bib}{}
 \bibliographystyle{plain}
+\label{endpage}
 \end{document}
